### 2015-03-16 draft.tex
initial commit

\documentclass[journal]{style/vgtc} 			          % final (journal style)
%\documentclass[review,journal]{style/vgtc}         % review (journal style)
%\documentclass[widereview]{style/vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{style/vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{style/vgtc}     % electronic version, journal
%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.
%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

% -- My Own Packages and Commands
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{\uline{#1}}}
\newcommand{\com}[1]{\textcolor{orange}{\uline{#1}}}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
%% TODO: Blind Review: Replace with online ID
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Interactive Visual Regression Cubes Analysis of Image-Centric Cohort Study Data}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernard Preim}
\authorfooter{
%% insert punctuation at end of each item
\item
 Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,niemann,preim\}@ovgu.de
\item
 Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
}

%other entries to be set up for journal
\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Interactive Visual Regression Cubes Analysis of Image-Centric Cohort Study Data}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{%
(The problem).
(New Solution).
(Validation).
(Results).
(Implications).
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Interactive Visual Analysis, Epidemiology, Breast Cancer, Hepatic Steatosis}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ % not used in journal version
 \CCScat{K.6.1}{Management of Computing and Information Systems}%
{Project and People Management}{Life Cycle};
 \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
}

%% Uncomment below to include a teaser figure.
  % \teaser{
  % \centering
  % \includegraphics[width=16cm]{CypressView}
  % \caption{In the Clouds: Vancouver from Cypress Mountain.}
  % }

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

\section{Epidemiological Background} \label{MedicalAndTechnicalBackground}
%%
This section covers the epidemiological workflow and requirements.

\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}

\subsection{Epidemiological Data} \label{EpidemiologicalData}

\subsection{The Study of Health in Pomerania (SHIP)}
After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
%
They slightly differ in clinical focus, acquired data and epidemiological research questions.
%
Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%
Data are collected without focus on a group of diseases.
%
This allows to query the data regarding many diseases and conditions.
%
Subjects were examined in a 5-year time span, continuously adding new parameters including MRI scans in the last iteration \cite{Hegenscheid2009}.
%
\section{Prior and Related Work}

\paragraph{Visual Analysis of Heterogenous Data.}
Zhang et al. \cite{Zhang2012} provide a web-based system for analyzing subject groups with linked views and batch-processing capabilities for categorizing new subject entries into the data set.
%
Their definition of a cohort differs from the understanding of the term in an epidemiological context by denoting every parameter-divided subject group as individual cohort.
%
Due to the short paper length, detail is missing on the data types and their algorithms of identifying similar subjects or whether they employ statistical measures.
%
We employ the idea of adding variables via drag and drop into a canvas area.

Generalized Pairs Plots (\texttt{GPLOM´S}) are an information visualization technique comparing heterogenous variables pairwise using a plot-matrix grouped by type \cite{GPLOMS, Francois2013}.
%%
They are useful to gain an overview over numerous variables and their distributions.
%%
Histograms, bar charts, scatter plots and heat maps are used to visualize variable combinations with regard to their type.
%%
The resulting matrix provides an \emph{overview visualization}, but requires a lot of screen space for many variables (127 in our application scenario).
%%
We incorporate the idea of adaptive type-dependent visualizations.
%%
Dai et al. \cite{Dai2005} explored risk factors by incorporating choropleth maps of epidemiological variables (e.g., mortality rates in a region) with parallel coordinates, bar charts and scatter plots with integrated regression lines.
%%
Their findings yielded a \emph{Concept Map}, which linked cancer-related associations via graph edges.
%%
While their goal to identify possible risk factors using socio-economic and health data is similar to ours, they focus on iteratively refining defined hypotheses and on geographical data.
%%
We employ the use of small multiples for incorporating heterogenous data types for comparability.
%%
Chui et al. \cite{Chui2011} visualized associations in time-dependent epidemiological data using time-series plots highlighting risk factor differences in age and gender.
%%
%The connection of time series plots with histograms and heat maps allows for fast insight into complex relationships and 
While the work shows how different visualization techniques provide insight into these data sets, it focuses on the time aspect, which is not present for our data.
%%
% *********************** Our own Work ***********************
% Prior Work
\paragraph{Prior Work.}
We visualized lumbar spine variabilities based on a semi-automatic shape detection algorithm of 490 participants of the \texttt{SHIP-2} cohort \cite{Klemm2013VMV}.
%
Hierarchical agglomerative clustering divided the population into shape-related groups.
%
As proof of concept, a relation between the size of the segmented shape and the measured size of the subjects was shown.
%
This work focuses on incorporating these derived data as new variables, enabling to include it into the hypothesis validation and generation process.
%
When applying clustering techniques to the non-image data it was found that \texttt{k-Prototypes} and \texttt{DBSCAN} are appropriate, but are strongly dependent on the chosen variables and distance measures \cite{Klemm2014BVM}.
%
Niemann et al. \cite{Niemann2014} presented an interactive data mining tool for the assessment of risk factors of hepatic steatosis, the fatty liver disease.
%
Association rules created by data mining methods can be analyzed interactively with their tool and highlight potentially overlooked variables.
% *********************** / Our own Work ***********************
\section{Regression Cubes} \label{Image Centric Cohort Study Data in Interactive Visual Analysis Context}
%%
\subsection{Analysis Workflow}
%%
\section{System Design and Implementation} \label{Interaction- and Visualization Techniques}
%
\subsection{Design and Visualization Techniques} \label{Structure and Workflow}

\paragraph{System Layout.}

\subsection{Implementation} \label{implementation}
%%
\com{Use VAST'14 as Guideline}
%
\section{Application} \label{application}
%
\subsection{The Breast Fat Data Set}

\subsubsection{Data Preprocessing} \label{application:Data Preprocessing}
The data processing follows the description in Section~\ref{Data Preprocessing}.
%

\subsection{Participants, Setup and Procedure}
%
\paragraph{Setup.} Due to the large geographical distance, the evaluation was done completely web-based.
%%
\paragraph{Procedure.}
%%
\com{Analyses here ...}
%%
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%

\section{Summary and Conclusion}
%%
\begin{small}
	\acknowledgments{Omitted due to blind review}
% \acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no. 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania. Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern. The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG. This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics. We thank Marko Rak and Klaus Toennies for providing the image detection data.}
\end{small}
\clearpage
\newpage
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{bibliography}
\end{document}

### 2015-03-16 draft.tex
first abstract draft

@@
 
-\title{Interactive Visual Regression Cubes Analysis of Image-Centric Cohort Study Data}
+\title{Regression Cube Analysis of Cohort Study Data}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Interactive Visual Regression Cubes Analysis of Image-Centric Cohort Study Data}
+\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Regression Cube Analysis of Cohort Study Data}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 \abstract{%
-(The problem).
-(New Solution).
-(Validation).
-(Results).
-(Implications).
+\com{Problem}.
+Epidemiological studies comprise of heterogenous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
+%%
+These data contain information (\emph{features}) about a subjects lifestyle, medical conditions and also medical images of the whole body.
+%%
+These features are analyzed using statistical regression analyses towards features indicating a disease (the \emph{target feature}).
+%%
+%These complex data are hard to analyze concurrently, only a few features can be considered in statistical analyses, such as regression analyses.
+These analyses usually include two to five features (\emph{independent features}) towards the disease of interest.
+\com{New Solution}.
+%(New Solution).
+We propose a new analyses approach of epidemiological data sets by incorporating all dimensions in a exhaustive regression-based analyses.
+%%
+It takes all possible combinations of \emph{independent features} towards a \emph{target feature} into account and provides a visualization, which allows for insights into the data by showing relationships possibly unknown to the domain expert.
+%%
+%The most meaningful features 
+A 3D-visualization of all possible combinations of two to three target variables acts as overview over the whole data set.
+%%
+We call it the \emph{Regression Cube}.
+%%
+Slicing through the regression cube allows for the detailed analysis of features towards the target disease.
+%%
+\com{Validation}.
+%%
+We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
+%%
+We conducted a qualitative analysis with three domain experts on a breast fat data set to derive insights into the correlation between breast lesions and non-image variables.
+%%
+\com{Results}. TODO.
+%%
+\com{Implications}.
+Epidemiological data sets can for the first time be visually overviewed using a regression-based analysis of all features towards a disease. 
+%With our technique it is possible for the first time to provide a overview visualization of variables regarding a target variable.
+
 } % end of abstract

### 2015-03-17 draft.tex
added introduction

@@
 %%
+The regression formula can be adjusted by the user to describe disease-specific hypotheses.
+%%
+Influences of features can be assessed using a difference view, comparing different calculation results.
 \com{Validation}.
@@
 %%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \maketitle
+%%
+Epidemiology aims to characterize health and disease conditions in defined populations (\emph{Cohorts}).
+%%
+Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
+%%
+They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide spread diseases.
+%%
+In the standard workflow, physicians translate observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.
+
+An important epidemiological tool for deriving such features are \emph{Cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
+%%
+To reduce any selection bias, subjects are invited at random and without a focus on a specific disease.
+%%
+The acquired features range from social and lifestyle factors to prior or current diseases and medications as well as medical parameters, such as blood pressure and also comprises of non-radiating medical image data.
+e.g.
+magnetic resonance imaging (MRI).
+%%
+Medical image data quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
+%%
+
+To assess the statistical resilience of a hypotheses using regression analyses rarely involves more than three features due to the required subject count.
+%%
+Due to missing overview techniques, possibly interesting correlations lie within the data, but are not made apparent.
+%%
+Explorative analyses and first overview visualizations of the data set as presented by Klemm et al.
+\cite{Klemm2014VIS} are not custom tailored to a specific target variable and mostly highlights correlations between variables, which are known to the domain expert (e.g.
+correlation between body size and spine shape).
+%%
+We incorporate the regression analysis, which is familiar to the domain experts into a overview visualizations, which can either be used for an hypothesis-free analysis or a analysis towards a specific disease or hypotheses.
+%%
+This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
+%%
+Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
+%%
+Difference views between regression formulas allow to assess the influences of individual variables in the process.
+
+Our contributions are:
+\begin{itemize}
+	\item An overview visualization technique describing feature interactions using target features.
+	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for detailed investigations of feature relationships.
+	\item freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
+	\item analysis of confounding variables by providing comparison views between different formula results
+	\item The open and web based approach of the system allows for analysis of any data using the presented method.
+\end{itemize}
 
@@
 %%
+\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction}
 \subsection{Analysis Workflow}
+\com{Different Formulas work as different steps in the analysis}
 %%

### 2015-03-18 draft.tex
first draft of whole structure

@@
 \com{Problem}.
-Epidemiological studies comprise of heterogenous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
+Epidemiological studies comprise heterogeneous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
 %%
-These data contain information (\emph{features}) about a subjects lifestyle, medical conditions and also medical images of the whole body.
+These data contain information (\emph{features}) about a subjects' lifestyle, medical conditions and also medical images of the whole body.
 %%
@@
 %(New Solution).
-We propose a new analyses approach of epidemiological data sets by incorporating all dimensions in a exhaustive regression-based analyses.
+We propose a new analyses approach of epidemiological data sets by incorporating all dimensions in an exhaustive regression-based analyses.
 %%
-It takes all possible combinations of \emph{independent features} towards a \emph{target feature} into account and provides a visualization, which allows for insights into the data by showing relationships possibly unknown to the domain expert.
+%It takes all possible combinations of \emph{independent features} towards a \emph{target feature} into account and provides a visualization, which allows for insights into the data by showing relationships possibly unknown to the domain expert.
+This approach combines all \emph{independent features} towards a \emph{target feature} and provides a visualization, which allows for insights into the data by showing the domain expert possibly unknown relationships.
 %%
 %The most meaningful features 
-A 3D-visualization of all possible combinations of two to three target variables acts as overview over the whole data set.
+%A 3D-visualization of all possible combinations of two to three target variables acts as overview over the whole data set.
+A 3D-visualization of all combinations of two to three target variables acts as an overview over the whole data set.
 %%
@@
 
+\subsection{Regression Analysis}
+\com{Regression cubes are an important tool for checking statistical resilience.}
+\com{Describe how regression cubes work (including notation).}
+\com{Move this to Regression Cube part?}
+
 \subsection{The Study of Health in Pomerania (SHIP)}
@@
 \section{Prior and Related Work}
-
+\com{From VAST'14 Paper:}
 \paragraph{Visual Analysis of Heterogenous Data.}
@@
 % *********************** / Our own Work ***********************
-\section{Regression Cubes} \label{Image Centric Cohort Study Data in Interactive Visual Analysis Context}
+\section{Regression Cube Analysis of Cohort Study Data}
 %%
-\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction}
+\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction.}
+
+\subsection{Target-Variable-dependent Dimension Reduction}
+\com{Explain the CFS algorithm and how it is applied to reduce the number of dimensions and therefore the calculation process.}
+
+\subsection{Cube Description using Regression Formula Notation}
+\com{Regression formulas are described using standard regression formula notation by introducing variables.}
+
+\subsection{Abstracting Regression Results using R$^2$}
+\com{R$^2$ are used to encode regression results to color spaces.}
+\com{R$^2$ is comparable between regression results of the same type, but not in between types. Therefore it is encoded using different colours.}
+
 \subsection{Analysis Workflow}
-\com{Different Formulas work as different steps in the analysis}
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=2.8in]{figures/placeholder}
+ \caption{
+ \com{Workflow for different analysis approaches.}
+ \com{a) Analysis without hypothesis about the data set (z ~ x+y)}
+ \com{b) Analysis without hypothesis with confounder (z ~ age)}
+ \com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
+ \com{d) Clarify using confounder (Var ~ x)}
+ }
+  \label{fig:Workflow}
+\end{figure}
+\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
 %%
 \section{System Design and Implementation} \label{Interaction- and Visualization Techniques}
-%
+%%
+
 \subsection{Design and Visualization Techniques} \label{Structure and Workflow}
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=2.8in]{figures/placeholder}
+ \caption{
+ \com{Screenshot of the cube visualization embedded in the framework}
+ \com{Also include active tooltip}
+ }
+  \label{fig:Cube}
+\end{figure}
+
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=2.8in]{figures/placeholder}
+ \caption{
+ \com{Screenshot of the cube visualization embedded in the framework}
+ }
+  \label{fig:comparison}
+\end{figure}
+%%
+\com{Simplicity of interface to allow for steep learning curve.}
+\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
+\com{Cube acts as visual mini map to give a global impression over the data.}
+\com{current pane shown using a heat map visualization.}
+\com{details on demand for selected regression formula.}
+\com{comparison view using reference cubes.}
 
@@
 \subsection{Implementation} \label{implementation}
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=2.8in]{figures/placeholder}
+ \caption{
+ \com{Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
+ }
+  \label{fig:Implementation}
+\end{figure}
 %%
 \com{Use VAST'14 as Guideline}
-%
+\com{Speed is merely a matter of available server machines due to parallelization process.}
+\com{Security enabled by hashing files.}
+\com{Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend (Fig.~\ref{fig:Implementation})}
+
 \section{Application} \label{application}
@@
 \subsection{The Breast Fat Data Set}
+\com{Describe how the data was acquired and preprocessed}
+
+\subsubsection{Data Preprocessing} \label{application:Data Preprocessing}
+The data processing follows the description in Section~\ref{Data Preprocessing}.
+%
+
+\subsection{The Hepatic Steatosis Data Set}
+\com{Uli: Describe the data set and also prior work on it.}
 
@@
 \subsection{Participants, Setup and Procedure}
+%%
+\com{Use of VDAR Technique}
 %
@@
 %%
-\com{Analyses here ...}
+\subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
+
+\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
 %%
@@
 %%
+\com{Feedback from the evaluation goes here.}
+\com{Time-aspect critical, interactive analysis requires for fast response. The method needs to be speeded up.}
 
 \section{Summary and Conclusion}
+\paragraph{Future Work.}
+\com{Implementing of regression analysis on the graphics card.}
 %%

### 2015-03-18 draft.tex
refined structure
added section od epidemiological background

@@
 This section covers the epidemiological workflow and requirements.
+%%
+\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
+Epidemiology unites experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
+%%
+Their goal is derive disease-specific risk factors by assessing epidemiological features using statistical methods.
+%%
+As described by \cite{Thew2009} is divided into three different steps:
+\begin{itemize}
+	\item Clinicians make observations in the daily practice which are translated into hypotheses.
+	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
+	\item Statisticians assess the association of the derived features towards the investigated disease.
+\end{itemize}
+%%
+Relative risks can be determined if a statistical resilient association of features towards a condition are extracted.
+%%
+They indicate the per-subject chance of developing the disease.
+%%
+Reproducibility of results is a epidemiological key requirement and guides all analyses steps.
+%%
+Statistical programs, such as \texttt{SPSS} are used to analyzed the data using the classical sequential epidemiological workflow, which uses images largely only to communicate results, rather than providing insight.
+%%
+A alternative data driven approach is described as follows.
 
-\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
+\subsubsection{Data Driven Hypothesis-Generation-based Analysis}
+Klemm et al. describe an approach for a Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
+%%
+Their methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
+%%
+Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
 
 \subsection{Epidemiological Data} \label{EpidemiologicalData}
+%%
+\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
+%%
+They yield a highly heterogenous and incomplete information space.
+%%
+Subject data are collected with the widest range possible, allowing the data set to be assessed towards different diseases.
+%%
+The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
+%%
+Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
+%%
+Other features are exclusive, such as women-specific questions, e.g. number of born children or period status.
+
+\paragraph{Data Types.}
+%%
+Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
+%%
+The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
+%%
+The data heterogeneity has to be taken into account when the analysis method is chosen.
+%%
+Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
+%%
+This is however avoided if possible, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
+%%
+
+\paragraph{Image Data.}
+%%
+Since the Rotterdam study, many modern cohort studies include medical image data.
+%%
+For ethical reasons, the imaging modalities must not include ionizing radiation.
+%%
+The image quality is often inferior to clinical standard, which is a tradeoff between time and cost \cite{Preim2014}.
+%%
+These data are hard to analyze as they require segmentation highlighting the structures of interest.
+%%
+This process is prone to inter and intra-observer variability when carried out manually.
+%%
+Automatic or semi-automatic solutions bypass this problem, but are costly and need to be custom tailored to the structure of interest \cite{Toennies2015}.
+%%
+Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
+
+\paragraph{Confounding Features}
+Features influencing the exposure as well as the outcome of a analysis are called confounders.
+%%
+The analysis model has to be adjusted by normalizing all included features towards the confounding feature.
+%%
+\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely to happen with increasing age.
+%%
+It also influences the general body condition and therefore almost all features acquired through cohort studies.
+%%
+Another important confounder is \emph{Gender}.
+%%
+Other confounder have to be selected by epidemiologists specific to the investigated condition.
 
 \subsection{Regression Analysis}
-\com{Regression cubes are an important tool for checking statistical resilience.}
-\com{Describe how regression cubes work (including notation).}
+%%
+Regression analyses are the most important statistical tool when analyzing epidemiological data.
+%%
+They are also the foundation of this work.
+%%
+A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
+%%
+The regression model yields a function describing the target feature by weighting the independent features.
+%%
+Metrics, such as the weightings itself and associated p values and describe the resulting function (the \emph{model}).
+%%
+R$^2$ values describe the goodness of fit; in other words how well the dependent features describe the target feature.
+%%
+The value is in the range [0,1], while 1 encodes a perfect fit.
+
+\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
+%%
+\begin{equation}
+Dependent \sim Independent_1 + Independent_2 + ... + Independent_n
+\end{equation}
+%%
+The most used regression operators comprise:
+%%
+\begin{itemize}
+	\item $+$ / $-$ inclusion/exclusion of the variable,
+	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
+	\item $*$ inclusion the variables as well as their interactions (e.g. $x*y$)
+	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
+\end{itemize}
+%%
+The type of the target feature restricts the regression type.
+
+\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating linear weightings for the \emph{independent} features.
+%%
+The dependent variable has to be of a continuous type.
+%%
+
+\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
+%%
+The target is described by fitting a logistic function.
+%%
+Logistic models do as opposed to linear models does not allow for extracting a R$^2$ goodness of fit value.
+%%
+Therefore, pseudo-R$^2$ values are extracted, such as the \emph{Nagelkerke R$^2$}, which mimics the behavior of the R$^2$.
+%%
+\emph{Nagelkerke R$^2$} cannot be compared to R$^2$ values extracted from linear regression model.
+
+%\com{Regression cubes are an important tool for checking statistical resilience.}
+%\com{Describe how regression cubes work (including notation).}
 \com{Move this to Regression Cube part?}
@@
 \subsection{The Study of Health in Pomerania (SHIP)}
-After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
+The \texttt{SHIP}, located in Northern Germany aims to aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+%%
+It does not focus on a specific disease, making the data set open for many diseases.
+%%
+Unique for the \texttt{SHIP} is the acquisition of medical image data per subject.
+%%
+A second cohort, \texttt{SHIP-TREND} was introduced in 2012.
+%%
+Data for both cohorts are examined in a in a 5-year time span.
+%%
+New parameters are added in each iteration, extending the range of investigated diseases.
+%%
+For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009}.
+%%
+\com{Include Hegenscheids Breast Fat Paper}
+%%
+%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
 %
-They slightly differ in clinical focus, acquired data and epidemiological research questions.
+%They slightly differ in clinical focus, acquired data and epidemiological research questions.
 %
-Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
 %
-Data are collected without focus on a group of diseases.
+%Data are collected without focus on a group of diseases.
 %
-This allows to query the data regarding many diseases and conditions.
+%This allows to query the data regarding many diseases and conditions.
 %
-Subjects were examined in a 5-year time span, continuously adding new parameters including MRI scans in the last iteration \cite{Hegenscheid2009}.
 %
@@
  \com{Workflow for different analysis approaches.}
+ \com{0. Select the Variables.}
+ \com{1. Select the Formulas:}
  \com{a) Analysis without hypothesis about the data set (z ~ x+y)}
@@
  \com{d) Clarify using confounder (Var ~ x)}
+ \com{3. Find interesting pane.}
+ \com{4. Get Details on Demand.}
  }

### 2015-03-19 draft.tex
revised abstract with Kais input

@@
 %%
-%These complex data are hard to analyze concurrently, only a few features can be considered in statistical analyses, such as regression analyses.
-These analyses usually include two to five features (\emph{independent features}) towards the disease of interest.
+Overview visualizations over a whole data sets towards a target feature are highly required, but not available to epidemiologists.
+%%
+%These analyses usually include two to five features (\emph{independent features}) towards the investigated condition.
 \com{New Solution}.
 %(New Solution).
-We propose a new analyses approach of epidemiological data sets by incorporating all dimensions in an exhaustive regression-based analyses.
+
+We propose a new analyses approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
 %%
-%It takes all possible combinations of \emph{independent features} towards a \emph{target feature} into account and provides a visualization, which allows for insights into the data by showing relationships possibly unknown to the domain expert.
-This approach combines all \emph{independent features} towards a \emph{target feature} and provides a visualization, which allows for insights into the data by showing the domain expert possibly unknown relationships.
+This approach combines all \emph{independent features} towards a \emph{target feature} and provides a visualization, which allows for insights into the data by highlighting relationships. %showing the
 %%
-%The most meaningful features 
-%A 3D-visualization of all possible combinations of two to three target variables acts as overview over the whole data set.
-A 3D-visualization of all combinations of two to three target variables acts as an overview over the whole data set.
+A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview over the whole data set, the \emph{Regression Cube}.
 %%
-We call it the \emph{Regression Cube}.
+%We call it the \emph{Regression Cube}.
 %%
-Slicing through the regression cube allows for the detailed analysis of features towards the target disease.
+Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
 %%
-The regression formula can be adjusted by the user to describe disease-specific hypotheses.
+The regression model can be adjusted by the user to describe disease-specific hypotheses.
 %%
-Influences of features can be assessed using a difference view, comparing different calculation results.
+Furthermore, the influences of features can be assessed using a difference view, comparing different calculation results.
 \com{Validation}.
@@
 %%
-\com{Results}. TODO.
+\com{Results}.
+%%
+With our technique we were able to reproduce results from a data mining analysis of a hepatic steatosis data set.
+%%
+By applying it to a breast fat data set we were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
 %%
@@
 \begin{itemize}
-	\item An overview visualization technique describing feature interactions using target features.
+	\item A overview visualization technique describing feature interactions using target features.
 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for detailed investigations of feature relationships.
@@
 %%
-For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009}.
-%%
-\com{Include Hegenscheids Breast Fat Paper}
+For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
 %%

### 2015-03-19 draft.tex
added first part of Regression Cube section

@@
 
-\subsection{Epidemiological Data} \label{EpidemiologicalData}
+\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
 %%
@@
 
-\subsection{Regression Analysis}
+\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
 %%
@@
 Dependent \sim Independent_1 + Independent_2 + ... + Independent_n
+\label{eq:RegressionNotation}
 \end{equation}
@@
 %%
-\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction.}
+The basic idea of our \emph{Regression Cube} is to provide a overview visualization of large cohort study data sets towards target variables.
+%% TODO BLIND
+Overview visualizations of feature relationships as presented by Klemm et al. \cite{Klemm2014VIS} are often focused on relationships between the visualized features.
+%%
+Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
+%%
+In epidemiology, these relationships are also of interest, but rather towards their explanatory power towards the target feature.
+%%
+These target features often indicates the presence of the investigated disease.
+%%
+As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
+%%
+A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
+%%
+\subsection{Cube Description using Regression Formula Notation}
+%%
+Expert knowledge is introduced into regression analyses using the regression formulas.
+%%
+As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
+%%
+%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
+%%
+
+Since we want to use the regression analyses with a overview visualization, we are are interested in all possible combinations of (two or more) independent features describing a target.
+%%
+We achieve this by introducing dynamic variables $X$, $Y$ and $Z$ into the regression notation.
+%%
+The method then replaces the dynamic variables with all features in the data set.
+%%
+In a data set with 100 features, the regression formula
+\begin{equation}
+Cancer \sim X + Y
+\end{equation}
+would yield 10.000 regression models, describing all possible combinations of two features in the data describing the target $Cancer$.
+%%
+The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, because it uses the same notation, all operators can be used as before.
+%%
+This allows for a fast adaptation and in the epidemiological application domain.
+%%
+With simple adjustments to the formula, different results can be achieved:
+%%
+\begin{itemize}
+	\item $Z \sim X + Y$ calculates all combinations of two features towards all possible target features.
+	\item $Cancer \sim X + Y + BodyWeight$ includes the $BodyWeight$ feature into all regression models as feature.
+	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features towards the target.
+\end{itemize}
+%%
+The problem with this brute-force approach lies in its complexity.
+%%
+The number calculated regression models increases exponentially for each dynamic variable added.
+%%
+If we assume a data set with 100 features with the formula $Z \sim X + Y$ we calculate 1,000,000 regression models.
+%%
+When each regression takes about 50~ms calculation, we have to wait roughly 14 hours for the calculation to complete.
+%%
+Therefore, the computational complexity needs to be reduced.
+%%
+%\com{Regression formulas are described using standard regression formula notation by introducing variables.}
+%%
+%\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction.}
 
 \subsection{Target-Variable-dependent Dimension Reduction}
-\com{Explain the CFS algorithm and how it is applied to reduce the number of dimensions and therefore the calculation process.}
+\com{Uli, Can you please proofread this?}
+%%
+The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
+%%
+Identifying these features und excluding them from the calculation can reduce the number of dimension significantly.
+%%
+The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for achieving this task \cite{CFS}.
+%%
+It uses information entropy to select the features which have the most explanatory power towards the target feature.
+%%
+At the same time it tries to reduce the correlation between the independent features.
+%%
+When for example the \emph{body weight} has a strong explanatory power towards the target it is likely that \emph{BMI} or \emph{waist circumference} behave similar towards it.
+%%
+They, however, correlate strongly with each other.
+%%
+The CFS algorithm would then select the feature which has the largest explanatory power and discards the other dimensions.
 
-\subsection{Cube Description using Regression Formula Notation}
-\com{Regression formulas are described using standard regression formula notation by introducing variables.}
+We apply the CFS algorithm for each target feature in a regression formula with dynamic variables.
+%%
+The formula $Cancer \sim X + Y$ would yield one initial CFS information space reduction.
+%%
+For $Z \sim X + Y$ the CFS algorithm is applied to the data every time $Z$ is replaced with another feature.
+%%
+
+The number of features calculated by the CFS algorithm is dependent on the information entropy in the data.
+%%
+In our epidemiological data we observed a usual number of 10 to 20 features.
+%%
+For features, such as age or gender, which affect most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}) the number is significantly larger (about half the features in the whole data set).
+%%
+
+Now we have a method to derive the interesting regression models in a reasonable time span.
+%%
+The next section shows ways of abstracting the results to make them visually feasible.
 

### 2015-03-19 draft.tex
wrote Encoding subsec on Regression Cube section

@@
 \subsection{Abstracting Regression Results using R$^2$}
-\com{R$^2$ are used to encode regression results to color spaces.}
-\com{R$^2$ is comparable between regression results of the same type, but not in between types. Therefore it is encoded using different colours.}
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3.0in]{figures/cube_sketch}
+ \caption{
+ (a) Overview visualization using a heatmap for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
+ %%
+ The R$^2$ values extracted from the regression formulas depict the goodness of fit and are mapped to color saturation.
+ %%
+ A saturated color shows a strong correlation.
+ %%
+ (b) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ heatmap visualizations.
+ %%
+ These represent the slices in our cube visualization.
+ }
+  \label{fig:Cube}
+\end{figure}
+%%
+The goal of a overview visualization is to provide a comprehensive view on the data, which is easy to understand.
+%% TODO: Cite another paper here!
+As described in \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
+%%
+Regression models are more complex, having many associated describing metrics.
+%%
+For the \emph{Regression Cube} analysis we are interested in the goodness of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
+%%
+As described in Section~\ref{sec:RegressionAnalysis}, the R$^2$ value is the metric allowing for this kind of assessment.
+%%
+Since it is scaled between [0,1] it also allows for comparison between regression models.
+%%
+We can apply the same information mapping by translating the R$^2$ values to color saturation (Fig.~\ref{fig:Cube} (a)).
+%%
+This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g. $Age \sim X + Y$).
+%%
+Introducing $Z$ creates a 3D \emph{Regression Cube}.
+%%
+The visualization of R$^2$ values derived from different regression cubes (e.g. $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%%
+Therefore, the R$^2$ results of different regression methods are encoded using different colors (e.g. blue for linear regression and red for logistic regression).
+%%
+This way, the cube can easily be extended using other regression types.
+%%
+For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings is required and the $z$ dimension can be compared directly.
+
+The goal is to create a overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
+%%
+These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
+%%
+Different analysis approaches require different starting points using the \emph{Regression Cube}.
 
@@
  \caption{
- \com{Screenshot of the cube visualization embedded in the framework}
+ \com{Screenshots of Cubes from the Evaluation}
  }

### 2015-03-19 draft.tex
added draft for workflow

@@
 Other features are exclusive, such as women-specific questions, e.g. number of born children or period status.
+
+\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
 
@@
  \centering
- \includegraphics[width=2.8in]{figures/placeholder}
+ \includegraphics[width=3.0in]{figures/workflow_sketch}
  \caption{
+ Workflow of the analysis using \emph{Regression Cubes}.
+ %%
+ The analysis starts with a set of features given by the data set.
+ %%
+ Then, the user may specify a formula regarding to specify a hypothesis, or use a predefined formula to start a explorative analysis of the data set.
+ %%
+ The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
+ %%
+ Insights into the data yield either an adjustment of the current formula.
+ %%
+ By applying a difference view, regression cubes can be compared to each other.
+ %%
+ The analysis yields insights and hypothesis about feature relations.
+ %%
+ }
+  \label{fig:Workflow}
+\end{figure}
+
+\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
  \com{Workflow for different analysis approaches.}
  \com{0. Select the Variables.}
- \com{1. Select the Formulas:}
+ \com{1. Select the Formulas}
  \com{a) Analysis without hypothesis about the data set (z ~ x+y)}
@@
  \com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
+ \com{d) Use Difference View}
  \com{d) Clarify using confounder (Var ~ x)}
@@
  \com{4. Get Details on Demand.}
- }
-  \label{fig:Workflow}
-\end{figure}
-\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
+%%
+Analyze first, Show Overview, Details on Demand
 %%

### 2015-03-19 draft.tex
revised version from Bernhard and Sylvia

@@
 %% TODO: Blind Review: Replace with online ID
-\onlineid{0}
+\onlineid{175}
 
@@
 %%
-These data contain information (\emph{features}) about a subjects' lifestyle, medical conditions and also medical images of the whole body.
+These data contain information (\emph{features}) about a subjects' lifestyle, medical conditions and also medical image data.
 %%
-These features are analyzed using statistical regression analyses towards features indicating a disease (the \emph{target feature}).
+These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
 %%
-Overview visualizations over a whole data sets towards a target feature are highly required, but not available to epidemiologists.
+Although there is a strong demand for overview visualizations of a whole data set towards a target feature, no suitable tool is available for epidemiological researchers.
+
+\com{New Solution}.
 %%
-%These analyses usually include two to five features (\emph{independent features}) towards the investigated condition.
-\com{New Solution}.
-%(New Solution).
-
-We propose a new analyses approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
+We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
 %%
-This approach combines all \emph{independent features} towards a \emph{target feature} and provides a visualization, which allows for insights into the data by highlighting relationships. %showing the
+This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insight into the data by highlighting relationships.
 %%
-A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview over the whole data set, the \emph{Regression Cube}.
-%%
-%We call it the \emph{Regression Cube}.
+A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
 %%
@@
 %%
-The regression model can be adjusted by the user to describe disease-specific hypotheses.
+Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
 %%
-Furthermore, the influences of features can be assessed using a difference view, comparing different calculation results.
-\com{Validation}.
+Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
+%%
+\com{Validation \& Results}.
 %%
@@
 %%
-We conducted a qualitative analysis with three domain experts on a breast fat data set to derive insights into the correlation between breast lesions and non-image variables.
+A qualitative analysis with three domain experts was conducted on a breast fat data set.
 %%
-\com{Results}.
-%%
-With our technique we were able to reproduce results from a data mining analysis of a hepatic steatosis data set.
-%%
-By applying it to a breast fat data set we were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
+We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
 %%
 \com{Implications}.
-Epidemiological data sets can for the first time be visually overviewed using a regression-based analysis of all features towards a disease. 
-%With our technique it is possible for the first time to provide a overview visualization of variables regarding a target variable.
-
+%%
+With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time a interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract

### 2015-03-19 draft.tex
Added workflow subsec to Cube section

@@
 \abstract{%
-\com{Problem}.
+%\com{Problem}.
 Epidemiological studies comprise heterogeneous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
@@
 
-\com{New Solution}.
+%\com{New Solution}.
 %%
@@
 %%
-\com{Validation \& Results}.
+%\com{Validation \& Results}.
 %%
@@
 %%
-\com{Implications}.
+%\com{Implications}.
 %%
@@
  These represent the slices in our cube visualization.
+ %%
+ The R$^2$ values of each slice voxel is mapped on opacity in the 3D-view, reducing the occlusion of other values.
  }
@@
 %%
-Since it is scaled between [0,1] it also allows for comparison between regression models.
+\paragraph{2D (Slice) View.}
 %%
-We can apply the same information mapping by translating the R$^2$ values to color saturation (Fig.~\ref{fig:Cube} (a)).
+Since R$^2$ is scaled between [0,1] it also allows for comparison between regression models.
+%%
+We can apply the same mosaic plot mapping by translating the R$^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
 %%
@@
 %%
-Introducing $Z$ creates a 3D \emph{Regression Cube}.
+
+\paragraph{3D (Cube) View.}
+%%
+Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
+%%
+R$^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
 %%
@@
 For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings is required and the $z$ dimension can be compared directly.
-
+%%
+\\\\
 The goal is to create a overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
@@
 \subsection{Analysis Workflow}
+\com{I'm not satisfied with the structure in this subsection, please proofread it with focus in the train of thought!}
 \begin{figure}[htb]
@@
 \end{figure}
+%%
+%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
+%\com{Workflow for different analysis approaches.}
+%\com{0. Select the Variables.}
+%\com{1. Select the Formulas}
+%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
+%\com{b) Analysis without hypothesis with confounder (z ~ age)}
+%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
+%\com{d) Use Difference View}
+%\com{d) Clarify using confounder (Var ~ x)}
+%\com{3. Find interesting pane.}
+%\com{4. Get Details on Demand.}
+%%
+The analysis workflow follows the Visual Analytics Mantra of Keim et al \cite{Keim2008}:
 
-\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
- \com{Workflow for different analysis approaches.}
- \com{0. Select the Variables.}
- \com{1. Select the Formulas}
- \com{a) Analysis without hypothesis about the data set (z ~ x+y)}
- \com{b) Analysis without hypothesis with confounder (z ~ age)}
- \com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
- \com{d) Use Difference View}
- \com{d) Clarify using confounder (Var ~ x)}
- \com{3. Find interesting pane.}
- \com{4. Get Details on Demand.}
+\textbf{Analyze First.} Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
+
+\textbf{Show the Important.} The 3D-cube visualization acts as overview over the whole data set.
 %%
-Analyze first, Show Overview, Details on Demand
+Regression models with large R$^2$ values can be spotted fast here, steering the attention to the respective slice.
+
+\textbf{Zoom, Filter and Analyze Further.} The slices of interest can then be analyzed using the 2D mosaic plot of the slice.
+
+\textbf{Details on Demand.} Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
+in a hover-modal on a currently selected data point).
+\\\\
+As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
 %%
+Observations in the 2D-mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
+%%
+This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
+%%
+\paragraph{Formula Types.}
+%%
+Input formulas reflect \emph{hypotheses} about the data.
+%%
+Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
+%%
+Here we show just a few examples on how formulas can serve different purposes.
+
+$Z \sim X + Y$ is the default if the user did not specify a formula.
+%%
+This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
+%%
+Each slice represents a different target feature.
+%%
+It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.
+
+$Cancer \sim X + Y + Z + feature_1:feature_2$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
+%%
+The interaction between $feature_1$ and $feature_2$ is also included in all regression formulas.
+%%
+
+$Cancer \sim X + Y + Z$ subtracted with results from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
+%%
+\com{TODO: divide using analysis approaches, e.g. Hypothesis-based \& Hypothesis-free with follow up on the formulas used?}
+%%
+\paragraph{Cube Comparison.}
+
 \section{System Design and Implementation} \label{Interaction- and Visualization Techniques}

### 2015-03-20 draft.tex
added first part of System Design and Implementation

@@
 \section{Regression Cube Analysis of Cohort Study Data}
+\com{Add Data Preprocessing Section.}
 %%
@@
 %%
+\paragraph{Cube Comparison.}
+%%
+Regression Cubes can be compared by creating difference views.
+%%
+One cube (formula) acts as reference.
+%%
+The absolute difference in R$^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
+%%
+\com{TODO: divide using analysis approaches, e.g. Hypothesis-based \& Hypothesis-free with follow up on the formulas used?}
+\paragraph{Hypothesis-Free Analysis}
+%%
+\paragraph{Hypothesis-Based Analysis}
+%%
+%%
 \paragraph{Formula Types.}
@@
 %%
-\com{TODO: divide using analysis approaches, e.g. Hypothesis-based \& Hypothesis-free with follow up on the formulas used?}
+
+\section{System Design and Implementation} \label{sec:SystemDesign}
 %%
-\paragraph{Cube Comparison.}
-
-\section{System Design and Implementation} \label{Interaction- and Visualization Techniques}
+We designed our system with openness and ease of use in mind.
+%%
+Using open formats as input interfaces does not restrict the application towards non-epidemiological data sets.
+%%
+The focus lies on creating a overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
+%%
+Therefore the system has to be intuitive and comprehensive in order to be adapted by domain experts.
 %%
@@
 \end{figure}
+%%
+Design choices and spaces are restricted by the underlying technologies and devices.
+%%
+Resting collaboration and exchange about data and methods on web-based technologies shows promising results, as there is no set-up time involved and domain experts can use the methods from any computer connected with the web.
+%%
+Web technology is also inherently designed with client-server architecture in mind, making it easy to outsource computation heavy tasks on server-clusters and transferring results to the client device.
+%%
+The design space of web-technologies is different to standard WIMP-Interfaces, right click menus, modal windows and menu bars are not established UI components in this context.
+%%
+Therefore we have to adapt our design respecting the affordances of web pages.
+%%
+Our prototype consists of different components:
+\begin{itemize}
+	%%
+	\item The \emph{file upload} section starting the analysis with providing a comma separated values (CSV) data set.
+	%%
+	\item The \emph{formula editor} allows to adjust the formula towards a current hypothesis or conducting a \emph{hypothesis-free} analysis
+	%%
+	It also allows to select a reference formula for creating a difference-cube.
+	%%
+	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube.
+	%%
+\end{itemize}
+
+\paragraph{File-Upload and Classification.}
+%%
+Popular analytics tools, such as WEKA \cite{WEKA} owe their popularity to their support of open file types.
+%%
+To allow other users even outside the epidemiological application domain access to our tool, we choose to do this using standard ASCII-based CSV files.
+%%
+The first line in a CSV file represents all features (columns) of the data set.
+%%
+Each line after that represents one subject (row) and their feature manifestations.
+%%
+Encoding variable types in CSV files is not standardized.
+%%
+We, however, need to ensure the correct variable type classification and have to enforce some basic standards.
+%%
+All categorical values have to be enclosed by quotation marks.
+%%
+Continuous variables are denoted as digits without enclosing quotation marks.
+%%
+This seems obvious, but in fact many cohort study data sets encode categorical features using ID-values, which are denoted in a data dictionary.
+%%
+Variables with only two manifestations are classified as dichotomous, leading to three possible data types: numerical, categorical and categorical/dichotomous.
+%%
+Missing values are denoted using no character at all, a whitespace, or an empty quotation mark encapsulated string.
+
+Security issues are raised by uploading data into a online service, such as our prototype.
+%%
+The use of epidemiological data is preceded by a detailed description on the analysis purpose and has to be approved by ethics committees.
+%%
+Preventive steps have to be taken to restrict access to unauthorized subjects.
+%%
+We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
+%%
+Instead we apply SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
+%% TODO: Implement this
+Data sets are deleted from the server after the closing a session.
+
+\paragraph{Formula Editor.}
+%%
+After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
+%%
+Entering a formula is provided via text input.
+%%
+On entering the text are, a context modal displays all data set variables as well as the available operators and their function.
+%%
+This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
+%%
+Auto-completing input variables also simplifies the approach and also works as spell-check of (sometimes hard to memorizable) variable names.
+%%
+
+The formula is checked for validity directly on input, indicating invalid formulas with a red halo around the text input, which turns to green for valid input.
+%%
+This prevents processing errors on the statistical processor backend.
+%%
+Confirming a formula triggers the cube calculation, which is preceded by determining all required formulas.
+%%
+These are then divided by the number of available statistical backend processors available, driving a \emph{cloud computing} based approach.
+%%
+In theory, the calculation duration is reduced by a factor of $0.5$ by every statistical processor.
+%%
+In praxis, data transmission and differences in machine specifications always influence the speed.
+
+%\com{Dynamic Type Checking, Auto-Completion Analysis Trigger, Comparison between Cubes, triggering of cube calculation, parallelization of formula calculation}
+
+\paragraph{Regression Cube Visualization.}
+%%
+\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
+
+\com{Simplicity of interface to allow for steep learning curve.}
+\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
+\com{Cube acts as visual mini map to give a global impression over the data.}
+\com{current pane shown using a heat map visualization.}
+\com{details on demand for selected regression formula.}
+\com{comparison view using reference cubes.}
 
@@
 \end{figure}
-%%
-\com{Simplicity of interface to allow for steep learning curve.}
-\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
-\com{Cube acts as visual mini map to give a global impression over the data.}
-\com{current pane shown using a heat map visualization.}
-\com{details on demand for selected regression formula.}
-\com{comparison view using reference cubes.}
 
@@
 \com{Use VAST'14 as Guideline}
+\com{Cloud computing}
 \com{Speed is merely a matter of available server machines due to parallelization process.}
@@
 %%
-\begin{small}
-	\acknowledgments{Omitted due to blind review}
-% \acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no.
+% \begin{small}
+	%\acknowledgments{Omitted due to blind review}
+\acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no.
 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania.
@@
 We thank Marko Rak and Klaus Toennies for providing the image detection data.}
-\end{small}
+% \end{small}
 \clearpage

### 2015-03-20 draft.tex
nearly finished System Design and Implementation

@@
 %%
-The analysis workflow follows the Visual Analytics Mantra of Keim et al \cite{Keim2008}:
+The analysis workflow follows the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:
 
@@
  \centering
- \includegraphics[width=2.8in]{figures/placeholder}
+ \includegraphics[width=3.5in]{figures/system_sketch}
  \caption{
- \com{Screenshot of the cube visualization embedded in the framework}
- \com{Also include active tooltip}
+ %%
+ \com{FIGURE NOT FINAL!}
+ %%
+ Breast fat data set loaded into the \emph{Regression Cube} prototype.
+ %%
+ (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
+ %%
+ (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
+ %%
+ The values of the current selected slice are mirrored and represented as orange data points on the slicing plane.
+ %%
+ (c) 2D mosaic plot visualization of the selected slice.
+ %%
+ \com{TODO: Adapt formula to final image.}
+ %%
+ Context information is shown for the regression model of $Parenchym \sim Bronchitis + AlcoholLast30Days$, depicting details about the confidence intervals, coefficients and associated p-values.
+ %%
+ %\com{Screenshot of the cube visualization embedded in the framework}
+ %\com{Also include active tooltip}
  }
-  \label{fig:Cube}
+  \label{fig:System}
 \end{figure}
@@
 %%
-The design space of web-technologies is different to standard WIMP-Interfaces, right click menus, modal windows and menu bars are not established UI components in this context.
+The design space of web-technologies is different to standard WIMP-applications--right click menus, modal windows and menu bars are not established UI components in this context.
 %%
@@
 %%
-Our prototype consists of different components:
+\paragraph{System Paradigm and Components.}
+%%
+The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of ui elements as much as possible.
+%%
+This allows for a fast learning of the individual system parts.
+%%
+More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
+%%
+Therefore, our prototype consists of three components:
 \begin{itemize}
@@
 \end{itemize}
-
+%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
 \paragraph{File-Upload and Classification.}
@@
 Each line after that represents one subject (row) and their feature manifestations.
+
+\textbf{Classification.}
 %%
@@
 
+\textbf{Data Security.}
+%%
 Security issues are raised by uploading data into a online service, such as our prototype.
@@
 %%
-Instead we apply SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
+Instead we apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
 %% TODO: Implement this
@@
 
-The formula is checked for validity directly on input, indicating invalid formulas with a red halo around the text input, which turns to green for valid input.
+\textbf{Formula Validation and Calculation.} The formula is checked for validity directly on input, indicating invalid formulas with a red halo around the text input, which turns to green for valid input.
 %%
@@
 
+%%
+\textbf{Difference-Cube.}
+Adding a formula also adds it to the reference selection for a difference cube.
+%%
+If the user selects a reference formula, the absolute difference of R$^2$ values between the cube described by the reference and the currently active formula are calculated and shown.
+%%
+
 %\com{Dynamic Type Checking, Auto-Completion Analysis Trigger, Comparison between Cubes, triggering of cube calculation, parallelization of formula calculation}
@@
 %%
-\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
+The visualization and interaction with the \emph{Regression Cube} is the heart of the prototype.
+%%
+Results from the statistical processors are uploaded into the visualization by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.
 
-\com{Simplicity of interface to allow for steep learning curve.}
-\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
-\com{Cube acts as visual mini map to give a global impression over the data.}
-\com{current pane shown using a heat map visualization.}
-\com{details on demand for selected regression formula.}
-\com{comparison view using reference cubes.}
+\textbf{Usage of a Regression Prism for information reduction.}
+%%
+Figure~\ref{fig:Cube} shows that all values are mirrored along the mosaic plot matrix diagonal.
+%%
+This is due to the symmetry of basic regression operators.
+%%
+$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
+%%
+Therefore we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
+%%
+This opens up space for displaying additional information.
+%%
+
+\textbf{3D-Prism as Data Mini-Map.}
+%%
+The 3D-cube representation acts overview over the whole data set, but its purpose is not to derive detailed information about data points.
+%%
+It serves a function similar to a mini-map, guiding the attention towards points of interest in the data as well as giving context information about adjacent data values when using the 2D mosaic plot.
+%%
+It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
+%%
+The displayed prism shows values above the matrix diagonal.
+%%
+\\\\
+%% TODO: Add reference for disadvantages of 3D information visualizations!
+\textbf{Tackling the disadvantages of 3D information visualization.}
+3D-Information visualizations are often criticized for introducing occlusion and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
+%%
+We aim to minimize these advantages as much as possible.
+%%
+Since the R$^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
+%%
+It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low R$^2$ values.
+%%
+Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
+%%
+Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
+%%
+
+%% INTERACTION
+Rotating with the cube is restricted to y axis, preserving the mental map of the position of the individual features.
+%%
+The cube is always oriented according to the 2D-representation, allowing for a easy mental combination of the two representations.
+%%
+Allowing more degrees of freedom was confusing to our users and also does not add value to the visualization.
+%%
+We provide also a zoom functionality using the mouse wheel input.
+%%
+\\\\
+\textbf{Cube Slice Selection.}
+%%
+In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
+%%
+We propose two ways to achieve this.
+\begin{itemize}
+	\item \emph{Applying the slicing metaphor from 3D volume data.}
+	%%
+	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
+	%%
+	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
+	%%
+	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
+	%%
+	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
+	%%
+\end{itemize}
+%%
+The current selected slice is denoted with a semi-transparent gray plane.
+%%
+The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
+%%
+the values are projected on this plane to give a overlapping-free view on the data points, which makes it easier to identify the current slice.
+%%
+%It also allows for a fast slicing through the cube and assessing the information.
+%%
+%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
+\\\\
+\textbf{2D Mosaic Plot Slice Visualization.}
+%%
+The 2D mosaic visualization shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
+%%
+To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
+%%
+The free space above the matrix diagonal is used to display the 3D cube.
+%%
+
+The purpose of this view is the detailed assessment of the underlying regression models.
+%%
+By hovering over a data entry in the plot, a tooltip model displays detailed information about a models coefficients, associated p values and confidence intervals.
+%%
+
+%\com{Simplicity of interface to allow for steep learning curve.}
+%\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
+%\com{Cube acts as visual mini map to give a global impression over the data.}
+%\com{current pane shown using a heat map visualization.}
+%\com{details on demand for selected regression formula.}
+%\com{comparison view using reference cubes.}
 
@@
 \end{figure}
-
-\paragraph{System Layout.}
 
@@
 \com{Cloud computing}
+\com{Volume Rendering vs Particles}
 \com{Speed is merely a matter of available server machines due to parallelization process.}

### 2015-03-20 draft.tex
added parts of Implementation section

@@
 %%
+
+\paragraph{Front-End.}
+The front end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
+%%
+\texttt{Angular.js} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+%%
+\texttt{Angular.js} is forcing developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
+%%
+The page layout is handled using \texttt{Twitter Bootstrap}, which also provides a rich set of UI elements with proper stylings.
+%%
+The 2D mosaic plot is implemented using \texttt{Data driven Documents}(\texttt{D3.js}), which is very popular in information visualization using vector graphics.
+%%
+It provides fast and easy methods for binding data to graphical elements.
+%%
+The 3D plot is created using the WebGL-based \texttt{Threejs} library.
+%%
+We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
+%% TODO
+Using a volume rendering algorithm for the cube \com{TODO}.
+%%
+Creating a cube primitive for each data point resulted in non-interactive framerates for data sets larger than 30 features (creating 30$^3$ cube primitives).
+%%
+We use a shader-based solution by rendering the cube as a sprite based particle system, allowing to customize color and opacity of every data point.
+%%
+It also is the fastest solution that we tested.
+
+\paragraph{Back-End.}
+Two server structures serve as back-end.
+%%
+First, the web-server is written in \texttt{Javascript} using \texttt{NodeJS}, running on Googles V8 Javascript runtime environment.
+%%
+It is hosted on \texttt{Heroku}, a cloud application platform.
+%%
+
+The statistical processors are the second structure.
+%%
+They rely on the statistical programming language \texttt{R}.
+%%
+It is widely adopted in the statistical analysis community, yielding a rich support of fast state-of-the-art statistics algorithms as well newly published methods.
+%%
+\texttt{OpenCPU} provides a API for accessing \texttt{R} via HTTP calls.
+%%
+
+\paragraph{Source and Access.}
+
 \com{Use VAST'14 as Guideline}

### 2015-03-21 draft.tex
added Implementation subsec

--- 2015-03-20 draft.tex
@@
 %%
-\com{Use VAST'14 as Guideline}
-\com{Cloud computing}
-\com{Volume Rendering vs Particles}
-\com{Speed is merely a matter of available server machines due to parallelization process.}
-\com{Security enabled by hashing files.}
-\com{Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend (Fig.~\ref{fig:Implementation})}
+Relying on web technologies paid off in previous collaboration with our clinical exerts, since it allows for:
+\begin{itemize}
+	\item easy exchange of software, without the need of installing anything other than a modern web browser,
+	\item fast iterative adaption of method changes, without the need of installing updates,
+	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
+\end{itemize}
+%%
+The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
+%%
+\paragraph{Front-End.}
+%%
+The front end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
+%%
+\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{www.angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+%%
+\texttt{Angular.js} is forcing developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
+%%
+The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of UI elements with proper stylings.
+%%
+The 2D mosaic plot is implemented using \texttt{Data driven Documents}(\texttt{D3.js}), which is very popular in information visualization using vector graphics \cite{D3}.
+%%
+It provides fast and easy methods for binding data to graphical elements.
+%%
+The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
+%%
+We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
+%%
+Proper open source volume rendering methods are not available for the web, the effort of writing one from scratch was not worth the counter value.
+%%
+Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
+%%
+We use a shader-based solution by rendering the cube as a sprite based particle system, allowing to customize color and opacity of every data point.
+%%
+It also is the fastest solution that we tested.
+
+\paragraph{Back-End.}
+Two server structures serve as back-end.
+%%
+The first one is the web-server, which is written in \texttt{Javascript} using \texttt{NodeJS}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
+%%
+It is hosted on \texttt{Heroku}\footnote{Owned by Salesforce.com, \href{https://www.heroku.com/}{\texttt{heroku.com}}}, a cloud application platform.
+%%
+
+The statistical processors yield the second structure.
+%%
+They rely on the statistical programming language \texttt{R}\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}.
+%%
+It is widely adopted in the statistical analysis community, yielding a rich support of fast state-of-the-art statistics algorithms as well newly published methods.
+%%
+\texttt{OpenCPU} is a \texttt{R} package and provides a API for accessing it via HTTP calls.
+%%
+This way, any computer, which runs \texttt{R} can be turned into a statistical processor for our project.
+%%
+The backend functions necessary for all cube calculations are provided via an \texttt{R} package.
+%%
+It uses multi-core optimization to use all machine CPUs to speed up the calculation process.
+%%
+The server workload balances are managed by the front-end code.
+
+\paragraph{Access and Source.}
+%%
+A running instance of the \emph{Regression Cube} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
+%%
+The source for the prototype is freely available at \texttt{Github}.
+%%
+Instructions and code on how to create a setup running the \emph{Regression Cube} statistical backend through a \texttt{Ubuntu} server using \texttt{OpenCPU} can be found here: \href{https://github.com/paulklemm/regression-cube-r-package}{\texttt{github.com/paulklemm/regression-cube-r-package}}.
+%%
+The front-end source as well as the \texttt{Nodejs} backend can be found here: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{github.com/paulklemm/regression-cube-prototype}}.
+%%
+Using \href{https://www.heroku.com/}{\texttt{www.heroku.com}}, the front-end setup is as easy as cloning the repository into a Heroku homepage. 
+%%
+% \com{Use VAST'14 as Guideline}
+% \com{Cloud computing}
+% \com{Volume Rendering vs Particles}
+% \com{Speed is merely a matter of available server machines due to parallelization process.}
+% \com{Security enabled by hashing files.}
+% \com{Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend (Fig.~\ref{fig:Implementation})}
 

### 2015-03-21 draft.tex
restructured Design and Implementation

@@
 
-\section{System Design and Implementation} \label{sec:SystemDesign}
-%%
-We designed our system with openness and ease of use in mind.
-%%
-Using open formats as input interfaces does not restrict the application towards non-epidemiological data sets.
-%%
-The focus lies on creating a overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
-%%
-Therefore the system has to be intuitive and comprehensive in order to be adapted by domain experts.
-%%
-
-\subsection{Design and Visualization Techniques} \label{Structure and Workflow}
+\section{System Design} \label{sec:SystemDesign}
 \begin{figure}[htb]
@@
 %%
-Design choices and spaces are restricted by the underlying technologies and devices.
+We designed our system with openness and ease of use in mind.
+%%
+Using open formats as input interfaces does not restrict the application towards non-epidemiological data sets.
+%%
+The focus lies on creating a overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
+%%
+Therefore the system has to be intuitive and comprehensive in order to be adapted by domain experts.
+%%
+\\\\
+% \subsection{Design and Visualization Techniques}
+%%
+\noindent Design choices and spaces are restricted by the underlying technologies and devices.
 %%
@@
 %%
-\paragraph{System Paradigm and Components.}
+\subsection{System Paradigm and Components}
 %%
@@
 
-\paragraph{Regression Cube Visualization.}
+\subsection{Regression Cube Visualization.}
 %%
@@
 
-\textbf{Usage of a Regression Prism for information reduction.}
+\paragraph{Usage of a Regression Prism for information reduction.}
 %%
@@
 
-\textbf{3D-Prism as Data Mini-Map.}
+\paragraph{3D-Prism as Data Mini-Map.}
 %%
@@
 %% TODO: Add reference for disadvantages of 3D information visualizations!
-\textbf{Tackling the disadvantages of 3D information visualization.}
+\paragraph{Tackling the disadvantages of 3D information visualization.}
 3D-Information visualizations are often criticized for introducing occlusion and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
@@
 %%
-\\\\
-\textbf{Cube Slice Selection.}
+%\\\\
+\paragraph{Cube Slice Selection.}
 %%
@@
 %\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
-\\\\
-\textbf{2D Mosaic Plot Slice Visualization.}
+%\\\\
+\paragraph{2D Mosaic Plot Slice Visualization.}
 %%
@@
 
-\subsection{Implementation} \label{implementation}
+\section{Implementation} \label{implementation}
 \begin{figure}[htb]
@@
 %%
-The source for the prototype is freely available at \texttt{Github}.
-%%
-Instructions and code on how to create a setup running the \emph{Regression Cube} statistical backend through a \texttt{Ubuntu} server using \texttt{OpenCPU} can be found here: \href{https://github.com/paulklemm/regression-cube-r-package}{\texttt{github.com/paulklemm/regression-cube-r-package}}.
-%%
-The front-end source as well as the \texttt{Nodejs} backend can be found here: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{github.com/paulklemm/regression-cube-prototype}}.
-%%
-Using \href{https://www.heroku.com/}{\texttt{www.heroku.com}}, the front-end setup is as easy as cloning the repository into a Heroku homepage. 
+The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\texttt{github.com/paulklemm/regression-cube-r-package}}}\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{github.com/paulklemm/regression-cube-prototype}}}.
+%%
+Instructions and code on how to create a setup running the \emph{Regression Cube} statistical backend through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
+%%
+The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku homepage. 
+%%
+%The front-end source as well as the \texttt{Nodejs} backend can be found here: .
 %%

### 2015-03-21 draft.tex
revised abstract with Kais notes

@@
 %%
-These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
+%These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
+%%
+Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
 %%
@@
 %%
-This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insight into the data by highlighting relationships.
+This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
 %%
@@
 %%
-With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time a interactive regression-based analysis of large feature sets with respect to a disease.
+%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time a interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract

### 2015-03-21 draft.tex
added Application Background for Breast Fat Data

@@
 \section{Application} \label{application}
-%
+%%
+In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
+%%
+The hepatic steatosis data set was analyzed using data mining algorithms, yielding risk groups, which we now analyze further.
+%%
+Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
+%%
+The breast fat data set is the foundation for a explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
+%%
+
+Both data sets are unusual for epidemiological analysis regarding their feature extend.
+%%
+Usually, only a few features depicting a hypotheses are compiled into a data set to assess them using statistical tools.
+%%
+Our method focuses on data exploration and knowledge extraction and requires a wide scope of sociodemographic, medical and lifestyle features.
+%%
+\subsection{Participants, Setup and Procedure}
+%%
+To assess the ability a system to discover knowledge discover knowledge is hard to measure.
+%%
+Lam et al.
+\cite{Lam2012} propose for this purpose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems ability to generate hypotheses and explore the data in order to extract information.
+%%
+\emph{VDAR} can be carried out using case studies using thinking-aloud techniques to comprehend the reasoning and thought process of the user.
+%%
+We conducted a study using two participants, who also co-authored this publication.
+%%
+UN is a data scientist who analyzed the hepatic steatosis data set by extracting decision rules.
+%%
+KH is a radiologist (10 years of experience) with specialization in epidemiological research and is also responsible for the SHIP MRI acquisition.
+%%
+\paragraph{Setup and Procedure.}
+%%
+One major advantage of focusing on web based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
+%%
+We conducted a web based research by using a online-meeting software, which features voice chat as well as screen sharing.
+%%
+Starting a analysis using these techniques take about 5-10 minutes of setup time.
+
+The sessions started with an initial overview of the system, showcasing its features and functionality.
+%%
+After that the experts use the system on their own computers.
+%%
+The screen-sharing function was still used to observe the actions of the expert.
+%%
+All sessions were video recorded to be processed later on.
+%%
 \subsection{The Breast Fat Data Set}
-\com{Describe how the data was acquired and preprocessed}
-
-\subsubsection{Data Preprocessing} \label{application:Data Preprocessing}
-The data processing follows the description in Section~\ref{Data Preprocessing}.
-%
-
+The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
+%%
+The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
+%%
+Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.
+
+The data comprises 1.186 subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
+%%
+It contains 230 dimensions, holding information about
+\begin{itemize}
+	\item \emph{somatometric features}, e.g. body size, weight or BMI, 
+	\item \emph{lifestyle features}, e.g. alcohol/tobacco consume, 
+	\item \emph{personal history}, e.g. occupation, marital status,
+	\item \emph{medical history}, e.g. current or prior diseases or laboratory values, such as cholesterol levels,
+	\item \emph{women specific features}, e.g. number of born children, contraception type or hormone replacement therapy and
+	\item \emph{mammography features} derived from the image data.
+\end{itemize}
+%%
+The latter were derived using a level set based segmentation \cite{Ivanovska2014} of MRI image data for each subject.
+%%
+\com{Check this with Mrs. Hegenscheid, probably most of them were segmented by hand!}
+%%
+%\com{Describe Parenchym tissue and its relation to breast cancer.}
+%\com{Describe how the data was acquired and preprocessed}
+
+The data of each cohort was presented as individual \texttt{SPSS} files.
+%%
+All data related to the mammography was stored in a additional file.
+%%
+We converted the \texttt{SPSS} data sets to CSV and used \texttt{R} to merge the data sets together using their ID.
+%%
+All features were renamed to be expressive, e.g. \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
+%%
+This avoids the need of defining a separate data dictionary file for the \emph{Regression Cube}, translating the feature names.
+%%
+All male subjects were removed as their data does not contribute to the analysis.
+%%
 \subsection{The Hepatic Steatosis Data Set}
 \com{Uli: Describe the data set and also prior work on it.}
-
-\subsubsection{Data Preprocessing} \label{application:Data Preprocessing}
-The data processing follows the description in Section~\ref{Data Preprocessing}.
-%
-
-\subsection{Participants, Setup and Procedure}
-%%
-\com{Use of VDAR Technique}
-%
-\paragraph{Setup.} Due to the large geographical distance, the evaluation was done completely web-based.
-%%
-\paragraph{Procedure.}
+%%
+\subsubsection{Data Preprocessing}
 %%
 \subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
-
+%%
 \subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}

### 2015-03-21 draft.tex
worked on structure of Related Work

@@
 \section{Prior and Related Work}
+\com{Add section on most similar publications}
+\com{TODO: Needs to be rewritten, this is only here for the references!}
 \com{From VAST'14 Paper:}
+
+\com{http://www.ii.uib.no/vis/publications/publication/2014/Angelelli14Interactive}
+%\com{http://link.springer.com/chapter/10.1007/978-3-642-39146-0_1#page-1}
+\com{Turkey Metrics Paper}
+
+\paragraph{Data Mining in Epidemiology.}
+\com{Problem: Only paper of co-authors in this section, will look one-sided!}
+\com{Uli Niemann, Tommy Hielscher, Myra Spiliopoulou, Henry Völzke, and Jens-Peter Kühn. Can we classify the participants of a longitudinal epidemiological study from their previous evolution?.}
+
+\com{Niemann, U.; Völzke, H.; Kühn, J.-P.
+Spiliopoulou, M.
+(2014), Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis, Expert Systems with Applications 41 (11) , 5405-5415 .}
+
+\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
+
+\com{Uli, can you write a short summary on these?}
+
+\paragraph{Statistical Analysis}
+\com{Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding}
+
+\com{[Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics]}
+
+\com{Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization}
+
 \paragraph{Visual Analysis of Heterogenous Data.}
@@
 %%
-\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{www.angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
 %%
@@
 %%
-The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\texttt{github.com/paulklemm/regression-cube-r-package}}}\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{github.com/paulklemm/regression-cube-prototype}}}.
+The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}.
 %%
@@
 \section{Summary and Conclusion}
-\paragraph{Future Work.}
-\com{Implementing of regression analysis on the graphics card.}
+We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
+%%
+\emph{Regression Cube} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
+%%
+Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
+%%
+These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
+%%
+Correlation based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
+%%
+The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
+%%
+All code associated to this project is also open source.
+%%
+Two case studies revealed that the presented approach is able to reproduce knowledge extracted using decision tree based data mining methods on a hepatic steatosis data set as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
+%%
+
+As future work, we want to introduce more regression models to the data set, which model different correlation types.
+%%
+We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
+%%
+
+The method was very well received with our project partners, allowing them for the first time to retrieve a overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
+%%
+The observed relationship are now subject of detailed statistical analyses.
+%%
+Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
+%%
+We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
+%%
+%%\paragraph{Future Work.}
+%%\com{Implementing of regression analysis on the graphics card.}
 %%

### 2015-03-23 draft.tex
added first part of related work

@@
 %%
-As described by \cite{Thew2009} is divided into three different steps:
+As described by Thew et al. \cite{Thew2009} it is divided into these different steps:
 \begin{itemize}
@@
 \section{Prior and Related Work}
-\com{Add section on most similar publications}
-\com{TODO: Needs to be rewritten, this is only here for the references!}
-\com{From VAST'14 Paper:}
-
-\com{http://www.ii.uib.no/vis/publications/publication/2014/Angelelli14Interactive}
+%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
+Tukey already stated in 1977 that data if too often only analyzed using confirmatory data analysis \cite{Tukey}.
+%%
+He emphasized the need to use data to derive hypotheses, which can then be tested again.
+%%
+In this section we present prior and related work trying to achieve this goal.
+%%
 %\com{http://link.springer.com/chapter/10.1007/978-3-642-39146-0_1#page-1}
 \com{Turkey Metrics Paper}
-
-\paragraph{Data Mining in Epidemiology.}
+\\\\
+The work of Piringer et al. \cite{Piringer}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
+%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
+%%
+Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
+%%
+Their main goal is to pairwise assess the influence of independent features towards the target.
+%%
+This is achieved by using a plot matrix displaying models as contours towards the target feature.
+%%
+They also incorporate 3D visualizations for each pairwise combination, but mainly because it is popular with the engineers in their target domains.
+%%
+Linked views of model deviations allow to select outlier residuals.
+%%
+Comparison of models is also included by mapping model values on line-style (e.g. solid for $model_a$ and dashed for $model_b$).
+%%
+This limits the method to comparing a few models at once.
+%%
+The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
+%%
+The main difference is their focus on analyzing one complex model in detail, yielding in extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
+%%
+They also focus on metric features, while we process also categorical data.
+%%
+We share the matrix-arrangement of results depicting relationships between independent features in a model.
+
+Zhang et al. \cite{Zhang2014} is closest to ours regarding the application of visual analysis on cohort study data.
+%%
+Their work present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} and its manipulation using operations - \emph{views} and \emph{analytics}.
+%%
+They use the system to find longitudinal pathways for diseases using subject health records.
+%%
+Sub-groups are created automatically by applying \emph{analytics} and expert-guided using \emph{views}.
+%%
+Unfortunately, they do not go into the detail into the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
+%%
+Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
+%%
+We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}, meaning that the system has to adapt to different requirements in different analysis steps.
+%%
+We implement this by supporting both hypothesis-based and hypothesis-free analysis.
+
+Angelelli et al. \cite{Angelelli} is similar to ours as they visualize both image-derived an non image data using cube data structures with focus on comparison and knowledge extraction.
+%%
+Their proposed engine represents multiple heterogenous cohort study data sets as multiple normalized data cubes to eliminate data redundancy between them and perform runtime measure aggregation when measures of the different cubes have to be combined.
+%%
+This allows for saving memory space.
+%%
+Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
+%%
+They use \emph{pearssons R} to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
+%%
+We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
+%%
+We employ the idea of fast conducting large scale correlation analysis between many features to derive insight into the data.
+
+\paragraph{Visual Analysis in Public Health.}
 \com{Problem: Only paper of co-authors in this section, will look one-sided!}
@@
 
+In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
+%%
+They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
+%%
+We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
+
+Shneiderman et al.
+\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data, e.g.
+using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
+%%
+They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
+%%
+We focus on the latter.
+%%
+%Shneiderman and colleagues also present challenges, 
+The presented challenges in this application comprise custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
+%%
+Futhermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
+%%
+We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
+
+Schreck and Keim \cite{Schreck} presented a method for analyzing social media data with focus on characterization and cause of an epidemic.
+%%
+They employ data aggregation techniques and use geographical data to map social media events, focusing on word clouds to derive insight into possible causes.
+%%
+They derive hypotheses using their high variation average message density algorithms, steering the users focus on particular events.
+%%
+The focus remains on geolocated social media data rather than subjects with multiple dimensions.
+%%
+These and similar systems follow the same analysis work as we do to derive insight into data.
+
+Perer et al. divide cohort analysis in three major parts, present \emph{CAVA}, a system which allows
+
+\com{Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies}
+
 \paragraph{Statistical Analysis}
+
 \com{Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding}
+\com{The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems}
 

### 2015-03-24 draft.tex
added all content to related work; needs restructuring

@@
 
-\section{Epidemiological Background} \label{MedicalAndTechnicalBackground}
+\section{Epidemiological Background} \label{sec:Background}
 %%
@@
 %%
-%\com{http://link.springer.com/chapter/10.1007/978-3-642-39146-0_1#page-1}
-\com{Turkey Metrics Paper}
+We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
+%%
 \\\\
-The work of Piringer et al. \cite{Piringer}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
+The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
 %% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
-%%
 Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
@@
 
+%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
+We use the term Regression Cube to refer to a comparative 3D representation of multiple regression models.
+%%
+Chan et al. \cite{Chan} use the same term to describe an extension of the 2D-Scatterplot representation of a linear regression models (incorporating solely metric features) to a 3D-Cube.
+%%
+They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
+%%
+Similar to \cite{Piringer}, they focus on highlighting details of the included models, rather than comparing models consisting of different features.
+%%
+Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
+%%
+In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
+%%
+We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
+%%
+\com{Does this suffice as reason on why we use the same term "Regression Cube" for different things?}
+
 Zhang et al. \cite{Zhang2014} is closest to ours regarding the application of visual analysis on cohort study data.
@@
 %%
-They use \emph{pearssons R} to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
+They use \emph{Pearsson's r} to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
 %%
@@
 \paragraph{Visual Analysis in Public Health.}
-\com{Problem: Only paper of co-authors in this section, will look one-sided!}
 \com{Uli Niemann, Tommy Hielscher, Myra Spiliopoulou, Henry Völzke, and Jens-Peter Kühn. Can we classify the participants of a longitudinal epidemiological study from their previous evolution?.}
-
-\com{Niemann, U.; Völzke, H.; Kühn, J.-P.
-Spiliopoulou, M.
-(2014), Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis, Expert Systems with Applications 41 (11) , 5405-5415 .}
-
+%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
+Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis (fatty liver disease) using decision trees with a interactive data mining tools.
+%%
+They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
+%%
+We try to reproduce their results and investigate their findings further.
+%%
 \com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
@@
 
-\com{Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies}
+%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
+Steenwijk et al. \cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
+%%
+They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
+%%
+They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select data.
+%%
+Medical image data can be incorporated to investigate potentially interesting individuals.
+%%
+We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
 
@@
 
-\com{Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding}
-\com{The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems}
-
-\com{[Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics]}
-
-\com{Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization}
+%\com{The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems}
+Ahmadi et al.
+\cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described using by their most reliable linear regression model.
+%%
+In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
+%%
+They focus on presenting a algebraic for efficient regression-model calculation to find the best fit for a subspace.
+%%
+The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.
+
+%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
+Albuquerque et al. \cite{Albuquerque} present a interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
+%%
+These subsets are then analyzed in detail, following a drilling-down approach.
+%%
+They incorporate scatter plot matrices (SPLOMs) of quality metrics describing the features.
+%%
+We incorporate the idea of abstracting features for comparison.
+%%
+Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
+%%
+The approach was applied to a cognitive aging studies, where new hypotheses could be derived.
+
+% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
+Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
+%%
+%According to their systematization we analyze correlations.
+%%
+Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
 
 \paragraph{Visual Analysis of Heterogenous Data.}
-Zhang et al. \cite{Zhang2012} provide a web-based system for analyzing subject groups with linked views and batch-processing capabilities for categorizing new subject entries into the data set.
-%
-Their definition of a cohort differs from the understanding of the term in an epidemiological context by denoting every parameter-divided subject group as individual cohort.
-%
-Due to the short paper length, detail is missing on the data types and their algorithms of identifying similar subjects or whether they employ statistical measures.
-%
-We employ the idea of adding variables via drag and drop into a canvas area.
-
-Generalized Pairs Plots (\texttt{GPLOM´S}) are an information visualization technique comparing heterogenous variables pairwise using a plot-matrix grouped by type \cite{GPLOMS, Francois2013}.
-%%
-They are useful to gain an overview over numerous variables and their distributions.
-%%
-Histograms, bar charts, scatter plots and heat maps are used to visualize variable combinations with regard to their type.
-%%
-The resulting matrix provides an \emph{overview visualization}, but requires a lot of screen space for many variables (127 in our application scenario).
-%%
-We incorporate the idea of adaptive type-dependent visualizations.
+%%
+Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+%%
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
 %%
 Dai et al.
-\cite{Dai2005} explored risk factors by incorporating choropleth maps of epidemiological variables (e.g., mortality rates in a region) with parallel coordinates, bar charts and scatter plots with integrated regression lines.
-%%
-Their findings yielded a \emph{Concept Map}, which linked cancer-related associations via graph edges.
-%%
-While their goal to identify possible risk factors using socio-economic and health data is similar to ours, they focus on iteratively refining defined hypotheses and on geographical data.
-%%
-We employ the use of small multiples for incorporating heterogenous data types for comparability.
-%%
-Chui et al. \cite{Chui2011} visualized associations in time-dependent epidemiological data using time-series plots highlighting risk factor differences in age and gender.
-%%
-%The connection of time series plots with histograms and heat maps allows for fast insight into complex relationships and 
-While the work shows how different visualization techniques provide insight into these data sets, it focuses on the time aspect, which is not present for our data.
-%%
-% *********************** Our own Work ***********************
-% Prior Work
+\cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
+%%
+A \emph{concept map} summarized features related to a specified disease.
+%%
+Time-dependent epidemiological data is visualized by Chui et al.
+\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
+%%
+We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
+%%
 \paragraph{Prior Work.}
-We visualized lumbar spine variabilities based on a semi-automatic shape detection algorithm of 490 participants of the \texttt{SHIP-2} cohort \cite{Klemm2013VMV}.
-%
-Hierarchical agglomerative clustering divided the population into shape-related groups.
-%
-As proof of concept, a relation between the size of the segmented shape and the measured size of the subjects was shown.
-%
-This work focuses on incorporating these derived data as new variables, enabling to include it into the hypothesis validation and generation process.
-%
-When applying clustering techniques to the non-image data it was found that \texttt{k-Prototypes} and \texttt{DBSCAN} are appropriate, but are strongly dependent on the chosen variables and distance measures \cite{Klemm2014BVM}.
-%
-Niemann et al. \cite{Niemann2014} presented an interactive data mining tool for the assessment of risk factors of hepatic steatosis, the fatty liver disease.
-%
-Association rules created by data mining methods can be analyzed interactively with their tool and highlight potentially overlooked variables.
-% *********************** / Our own Work ***********************
+In recent publications, we focused on analysis on characterization of the healthy aging process of the lumbar spine.
+%%
+Therefore, we utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
+%%
+The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
+%%
+
+We defined a Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (see \ref{sec:Background}) with a iterative analysis loop incorporating group selection and visualization using expert-input as well as clustering methods \cite{Klemm2014VIS}.
+%%
+Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image-and non-image information.
+%%
+The popularity of the mosaic plot depicting \emph{Cram\'{e}r's V} contingency values with our domain experts was the inspiration for this work.
+%%
+The representation presented in \cite{Klemm2014VIS} only highlights pairwise correlation between features without a target disease.
+
+%% TODO: Check DLPD for correct page numbers
+While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
+%%
+These measures were analyzed using a novel decision tree quality plot towards non-image data.
+%%
+We concluded that the model quality is not enough to characterize back pain.
+\\\\
+The difference of our presented approach towards our previous and other herein presented work is twofold:
+\begin{enumerate}
+	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
+\end{enumerate}
+\com{Main Difference to our and other related work}
 \section{Regression Cube Analysis of Cohort Study Data}
@@
 %%
-\texttt{OpenCPU} is a \texttt{R} package and provides a API for accessing it via HTTP calls.
+\texttt{OpenCPU} is a \texttt{R} package and provides a API for accessing it via HTTP calls \cite{Ooms}.
 %%

### 2015-03-24 draft.tex
gave related work proper structure

@@
 \paragraph{Visual Analysis in Public Health.}
-\com{Uli Niemann, Tommy Hielscher, Myra Spiliopoulou, Henry Völzke, and Jens-Peter Kühn. Can we classify the participants of a longitudinal epidemiological study from their previous evolution?.}
-%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
-Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis (fatty liver disease) using decision trees with a interactive data mining tools.
-%%
-They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
-%%
-We try to reproduce their results and investigate their findings further.
-%%
-\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
-
-\com{Uli, can you write a short summary on these?}
-
-In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
-%%
-They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
-%%
-We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
-
+%%--
 Shneiderman et al.
@@
 
+%%
+In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
+%%
+They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
+%%
+We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
+
+%%
 Schreck and Keim \cite{Schreck} presented a method for analyzing social media data with focus on characterization and cause of an epidemic.
@@
 
-\paragraph{Statistical Analysis}
-
-%\com{The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems}
+%%
+Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+%%
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
+%%
+Dai et al.
+\cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
+%%
+A \emph{concept map} summarized features related to a specified disease.
+%%
+Time-dependent epidemiological data is visualized by Chui et al.
+\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
+%%
+We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
+%%
+
+\paragraph{Statistical based Analysis.}
+%%
+Statistical analysis are an inherent part of Visual Analysis.
+%%
+Here we present work showing a strong focus on statistical analysis in this context.
+%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
+Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
+%%
+Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
+%%
+%% TODO: Use this one?
+"Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich ändert, wenn man per Zufall die Populationen verändert, um zu schauen wie zuverlässig diese Statistik ist.
+%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
 Ahmadi et al.
@@
 %%
-They incorporate scatter plot matrices (SPLOMs) of quality metrics describing the features.
+They incorporate scatter plot matrices (\texttt{SPLOM´S}) of quality metrics describing the features.
 %%
@@
 
-% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
-Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
-%%
-%According to their systematization we analyze correlations.
-%%
-Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
-
-\paragraph{Visual Analysis of Heterogenous Data.}
-%%
-Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
-%%
-The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
-%%
-Dai et al.
-\cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
-%%
-A \emph{concept map} summarized features related to a specified disease.
-%%
-Time-dependent epidemiological data is visualized by Chui et al.
-\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
-%%
-We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
-%%
+Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
+\com{Uli, can you please complete this, I do not have the pdf available!}
+%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
+Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with a interactive data mining tools.
+%%
+They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
+%%
+We try to reproduce their results and investigate their findings further.
+%%
+%\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
+
 \paragraph{Prior Work.}
@@
 \\\\
-The difference of our presented approach towards our previous and other herein presented work is twofold:
+The difference of the presented approach towards our previous and other herein presented work is twofold:
 \begin{enumerate}

### 2015-03-24 draft.tex
revised using Bernhards notes

@@
 %% Paper title.
+%% TODO: Replace towards with regarding?
 
@@
 \maketitle
+\com{Bernhard: Comment on prior occurrences of "Regression Cube" and how we differ}
 %%
@@
 %%
-They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide spread diseases.
+They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide-spread diseases.
 %%
@@
 %%
-The acquired features range from social and lifestyle factors to prior or current diseases and medications as well as medical parameters, such as blood pressure and also comprises of non-radiating medical image data.
+The acquired features range from social and lifestyle factors to prior or current diseases and medications as well as medical parameters, such as blood pressure and also comprises non-radiating medical image data.
 e.g.
@@
 %%
-Medical image data quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
+Medical image data may be quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
 %%
@@
 %%
-Due to missing overview techniques, possibly interesting correlations lie within the data, but are not made apparent.
-%%
-Explorative analyses and first overview visualizations of the data set as presented by Klemm et al.
-\cite{Klemm2014VIS} are not custom tailored to a specific target variable and mostly highlights correlations between variables, which are known to the domain expert (e.g.
+Due to missing overview techniques, possibly interesting correlations lie within the data, but are not apparent.
+%%
+Explorative analyses and overview visualizations of the data set as presented by Klemm et al.
+\cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g.
 correlation between body size and spine shape).
 %%
-We incorporate the regression analysis, which is familiar to the domain experts into a overview visualizations, which can either be used for an hypothesis-free analysis or a analysis towards a specific disease or hypotheses.
+We incorporate the regression analysis, which is familiar to the domain experts into overview visualizations, which can either be used for an hypothesis-free analysis or an analysis towards a specific disease or hypotheses.
 %%
@@
 \begin{itemize}
-	\item A overview visualization technique describing feature interactions using target features.
-	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for detailed investigations of feature relationships.
+	\item An overview visualization technique representing feature interactions using target features.
+	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
 	\item freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
@@
 \subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
-Epidemiology unites experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
-%%
-Their goal is derive disease-specific risk factors by assessing epidemiological features using statistical methods.
-%%
-As described by Thew et al. \cite{Thew2009} it is divided into these different steps:
+Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
+%%
+Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
+%%
+As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
 \begin{itemize}
@@
 %%
-Reproducibility of results is a epidemiological key requirement and guides all analyses steps.
-%%
-Statistical programs, such as \texttt{SPSS} are used to analyzed the data using the classical sequential epidemiological workflow, which uses images largely only to communicate results, rather than providing insight.
-%%
-A alternative data driven approach is described as follows.
-
-\subsubsection{Data Driven Hypothesis-Generation-based Analysis}
-Klemm et al. describe an approach for a Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
+Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
+%%
+Statistical programs, such as \texttt{SPSS} are used to analyze the data using the classical sequential epidemiological workflow, which uses images largely only to communicate results, rather than providing insight.
+%%
+An alternative data-driven approach is described as follows.
+
+\paragraph{Data Driven Hypothesis-Generation-based Analysis.}
+We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
 %%
@@
 %%
+\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
+%%
 \emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
@@
 %%
-Subject data are collected with the widest range possible, allowing the data set to be assessed towards different diseases.
-%%
-The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
+Subject data are collected either towards a specific disease or with the widest range possible.
+%%
+The latter allows the data set to be assessed towards different diseases.
+%%
+The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
+These features are derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
 %%
@@
 %%
-Other features are exclusive, such as women-specific questions, e.g. number of born children or period status.
+Other features are exclusive for a sub-group, such as women-specific questions, e.g. number of born children or period status.
 
@@
 \paragraph{Image Data.}
+\com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
 %%
@@
 %%
-For ethical reasons, the imaging modalities must not include ionizing radiation.
-%%
-The image quality is often inferior to clinical standard, which is a tradeoff between time and cost \cite{Preim2014}.
+For ethical reasons, the imaging modalities usually not include ionizing radiation.
+%%
+The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
 %%
@@
 %%
-Automatic or semi-automatic solutions bypass this problem, but are costly and need to be custom tailored to the structure of interest \cite{Toennies2015}.
+Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
 %%
@@
 %%
-\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely to happen with increasing age.
+\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
 %%
@@
 %%
-Regression analyses are the most important statistical tool when analyzing epidemiological data.
+Regression analysis is the most important statistical tool when analyzing epidemiological data.
 %%
@@
 For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
+%%
+However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study ("\emph{method loyality}").
 %%
@@
 %%
-The basic idea of our \emph{Regression Cube} is to provide a overview visualization of large cohort study data sets towards target variables.
-%% TODO BLIND
-Overview visualizations of feature relationships as presented by Klemm et al. \cite{Klemm2014VIS} are often focused on relationships between the visualized features.
+The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target variables.
+%%
+Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
 %%
@@
 %%
-In epidemiology, these relationships are also of interest, but rather towards their explanatory power towards the target feature.
+In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
 %%
@@
 
-Since we want to use the regression analyses with a overview visualization, we are are interested in all possible combinations of (two or more) independent features describing a target.
+Since we want to use the regression analyses with an overview visualization, we are are interested in all possible combinations of (two or more) independent features describing a target.
 %%
@@
 %%
-In a data set with 100 features, the regression formula
+In a data set with $n$ (e.g. 100) features, the regression formula
 \begin{equation}
@@
 \end{equation}
-would yield 10.000 regression models, describing all possible combinations of two features in the data describing the target $Cancer$.
+would yield $n^2$ (10.000) regression models, describing all possible combinations of two features in the data describing the target $Cancer$.
 %%
@@
 %%
-The number calculated regression models increases exponentially for each dynamic variable added.
+The number of calculated regression models increases exponentially for each dynamic variable added.
 %%
@@
 %%
-The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for achieving this task \cite{CFS}.
+The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for this task \cite{CFS}.
 %%
@@
 %%
-When for example the \emph{body weight} has a strong explanatory power towards the target it is likely that \emph{BMI} or \emph{waist circumference} behave similar towards it.
+When for example the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} behave similar towards it.
 %%
@@
 For features, such as age or gender, which affect most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}) the number is significantly larger (about half the features in the whole data set).
+%%
+The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
+%%
+\com{Uli, is this correct this way?}
 %%
@@
 %%
-The goal of a overview visualization is to provide a comprehensive view on the data, which is easy to understand.
+The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
 %% TODO: Cite another paper here!
@@
 %%
-For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings is required and the $z$ dimension can be compared directly.
+For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings are required and the $z$ dimension can be compared directly.
 %%
@@
 \noindent Design choices and spaces are restricted by the underlying technologies and devices.
-%%
+%% TODO: Umschreiben
 Resting collaboration and exchange about data and methods on web-based technologies shows promising results, as there is no set-up time involved and domain experts can use the methods from any computer connected with the web.
@@
 %%
-The design space of web-technologies is different to standard WIMP-applications--right click menus, modal windows and menu bars are not established UI components in this context.
+The design space spanned by web-technologies is different to standard WIMP-applications--right click menus, modal windows and menu bars are not established UI components in this context.
 %%
@@
 %%
-The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of ui elements as much as possible.
+The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
 %%
@@
 %%
-To allow other users even outside the epidemiological application domain access to our tool, we choose to do this using standard ASCII-based CSV files.
+To allow other users even outside the epidemiological application domain to access to our tool, we choose to do this using standard ASCII-based CSV files.
 %%
@@
 %%
-Security issues are raised by uploading data into a online service, such as our prototype.
-%%
-The use of epidemiological data is preceded by a detailed description on the analysis purpose and has to be approved by ethics committees.
+Security issues are raised by uploading data into an online service, such as our prototype.
+%%
+The use of epidemiological data is preceded by a detailed description of the analysis purpose and has to be approved by ethics committees.
 %%
@@
 %%
-We aim to minimize these advantages as much as possible.
+We aim to minimize these problems as much as possible.
 %%
@@
 %%
-The current selected slice is denoted with a semi-transparent gray plane.
+The currently selected slice is denoted with a semi-transparent gray plane.
 %%
@@
 %%
-To assess the ability a system to discover knowledge discover knowledge is hard to measure.
+To assess the ability of a system to discover knowledge is hard to measure.
 %%
@@
 %%
-We conducted a study using two participants, who also co-authored this publication.
+We conducted a study with two participants, who also co-authored this publication.
 %%
@@
 %%
-Correlation based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
+Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
 %%
@@
 %%
-Two case studies revealed that the presented approach is able to reproduce knowledge extracted using decision tree based data mining methods on a hepatic steatosis data set as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
+Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
 %%
@@
 The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
-This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.
-We thank Marko Rak and Klaus Toennies for providing the image detection data.}
+This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.}
 % \end{small}

### 2015-03-25 draft.tex
added Kais revision

@@
 %%
-Epidemiology aims to characterize health and disease conditions in defined populations (\emph{Cohorts}).
+Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
 %%
@@
 
-An important epidemiological tool for deriving such features are \emph{Cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
+An important epidemiological tool for deriving such features are \emph{cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
 %%
@@
 %%
-The acquired features range from social and lifestyle factors to prior or current diseases and medications as well as medical parameters, such as blood pressure and also comprises non-radiating medical image data.
-e.g.
-magnetic resonance imaging (MRI).
-%%
-Medical image data may be quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
-%%
-
-To assess the statistical resilience of a hypotheses using regression analyses rarely involves more than three features due to the required subject count.
-%%
-Due to missing overview techniques, possibly interesting correlations lie within the data, but are not apparent.
+The acquired features comprise a wide range.
+%
+Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure are gathered.
+%%
+Medical image data from non-radiating modalities. e.g. from magnetic resonance imaging (MRI) is also acquired in modern studies.
+%%
+These data may be quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
+%%
+
+Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
+%%
+Using it to assess the statistical resilience of a hypotheses rarely involves \emph{more than three features} due to the required subject count.
+%%
+Due to missing overview techniques, possibly interesting correlations may be missed.
 %%
@@
 %%
-We incorporate the regression analysis, which is familiar to the domain experts into overview visualizations, which can either be used for an hypothesis-free analysis or an analysis towards a specific disease or hypotheses.
+We incorporate the regression analysis, which is familiar to the domain experts into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypotheses.
 %%
@@
 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
-	\item freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
-	\item analysis of confounding variables by providing comparison views between different formula results
-	\item The open and web based approach of the system allows for analysis of any data using the presented method.
+	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
+	\item Analysis of confounding variables by providing comparison views between different formula results
+	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
 \end{itemize}
@@
 \begin{itemize}
-	\item Clinicians make observations in the daily practice which are translated into hypotheses.
+	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
 	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
@@
 %%
-Other features are exclusive for a sub-group, such as women-specific questions, e.g. number of born children or period status.
+Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
 
@@
 %%
-These data are hard to analyze as they require segmentation highlighting the structures of interest.
-%%
-This process is prone to inter and intra-observer variability when carried out manually.
+These data are hard to analyze as they require segmentations to outline structures of interest.
+%%
+This process is prone to inter- and intra-observer variability when carried out manually.
 %%
@@
 \paragraph{Confounding Features}
-Features influencing the exposure as well as the outcome of a analysis are called confounders.
+Features influencing the exposure as well as the outcome of an analysis are called confounders.
 %%
@@
 %%
-It also influences the general body condition and therefore almost all features acquired through cohort studies.
+It also influences the general body condition and thereby almost all features acquired through cohort studies.
 %%
@@
 %%
-Other confounder have to be selected by epidemiologists specific to the investigated condition.
+Other confounders have to be selected by epidemiologists specific to the investigated condition.
 
@@
 %%
-Metrics, such as the weightings itself and associated p values and describe the resulting function (the \emph{model}).
-%%
-R$^2$ values describe the goodness of fit; in other words how well the dependent features describe the target feature.
-%%
-The value is in the range [0,1], while 1 encodes a perfect fit.
+Different metrics, such as the weightings itself and associated $p$ values describe the resulting function (the \emph{model}).
+%%
+$R^2$ values describe the goodness of fit; in other words how well the dependent features describe the target feature.
+%%
+The value is in the range [0,1], where 1 encodes a perfect fit.
 
@@
 %%
+\com{Kai, + in this case is a regression operator including a feature. The formula should denote any operator (maybe as Var1 omega Var2).}
 \begin{equation}
-Dependent \sim Independent_1 + Independent_2 + ... + Independent_n
+Dependent \sim Independent_1 + ... + Independent_n
+%Dependent \sim \sum_{i=1}^n Independent_i
 \label{eq:RegressionNotation}
@@
 \begin{itemize}
-	\item $+$ / $-$ inclusion/exclusion of the variable,
+	\item \fbox{$+,-$} inclusion/exclusion of the variable,
 	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
@@
 
-\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating linear weightings for the \emph{independent} features.
-%%
-The dependent variable has to be of a continuous type.
+\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
+%%
+The \emph{dependent} variable has to be of a continuous type.
 %%
@@
 %%
-Logistic models do as opposed to linear models does not allow for extracting a R$^2$ goodness of fit value.
-%%
-Therefore, pseudo-R$^2$ values are extracted, such as the \emph{Nagelkerke R$^2$}, which mimics the behavior of the R$^2$.
-%%
-\emph{Nagelkerke R$^2$} cannot be compared to R$^2$ values extracted from linear regression model.
+Logistic models as opposed to linear models do not allow for extracting an $R^2$ goodness of fit value.
+%%
+Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
+%%
+\emph{Nagelkerke $R^2$} cannot be compared to $R^2$ values extracted from linear regression model.
 
@@
 \subsection{The Study of Health in Pomerania (SHIP)}
-The \texttt{SHIP}, located in Northern Germany aims to aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+The \texttt{SHIP}, located in Northern Germany aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
 %%
@@
 %%
-Data for both cohorts are examined in a in a 5-year time span.
+Data for both cohorts are examined in a 5-year time span.
 %%
@@
 %%
-The method then replaces the dynamic variables with all features in the data set.
+Our method replaces the dynamic variables with all features in the data set.
 %%
@@
 %%
-The problem with this brute-force approach lies in its complexity.
+The problem with this approach lies in its complexity.
 %%
@@
 
-\subsection{Abstracting Regression Results using R$^2$}
+\subsection{Abstracting Regression Results using $R^2$}
 \begin{figure}[htb]
@@
  %%
- The R$^2$ values extracted from the regression formulas depict the goodness of fit and are mapped to color saturation.
+ The $R^2$ values extracted from the regression formulas depict the goodness of fit and are mapped to color saturation.
  %%
@@
  %%
- The R$^2$ values of each slice voxel is mapped on opacity in the 3D-view, reducing the occlusion of other values.
+ The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view, reducing the occlusion of other values.
  }
@@
 %% TODO: Cite another paper here!
-As described in \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
+As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
 %%
@@
 %%
-As described in Section~\ref{sec:RegressionAnalysis}, the R$^2$ value is the metric allowing for this kind of assessment.
+As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
 %%
@@
 %%
-Since R$^2$ is scaled between [0,1] it also allows for comparison between regression models.
-%%
-We can apply the same mosaic plot mapping by translating the R$^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
+Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
+%%
+We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
 %%
@@
 %%
-R$^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
-%%
-The visualization of R$^2$ values derived from different regression cubes (e.g. $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
-%%
-Therefore, the R$^2$ results of different regression methods are encoded using different colors (e.g. blue for linear regression and red for logistic regression).
+$R^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
+%%
+The visualization of $R^2$ values derived from different regression cubes (e.g. $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%%
+Therefore, the $R^2$ results of different regression methods are encoded using different colors (e.g. blue for linear regression and red for logistic regression).
 %%
@@
 \\\\
-The goal is to create a overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
+The goal is to create an overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
 %%
@@
 \subsection{Analysis Workflow}
+%% TODO: Anmerkung Kai einarbeiten Ja, gefällt mir auch nicht so gut.
+Ich würde zunächst Keims Mantra vorstellen (mit itemize) dann dazu erklären was er damit meint und was es bedeutet.
+Diese Mantrapunkte nennst du M1-M4 (oder ein anderen Buchstaben).
+Und dann würde ich bei der Erklärung deines Tools immer wieder darauf verweisen, welcher Punkt nun erfüllt ist (..., which fulfills M2)
+\com{ToDo. Include Kais revision, renaming Matnrapoints and refer to them}
 \com{I'm not satisfied with the structure in this subsection, please proofread it with focus in the train of thought!}
@@
 
-\textbf{Show the Important.} The 3D-cube visualization acts as overview over the whole data set.
-%%
-Regression models with large R$^2$ values can be spotted fast here, steering the attention to the respective slice.
+\textbf{Show the Important.} The 3D-cube visualization acts as an overview over the whole data set.
+%%
+Regression models with large $R^2$ values can be spotted here fast, steering the users attention to the respective slice.
 
@@
 %%
-The absolute difference in R$^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
+The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
 %%
@@
 
-$Z \sim X + Y$ is the default if the user did not specify a formula.
+$Z \sim X + Y$ is the default formula.
 %%
@@
 %%
-The interaction between $feature_1$ and $feature_2$ is also included in all regression formulas.
-%%
-
-$Cancer \sim X + Y + Z$ subtracted with results from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
+% The interaction between $feature_1$ and $feature_2$ is also included in all regression formulas.
+%%
+This formula models the hypothesis, of a interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result, hence the cube using $x$,$y$ and $z$ as independent features.
+%%
+
+$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
 %%
@@
 %%
-The focus lies on creating a overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
-%%
-Therefore the system has to be intuitive and comprehensive in order to be adapted by domain experts.
+The focus lies on creating an overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
+%%
+Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
 %%
@@
 %%
-Web technology is also inherently designed with client-server architecture in mind, making it easy to outsource computation heavy tasks on server-clusters and transferring results to the client device.
-%%
-The design space spanned by web-technologies is different to standard WIMP-applications--right click menus, modal windows and menu bars are not established UI components in this context.
-%%
-Therefore we have to adapt our design respecting the affordances of web pages.
+Web technology is by design based on a client-server architecture, making it easy to outsource computationally heavy tasks on server-clusters and transferring results to the client device.
+%%
+The design space spanned by web-technologies is different to standard WIMP-applications: right click menus, modal windows and menu bars are not established UI components in this context.
+%%
+Therefore, we have to adapt our design respecting the affordances of web pages.
 %%
@@
 	%%
-	\item The \emph{file upload} section starting the analysis with providing a comma separated values (CSV) data set.
-	%%
-	\item The \emph{formula editor} allows to adjust the formula towards a current hypothesis or conducting a \emph{hypothesis-free} analysis
+	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file.
+	%%
+	\item The \emph{formula editor} allows to adjust the formula towards a current hypothesis or conducting a \emph{hypothesis-free} analysis.
 	%%
@@
 %%
-Popular analytics tools, such as WEKA \cite{WEKA} owe their popularity to their support of open file types.
-%%
-To allow other users even outside the epidemiological application domain to access to our tool, we choose to do this using standard ASCII-based CSV files.
+Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
+%%
+To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
 %%
@@
 %%
-Missing values are denoted using no character at all, a whitespace, or an empty quotation mark encapsulated string.
+Missing values are denoted by using no character at all, a whitespace, or an empty quotation mark encapsulated string.
 
@@
 %%
-We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
-%%
-Instead we apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
+%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
+%%
+We apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
 %% TODO: Implement this
-Data sets are deleted from the server after the closing a session.
+Data sets are deleted from the server after closing a session.
 
@@
 %%
-On entering the text are, a context modal displays all data set variables as well as the available operators and their function.
+On entering the text are, a context modal displays all data set variables, as well as the available operators and their function.
 %%
@@
 %%
-Auto-completing input variables also simplifies the approach and also works as spell-check of (sometimes hard to memorizable) variable names.
-%%
-
-\textbf{Formula Validation and Calculation.} The formula is checked for validity directly on input, indicating invalid formulas with a red halo around the text input, which turns to green for valid input.
-%%
-This prevents processing errors on the statistical processor backend.
+Auto-completing input variables also simplifies the approach and also works as spell-check of variable names.
+%%
+
+\textbf{Formula Validation and Calculation.} The formula is checked for validity directly on input.
+%%
+The text input containing the formula is marked using a red halo to indicate invalid input, which turns green for valid formulas.
+%%
+This prevents processing errors on the statistical processor back-end.
 %%
@@
 %%
-These are then divided by the number of available statistical backend processors available, driving a \emph{cloud computing} based approach.
+These are then divided by the number of available statistical back-end processors available, driving a \emph{cloud-computing}-based approach.
 %%
@@
 %%
-If the user selects a reference formula, the absolute difference of R$^2$ values between the cube described by the reference and the currently active formula are calculated and shown.
+Since all cells in the cube are represented using $R^2$ values, a difference cube is calculated by denoting the absolute difference of $R^2$ for each cell.
+%%
+%%If the user selects a reference formula, the absolute difference of $R^2$ values between the cube described by the reference and the currently active formula are calculated and shown.
 %%
@@
 %%
-Figure~\ref{fig:Cube} shows that all values are mirrored along the mosaic plot matrix diagonal.
+Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the mosaic plot matrix.
 %%
@@
 %%
+Along the diagonal, $X$ and $Y$ assume the same feature, $Z ~ X + Y$ turns into $Z ~ X$ because the regression automatically ignores doublings.
+%%
+The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.
 
@@
 %%
-The 3D-cube representation acts overview over the whole data set, but its purpose is not to derive detailed information about data points.
-%%
-It serves a function similar to a mini-map, guiding the attention towards points of interest in the data as well as giving context information about adjacent data values when using the 2D mosaic plot.
-%%
+The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
+%%
+It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
+%%
+\com{TODO Mantra}
 It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
@@
 \paragraph{Tackling the disadvantages of 3D information visualization.}
-3D-Information visualizations are often criticized for introducing occlusion and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
+3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
 %%
@@
 %%
-Since the R$^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
-%%
-It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low R$^2$ values.
+Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
+%%
+It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
 %%
@@
 %% INTERACTION
-Rotating with the cube is restricted to y axis, preserving the mental map of the position of the individual features.
-%%
-The cube is always oriented according to the 2D-representation, allowing for a easy mental combination of the two representations.
+Rotating with the cube is restricted to y-axis, preserving the mental map to position individual features.
+%%
+The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
 %%
@@
 %%
-the values are projected on this plane to give a overlapping-free view on the data points, which makes it easier to identify the current slice.
+the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
 %%
@@
 \paragraph{2D Mosaic Plot Slice Visualization.}
+\com{Add Figure reference!}
 %%
@@
 %%
-By hovering over a data entry in the plot, a tooltip model displays detailed information about a models coefficients, associated p values and confidence intervals.
+By hovering over a data entry in the plot, a tooltip model displays detailed information about a models coefficients, associated $p$ values and confidence intervals.
 %%
@@
 %%
-Relying on web technologies paid off in previous collaboration with our clinical exerts, since it allows for:
+Using web technologies for prototyping allows for:
 \begin{itemize}
@@
 %%
-The front end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
+The front-end is created using \texttt{HTML5}, \texttt{CSS3} and \texttt{Javascript}.
 %%
@@
 %%
-The 2D mosaic plot is implemented using \texttt{Data driven Documents}(\texttt{D3.js}), which is very popular in information visualization using vector graphics \cite{D3}.
+The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is very popular in information visualization using vector graphics \cite{D3}.
 %%
@@
 %%
-Proper open source volume rendering methods are not available for the web, the effort of writing one from scratch was not worth the counter value.
+Open source volume rendering methods are available but do not satisfy our requirements.
 %%
@@
 %%
-The backend functions necessary for all cube calculations are provided via an \texttt{R} package.
+The back-end functions necessary for all cube calculations are provided via an \texttt{R} package.
 %%
@@
 %%
-The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}.
-%%
-Instructions and code on how to create a setup running the \emph{Regression Cube} statistical backend through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
+The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}.
+%%
+Instructions and code on how to create a setup running the \emph{Regression Cube} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
 %%
@@
 %%
-The breast fat data set is the foundation for a explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
+The breast fat data set is the foundation for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
 %%
@@
 Lam et al.
-\cite{Lam2012} propose for this purpose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems ability to generate hypotheses and explore the data in order to extract information.
+\cite{Lam2012} propose for this purpose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
 %%
@@
 %%
-One major advantage of focusing on web based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
-%%
-We conducted a web based research by using a online-meeting software, which features voice chat as well as screen sharing.
-%%
-Starting a analysis using these techniques take about 5-10 minutes of setup time.
+One major advantage of focusing on web-based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
+%%
+We conducted a web-based research by using an online-meeting software, which features voice chat as well as screen sharing.
+%%
+Starting an analysis using these techniques take about 5-10 minutes of setup time.
 
@@
 \begin{itemize}
-	\item \emph{somatometric features}, e.g. body size, weight or BMI, 
-	\item \emph{lifestyle features}, e.g. alcohol/tobacco consume, 
-	\item \emph{personal history}, e.g. occupation, marital status,
-	\item \emph{medical history}, e.g. current or prior diseases or laboratory values, such as cholesterol levels,
-	\item \emph{women specific features}, e.g. number of born children, contraception type or hormone replacement therapy and
+	\item \emph{somatometric features}, e.g., body size, weight or BMI, 
+	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume, 
+	\item \emph{personal history}, e.g., occupation, marital status,
+	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
+	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
 	\item \emph{mammography features} derived from the image data.
@@
 %%
-All features were renamed to be expressive, e.g. \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
+All features were renamed to be expressive, e.g., \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
 %%
@@
 \section{Summary and Conclusion}
+\com{Rewrite, focus on precise definition on what was done and why it is awesome.}
 We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
 %%
-\emph{Regression Cube} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
+\emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
 %%
@@
 %%
-Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
+Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
 %%
@@
 
-The method was very well received with our project partners, allowing them for the first time to retrieve a overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
+The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
 %%

### 2015-03-25 draft.tex
added more complex revisions of Kai and Bernhard

@@
 \maketitle
-\com{Bernhard: Comment on prior occurrences of "Regression Cube" and how we differ}
 %%
@@
 \end{itemize}
+%%
+The term Regression Cube was used before to define for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
+%%
+We refer to it using as visualization technique for comparing and analyzing a large number of different regression models with different input features.
+%%
+\com{Bernhard, does this suffice as reasoning?}
 
@@
 %%
-\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
+%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
 %%
@@
 Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
-
-\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
-
-\paragraph{Data Types.}
+%%
+%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
+\\\\
+%\paragraph{Data Types.}
 %%
@@
 
-\paragraph{Image Data.}
-\com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
-%%
-Since the Rotterdam study, many modern cohort studies include medical image data.
-%%
-For ethical reasons, the imaging modalities usually not include ionizing radiation.
-%%
-The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
-%%
-These data are hard to analyze as they require segmentations to outline structures of interest.
-%%
-This process is prone to inter- and intra-observer variability when carried out manually.
-%%
-Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
-%%
-Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
+Medical image data is also acquired and analyzed in modern cohort studies.
+%%
+We incorporate image-derived features, but do not put focus on analyzing image data.
+%%
+More discussion of data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
+
+%\paragraph{Image Data.}
+% \com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
+% %%
+% Since the Rotterdam study, many modern cohort studies include medical image data.
+% %%
+% For ethical reasons, the imaging modalities usually not include ionizing radiation.
+% %%
+% The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
+% %%
+% These data are hard to analyze as they require segmentations to outline structures of interest.
+% %%
+% This process is prone to inter- and intra-observer variability when carried out manually.
+% %%
+% Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
+% %%
+% Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
 
@@
 %%
-\com{Kai, + in this case is a regression operator including a feature. The formula should denote any operator (maybe as Var1 omega Var2).}
 \begin{equation}
@@
 %\com{Describe how regression cubes work (including notation).}
-\com{Move this to Regression Cube part?}
+%\com{Move this to Regression Cube part?}
 
@@
 \end{enumerate}
-\com{Main Difference to our and other related work}
+%%
 \section{Regression Cube Analysis of Cohort Study Data}
-\com{Add Data Preprocessing Section.}
+% TODO: Do we need a preprocessing section?
+%\com{Add Data Preprocessing Section.}
 %%
@@
 \subsection{Analysis Workflow}
-%% TODO: Anmerkung Kai einarbeiten Ja, gefällt mir auch nicht so gut.
-Ich würde zunächst Keims Mantra vorstellen (mit itemize) dann dazu erklären was er damit meint und was es bedeutet.
-Diese Mantrapunkte nennst du M1-M4 (oder ein anderen Buchstaben).
-Und dann würde ich bei der Erklärung deines Tools immer wieder darauf verweisen, welcher Punkt nun erfüllt ist (..., which fulfills M2)
-\com{ToDo. Include Kais revision, renaming Matnrapoints and refer to them}
-\com{I'm not satisfied with the structure in this subsection, please proofread it with focus in the train of thought!}
+%%
 \begin{figure}[htb]
@@
  %%
- The analysis starts with a set of features given by the data set.
+ [\textbf{1.AF}] The analysis starts with a set of features given by the data set.
  %%
@@
  %%
- The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
- %%
- Insights into the data yield either an adjustment of the current formula.
+ [\textbf{2.SI}] The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
+ %%
+ [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula.
  %%
@@
  %%
- The analysis yields insights and hypothesis about feature relations.
- %%
+ [\textbf{4.DD}] Details about features using the 2D-heatmap representation yields insights and hypothesis about feature relations.
  }
@@
 
-\textbf{Analyze First.} Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
-
-\textbf{Show the Important.} The 3D-cube visualization acts as an overview over the whole data set.
+\textbf{Analyze First} [\textbf{1.AF}].
+Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
+
+\textbf{Show the Important} [\textbf{2.SI}]. The 3D-cube visualization acts as an overview over the whole data set.
 %%
@@
 
-\textbf{Zoom, Filter and Analyze Further.} The slices of interest can then be analyzed using the 2D mosaic plot of the slice.
-
-\textbf{Details on Demand.} Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
+\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D mosaic plot of the slice.
+
+\textbf{Details on Demand} [\textbf{4.DD}].
+Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
 in a hover-modal on a currently selected data point).
 \\\\
+We use the squared bracket abbreviation for each step to denote the affiliation the system design section later on.
+%%
 As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
@@
 %%
+\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
+%%
+Input formulas reflect \emph{hypotheses} about the data.
+%%
+Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
+%%
+To support \emph{hypothesis-free} analysis by providing a default formula:
+
+$Z \sim X + Y$.
+%%
+This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
+%%
+Each slice represents a different target feature.
+%%
+It is therefore suitable for a explorative analysis to give a general impression about relationships in the data set.
+
+Hypotheses about the data are reflected in either how dynamic variables are related using the regression operators.
+%%
+Also, static features can be added for each regression formula.
+%%
+Here are a few examples:
+
+%%
+$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
+%%
+All combinations of three independent features towards the target are analyzed through the cube.
+
+$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
+%%
+This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result.
+%%
+Therefore, the cube incorporates this interaction for all $X$,$Y$ and $Z$ values as independent features.
+%%
+
+$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
+%%
+This is achieved through cube comparison.
+%%
+
 \paragraph{Cube Comparison.}
@@
 %%
-\com{TODO: divide using analysis approaches, e.g. Hypothesis-based \& Hypothesis-free with follow up on the formulas used?}
-\paragraph{Hypothesis-Free Analysis}
-%%
-\paragraph{Hypothesis-Based Analysis}
-%%
-%%
-\paragraph{Formula Types.}
-%%
-Input formulas reflect \emph{hypotheses} about the data.
-%%
-Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
-%%
-Here we show just a few examples on how formulas can serve different purposes.
-
-$Z \sim X + Y$ is the default formula.
-%%
-This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
-%%
-Each slice represents a different target feature.
-%%
-It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.
-
-$Cancer \sim X + Y + Z + feature_1:feature_2$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
-%%
-% The interaction between $feature_1$ and $feature_2$ is also included in all regression formulas.
-%%
-This formula models the hypothesis, of a interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result, hence the cube using $x$,$y$ and $z$ as independent features.
-%%
-
-$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
-%%
+It can for example be utilized for comparing the influence of one feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
+%%
+% \paragraph{Formula Types.}
+% %%
+% Input formulas reflect \emph{hypotheses} about the data.
+% %%
+% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
+% %%
+% Here we show just a few examples on how formulas can serve different purposes.
+%
+% $Z \sim X + Y$ is the default formula.
+% %%
+% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
+% %%
+% Each slice represents a different target feature.
+% %%
+% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.
 
@@
 	%%
-	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file.
+	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file [1.AF].
+	%%
+	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube [2.SI].
 	%%
@@
 	%%
-	It also allows to select a reference formula for creating a difference-cube.
-	%%
-	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube.
+	It also allows to select a reference formula for creating a difference-cube [3.ZF].
 	%%
@@
 %%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
-\paragraph{File-Upload and Classification.}
+\paragraph{File-Upload and Classification [1.AF].}
 %%
@@
 
-\paragraph{Formula Editor.}
+\paragraph{Formula Editor [1.AF, 3.ZF].}
 %%
@@
 
-\subsection{Regression Cube Visualization.}
+\subsection{Regression Cube Visualization [2.SI].}
 %%
@@
 %%
-\com{TODO Mantra}
 It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
@@
 %%
-\\\\
 %% TODO: Add reference for disadvantages of 3D information visualizations!
@@
 %\\\\
-\paragraph{Cube Slice Selection.}
+\paragraph{Cube Slice Selection [3.ZF].}
 %%
@@
 %\\\\
-\paragraph{2D Mosaic Plot Slice Visualization.}
+\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
 \com{Add Figure reference!}
@@
 \section{Summary and Conclusion}
-\com{Rewrite, focus on precise definition on what was done and why it is awesome.}
 We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
 %%
-\emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
-%%
-Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
-%%
-These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
-%%
-Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
-%%
-The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
-%%
-All code associated to this project is also open source.
-%%
-Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
-%%
-
-As future work, we want to introduce more regression models to the data set, which model different correlation types.
+Dimension reduction using the target restricts the analysis to most important features.
+%%
+\emph{Hypothesis-free} analysis employs default regression models.
+%%
+Model expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
+%%
+A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
+%%
+These can then be analyzed further using the 2D mosaic plot for each cube slice.
+%%
+Details on demand for each model allow for detailed assessment of regression models.
+%%
+We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
+%%
+The method was well received by our clinical partners, triggering detailed investigations of the findings.
+
+As next step, we want to introduce more regression types, which model different kinds of correlations.
 %%
@@
 
-The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
-%%
-The observed relationship are now subject of detailed statistical analyses.
-%%
-Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
+We published all associated code and provide an analysis platform open to heterogenous data types.
 %%
 We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
+
+% \com{Rewrite, focus on precise definition on what was done and why it is awesome.}
+% We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
+% %%
+% \emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
+% %%
+% Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
+% %%
+% These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
+% %%
+% Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
+% %%
+% The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
+% %%
+% All code associated to this project is also open source.
+% %%
+% Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
+% %%
+%
+% As future work, we want to introduce more regression models to the data set, which model different correlation types.
+% %%
+% We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
+% %%
+%
+% The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
+% %%
+% The observed relationship are now subject of detailed statistical analyses.
+% %%
+% Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
+% %%
+% We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
 %%

### 2015-03-25 draft.tex
added Ulis data description and revisions

@@
 
-Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
-\com{Uli, can you please complete this, I do not have the pdf available!}
+%Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
+%\com{Uli, can you please complete this, I do not have the pdf available!}
 %% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
@@
 They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
+%%
+In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{Evolution Features}) that describe latent temporal information across the study waves.
 %%
@@
 \subsection{Target-Variable-dependent Dimension Reduction}
-\com{Uli, Can you please proofread this?}
-%%
-The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
-%%
-Identifying these features und excluding them from the calculation can reduce the number of dimension significantly.
-%%
-The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for this task \cite{CFS}.
-%%
-It uses information entropy to select the features which have the most explanatory power towards the target feature.
-%%
-At the same time it tries to reduce the correlation between the independent features.
+% \com{Uli, Can you please proofread this?}
+% %%
+% The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
+% %%
+% Identifying these features und excluding them from the calculation can reduce the number of dimension significantly.
+% %%
+% The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for this task \cite{CFS}.
+% %%
+% It uses information entropy to select the features which have the most explanatory power towards the target feature.
+% %%
+% At the same time it tries to reduce the correlation between the independent features.
+% %%
+In epidemiological studies, manifold recordings lead to an abundance of features.
+%%
+Generally many of them exhibit low or none correlation towards the target feature.
+%%
+Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and contributes to presentiveness of the Regression Cube.
+%%
+The Correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in F.
+%%
+The dependency of a set of features utilizes the entropy-based information gain to measure the explanatory power towards the target feature.
+%%
+Beginning with an empty set of features $F$ the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when adding any feature does not increase the merit.
 %%
@@
 \subsection{The Hepatic Steatosis Data Set}
-\com{Uli: Describe the data set and also prior work on it.}
-%%
-\subsubsection{Data Preprocessing}
+We employ a subset of the data set which was used in \cite{Niemann2014} to identify predictive features with respect to the reversible disorder hepatic steatosis (fatty liver).
+%%
+The binary target feature is derived from the liver fat concentration which was measured with Magnetic Resonance Imaging (MRI).
+%%
+In particular, participants with a liver fat concentration of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
+%%
+The data set contains labels for 578 participants for which the MRI data was available to that point.
+%%
+Note that MRI was performed only in SHIP-2, so the target feature is present for just one of three study waves.
+
+%%
+Apart from the target feature, the data set contains 199 features extracted from participants' questionnaire answers and medical tests.
+%%
+They are features on socio-demographics (sex, age, etc.), features on consumption behavior (e.g.
+alcohol and cigarettes), SNPs, features extracted from laboratory data (e.g.
+sera concentrations), and two features on the results of the liver ultrasound.
+%%
+The names of 85 features have the suffix \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features exhibit the suffix \emph{s1} and 55 features have the suffix \emph{s2}, alongside 10 time-independent SNPs.
+%%
+
+In \cite{Niemann2014}, the authors show that women and men exhibit differences in the class distribution of the liver fat concentration.
+%%
+Also, for women they identify that there is a association between age and liver fat concentration.
+%%
+They compute an appropriate cut-off value of 52 years for which the class distribution is most homogeneous within the resulting subsets.
+%%
+Based on this observations, we perform our analysis on three sub-populations, $PartitionF$, $PartitionM$ and \emph{$F:age\>52$}.
 %%

### 2015-03-26 draft.tex
added details on the application section

@@
 %%
-\paragraph{Setup and Procedure.}
+\paragraph{Participants, Setup and Procedure.}
 %%
@@
 All sessions were video recorded to be processed later on.
+%%
+
+Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.
+%%
+KH is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
+
+The hepatic steatosis data set is analyzed by UN, a data scientists responsible for prior analysis of the data with.
+%%
+
+TI, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
 %%
@@
 %%
-It contains 230 dimensions, holding information about
+It contains 231 dimensions, holding information about
 \begin{itemize}
@@
 %%
+The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
+%%
+Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
+%%
+
 \subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
@@
 \subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
+%%
+\textbf{Extracted Hypotheses have to investigated further.}
+\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
+%%
+
+\textbf{Overview Visualizations gain importance with increasing feature count.}
+%%
+While the classical epidemiological workflow (see Sec.~\ref{sec:Background}) represents the foundation of the research in this area, explorative analysis based on the data gains importance.
+%%
+Incorporating data mining algorithms with statistically established and overview visualizations allow for reasoning about risk factors using underlying epidemiological data sets.
+%%
+One major advantage we observed is the confidence of an expert in the results he or she actually observe.
+%%
+Their participation in the analysis using human pattern-detection and expert knowledge is for experts preferred over results derived using a black-box method.
+%%
+By observing expected correlations matching the expert knowledge strengthens the confidence in the method and subsequently in the hypothesis generated from unexpected relationships.
+%%
+%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms.
+People believe more in what they see, rather than in what they are presented by an algorithm.
+Regression cubes incorporate this by showing known associations as well as new ones.
+Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}
+
+%%
+\textbf{Using non-discretized features reduces information bias.}
+%%
+\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
+%%
+
+\textbf{Support for different subject groups.}
+\com{Include this?}
 %%
@@
 \section{Summary and Conclusion}
+%%
 We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.

### 2015-03-26 draft.tex
start of Sylvias revision
added first part of Sylvias revision

@@
 \item
- Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,niemann,preim\}@ovgu.de
+ Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,niemann,preim\}@ovgu.de
 \item
@@
 %%
-These data contain information (\emph{features}) about a subjects' lifestyle, medical conditions and also medical image data.
+These data contain information (\emph{features}) about a subject's lifestyle, medical condition as well as medical image data.
 %%
@@
 %%
-%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time a interactive regression-based analysis of large feature sets with respect to a disease.
+%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract
@@
 %%
-They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide-spread diseases.
-%%
-In the standard workflow, physicians translate observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.
-
-An important epidemiological tool for deriving such features are \emph{cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
-%%
-To reduce any selection bias, subjects are invited at random and without a focus on a specific disease.
-%%
-The acquired features comprise a wide range.
+Furthermore, they can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
+%%
+During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.
+
+An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
+%%
+To reduce any selection bias, subjects are randomly invited without a focus on a specific disease.
+%%
+Hence, a wide range of features is acquired.
 %
-Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure are gathered.
-%%
-Medical image data from non-radiating modalities. e.g. from magnetic resonance imaging (MRI) is also acquired in modern studies.
-%%
-These data may be quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
+Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure, are gathered.
+%%
+Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
+%%
+These data may be quantified using user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
 %%
@@
 %%
-Using it to assess the statistical resilience of a hypotheses rarely involves \emph{more than three features} due to the required subject count.
-%%
-Due to missing overview techniques, possibly interesting correlations may be missed.
+Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
+%%
+% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
+Due to the amount of data and only limited overview techniques, possible correlations may be missed.
 %%
 Explorative analyses and overview visualizations of the data set as presented by Klemm et al.
-\cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g.
-correlation between body size and spine shape).
-%%
-We incorporate the regression analysis, which is familiar to the domain experts into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypotheses.
+\cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
+%%
+We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
 %%
@@
 %%
-Difference views between regression formulas allow to assess the influences of individual variables in the process.
+Difference views between regression formulas allow to assess the influences of individual variables on the process.
 
@@
 	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
-	\item Analysis of confounding variables by providing comparison views between different formula results
-	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
+	\item The analysis of confounding variables by providing comparison views between different formula results.
+%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
+	\item The expansion and adaption of our method to any data with our open and web-based system.
 \end{itemize}
 %%
-The term Regression Cube was used before to define for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
-%%
-We refer to it using as visualization technique for comparing and analyzing a large number of different regression models with different input features.
+The term Regression Cube has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
+%%
+In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
 %%
@@
 %%
-Relative risks can be determined if a statistical resilient association of features towards a condition are extracted.
+Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
 %%
@@
 %%
-Statistical programs, such as \texttt{SPSS} are used to analyze the data using the classical sequential epidemiological workflow, which uses images largely only to communicate results, rather than providing insight.
+Statistical programs, such as \texttt{SPSS}, are used to analyze the data using the classical sequential epidemiological workflow.
+Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
 %%
@@
 The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
-These features are derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
+%%
+These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
 %%
@@
 %%
-This is however avoided if possible, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
+However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
 %%
@@
 
-\paragraph{Confounding Features}
+\paragraph{Confounding Features.}
 Features influencing the exposure as well as the outcome of an analysis are called confounders.
@@
 %%
-They are also the foundation of this work.
+The presented work is based on regression analysis as well.
 %%
@@
 %%
-$R^2$ values describe the goodness of fit; in other words how well the dependent features describe the target feature.
+$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
 %%
@@
 \subsection{The Study of Health in Pomerania (SHIP)}
-The \texttt{SHIP}, located in Northern Germany aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
 %%
@@
 %%
-Unique for the \texttt{SHIP} is the acquisition of medical image data per subject.
+Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
 %%
@@
 %%
-For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
+For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
 %%
@@
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-Tukey already stated in 1977 that data if too often only analyzed using confirmatory data analysis \cite{Tukey}.
+In 1977, Tukey already stated that data is too often only analyzed using confirmatory data analysis \cite{Tukey}.
 %%
@@
 %%
+\com{1. Can we reconstruct relationships based on the regression cube?}
+
+\com{2. How does the numeric target feature influence the result?}
 \subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
@@
 
-\textbf{Overview Visualizations gain importance with increasing feature count.}
+\textbf{Overview Visualizations are preferred over Black-box Methods.}
 %%

### 2015-03-26 draft.tex
minor changes on Evaluation
added first part of Application section with Ulis analysis

@@
 %%
-Based on this observations, we perform our analysis on three sub-populations, $PartitionF$, $PartitionM$ and \emph{$F:age\>52$}.
+Based on this observations, we perform our analysis on three sub-populations, \emph{females}, \emph{males} and \emph{females} older than 52 years.
 %%
@@
 %%
+The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
+%%
+Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
+%%
+The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
+%%
+UN therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$.
+%%
+The data set containing all \emph{males} showed broad consensus with the expected results.
+%%
+The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
+%%
+Using the mouse-slicing allowed for fast navigation to the respective features.
+%%
+They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
+%%
+Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
+
+%%
+UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
+%%
+Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
+%%
+
+For most features, UN found at least tendencies using our approach.
+%%
+Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
+
+%Regression cube calculations were carried out for
+
+\textbf{Analysis of non-discretized target feature.}
+The target feature derived from the MRI image data was discretized in \cite{Niemann2014} as decision trees algorithms require a categorical target feature.
+%%
+Since the \emph{regression cubes} analysis is not restricted in this way, UN applied the same analysis for \emph{males} and the two female group using its whole data range.
+%%
+
+Using the \emph{Regression Cube}, UN 
+\textbf{Interleukin-6 correlation with liver fat.}
+
+\com{ToDo: Reflect on the different methods - one is used to define subsets which define the target feature very well, the other looks how good the features describe the target itself}
 \com{1. Can we reconstruct relationships based on the regression cube?}
 
-\com{2. How does the numeric target feature influence the result?}
+\com{2. Does discretization of the numeric target feature influence the result?}
+
 \subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
@@
 %%
+
+\textbf{Attention steering is crucial.}
+\com{Bring important events to the users attention, otherwise it might be overlooked!}
 

### 2015-03-27 draft.tex
revised Hepatic Steatosis analysis section

--- 2015-03-26 draft.tex
@@
 %%
-KH is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
-
-The hepatic steatosis data set is analyzed by UN, a data scientists responsible for prior analysis of the data with.
-%%
-
-TI, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
+\emph{KH} is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
+
+The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data with.
+%%
+
+\emph{TI}, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
 %%
@@
 %%
+%%
 The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
@@
 %%
-\com{1. Can we reconstruct relationships based on the regression cube?}
-
-\com{2. How does the numeric target feature influence the result?}
+We refer to each analysis step regarding to their belonging in the VA-Mantra (Subsec.~\ref{sec:Workflow}).
+%%
+The analysis goal is reproducing results based \cite{Niemann2014}.
+%%
+Therefore, \emph{UN} started the \textbf{AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
+%%
+The \emph{SI} step using the 3D cube representations hotspots at the end of the cube (Fig.~\ref{fig:application} a).
+%%
+\textbf{ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
+%%
+Similar results present for analyzing the \emph{female} groups.
+%%
+\emph{UN} could reproduce most results, some features features exhibit lower correlations though, e.g., creatinine magnitudes.
+%%
+Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
+%%
+Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
+
+% %%
+% The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
+% %%
+% Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
+% %%
+% The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
+% %%
+% \emph{UN} therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subects.
+% %%
+% %%
+% %%
+% The data set containing all \emph{males} showed broad consensus with the expected results.
+% %%
+% The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
+% %%
+% Using the mouse-slicing allowed for fast navigation to the respective features.
+% %%
+% They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
+% %%
+% Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
+%
+% %%
+% UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
+% %%
+% Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
+% %%
+%
+% For most features, UN found at least tendencies using our approach.
+% %%
+% Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
+%
+% %Regression cube calculations were carried out for
+
+\textbf{Analysis of non-discretized target feature.}
+%%
+Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
+%%
+The 3D-cube representation already showed generally lower $R^2$ values, but since the analysis is now based on linear regression, the $R^2$ values can not be directly compared.
+%%
+Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of XX for somatometric features as opposed to XX).
+%%
+We conclude that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
+%%
+
+\textbf{Interleukin-6 correlation with liver fat.}
+%%
+During the analysis, one hotspot was always observed in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
+%%
+The correlation was high for both the dichotomized and continuous target feature.
+%%
+In the literature, relations between liver cancer \cite{He2013} and also chronic liver diseases \cite{Streetz2003}.
+%%
+Strong effects of \emph{IL-6} with hepatic steatosis were also prior described with mice \cite{Hong2004}.
+%%
+The finding is subject of further analysis.
+%%
 \subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}

### 2015-03-28 draft.tex
added new figures
added application image
revised figures

@@
 %%
-These data may be quantified using user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
+These data may be quantified based on user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
 %%
@@
 %%
-We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
+We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
 %%
@@
 %%
+They extend 2D scatter plots to display detailed information about a regression model.
+%%
 In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
 %%
-\com{Bernhard, does this suffice as reasoning?}
-
 \section{Epidemiological Background} \label{sec:Background}
@@
 %%
-The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
+The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
 %%
@@
 %%
-A second cohort, \texttt{SHIP-TREND} was introduced in 2012.
+A second cohort, \texttt{SHIP-TREND} was started in 2012.
 %%
@@
  \centering
- \includegraphics[width=3.0in]{figures/cube_sketch}
+ \includegraphics[width=3.5in]{figures/cube}
  \caption{
@@
  \centering
- \includegraphics[width=3.0in]{figures/workflow_sketch}
+ \includegraphics[width=3.0in]{figures/workflow}
  \caption{
@@
  \centering
- \includegraphics[width=3.5in]{figures/system_sketch}
+ \includegraphics[width=3.5in]{figures/system}
  \caption{
- %%
- \com{FIGURE NOT FINAL!}
  %%
@@
 
-\begin{figure}[htb]
- \centering
- \includegraphics[width=2.8in]{figures/placeholder}
- \caption{
- \com{Screenshots of Cubes from the Evaluation}
- }
-  \label{fig:comparison}
-\end{figure}
-
 \section{Implementation} \label{implementation}
@@
  \centering
- \includegraphics[width=2.8in]{figures/placeholder}
+ \includegraphics[width=3.5in]{figures/implementation}
  \caption{
@@
 \section{Application} \label{application}
+%%
+\begin{figure*}[htb]
+ \centering
+ \includegraphics[width=7.0in]{figures/application}
+ \caption{
+ \com{Screenshots of Cubes from the Evaluation}
+ }
+  \label{fig:Application}
+\end{figure*}
 %%

### 2015-03-28 draft.tex
revised using Bernhards notes

@@
 %%
-Their main goal is to pairwise assess the influence of independent features towards the target.
+Their main goal is to assess the influence of independent features towards the target pairwise.
 %%
@@
 
-Zhang et al. \cite{Zhang2014} is closest to ours regarding the application of visual analysis on cohort study data.
-%%
-Their work present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} and its manipulation using operations - \emph{views} and \emph{analytics}.
+Zhang et al. \cite{Zhang2014} is closest to our work regarding the application of visual analysis on cohort study data.
+%%
+They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} and its manipulation using operations - \emph{views} and \emph{analytics}.
 %%
@@
 %%
-Sub-groups are created automatically by applying \emph{analytics} and expert-guided using \emph{views}.
-%%
-Unfortunately, they do not go into the detail into the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
+Sub-groups are created automatically by applying \emph{analytics} and expert-guided \emph{views}.
+%%
+They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
 %%
@@
 
-Angelelli et al. \cite{Angelelli} is similar to ours as they visualize both image-derived an non image data using cube data structures with focus on comparison and knowledge extraction.
+Angelelli et al. \cite{Angelelli} visualize both image-derived an non image data using cube data structures with focus on comparison and knowledge extraction.
 %%
@@
 %%
-This allows for saving memory space.
-%%
-Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
+%This allows for saving memory space.
+%%
+%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
 %%
@@
 %%
-Futhermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
-%%
-We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
+Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
+%%
+%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
 
@@
 %%
-The focus remains on geolocated social media data rather than subjects with multiple dimensions.
-%%
-These and similar systems follow the same analysis work as we do to derive insight into data.
-
-Perer et al. divide cohort analysis in three major parts, present \emph{CAVA}, a system which allows
+%The focus remains on geolocated social media data rather than subjects with multiple dimensions.
+%%
+%These and similar systems follow the same analysis work as we do to derive insight into data.
 
@@
 %%
-Medical image data can be incorporated to investigate potentially interesting individuals.
+%Medical image data can be incorporated to investigate potentially interesting individuals.
 %%
@@
 %%
-The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for small epidemiological data sets.
 %%
@@
 
-\paragraph{Statistical based Analysis.}
-%%
-Statistical analysis are an inherent part of Visual Analysis.
+\paragraph{Statistical Analysis.}
+%%
+Statistical analysis is an inherent part of Visual Analysis.
 %%
@@
 %%
-Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
+Their research agenda acts as guideline for our design, comprising applications using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
 %%
@@
 %%
-In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
+%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
 %%
@@
 %% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
-Albuquerque et al. \cite{Albuquerque} present a interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
+Albuquerque et al. \cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
 %%
@@
 \paragraph{Prior Work.}
-In recent publications, we focused on analysis on characterization of the healthy aging process of the lumbar spine.
+In recent publications, we analyzed the healthy aging process of the lumbar spine.
 %%
@@
 
-We defined a Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (see \ref{sec:Background}) with a iterative analysis loop incorporating group selection and visualization using expert-input as well as clustering methods \cite{Klemm2014VIS}.
+We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (see \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert-input as well as clustering methods \cite{Klemm2014VIS}.
 %%
@@
 %%
-The popularity of the mosaic plot depicting \emph{Cram\'{e}r's V} contingency values with our domain experts was the inspiration for this work.
-%%
-The representation presented in \cite{Klemm2014VIS} only highlights pairwise correlation between features without a target disease.
+The popularity of the mosaic plot depicting \emph{Cram\'{e}r's V} contingency values among our domain experts was the inspiration for this work.
+%%
+The mosaic plot presented in \cite{Klemm2014VIS} only highlights pairwise correlation between features without a target disease.
 
 %% TODO: Check DLPD for correct page numbers
-While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
-%%
-These measures were analyzed using a novel decision tree quality plot towards non-image data.
+While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
+%%
+%These measures were analyzed using a novel decision tree quality plot towards non-image data.
 %%
@@
 \\\\
-The difference of the presented approach towards our previous and other herein presented work is twofold:
+The difference of the presented approach towards our previous and related work is twofold:
 \begin{enumerate}
@@
 %%
-These target features often indicates the presence of the investigated disease.
-%%
-As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
+These target features often indicate the presence of the investigated disease.
+%%
+As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
 %%
@@
 %%
-The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, because it uses the same notation, all operators can be used as before.
-%%
-This allows for a fast adaptation and in the epidemiological application domain.
+The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, because it uses the same notation. 
+%%
+%This allows for a fast adaptation and in the epidemiological application domain.
 %%
@@
 %%
-Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and contributes to presentiveness of the Regression Cube.
+Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{Regression Cube} representation.
 %%
@@
 %%
-When for example the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} behave similar towards it.
+When for example the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
 %%
@@
 \\\\
-The goal is to create an overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
+Our goal is to create an overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
 %%
@@
 %%
-To support \emph{hypothesis-free} analysis by providing a default formula:
+To support \emph{hypothesis-free} analysis we provide a default formula:
 
@@
 %%
-It is therefore suitable for a explorative analysis to give a general impression about relationships in the data set.
+It is therefore suitable for an explorative analysis to give a general impression about relationships in the data set.
 
@@
  %%
- The values of the current selected slice are mirrored and represented as orange data points on the slicing plane.
+ The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
  %%
@@
 %%
-More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
-%%
-Therefore, our prototype consists of three components:
+%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
+%%
+Our prototype consists of three components:
 \begin{itemize}
@@
 	%%
-	\item The \emph{formula editor} allows to adjust the formula towards a current hypothesis or conducting a \emph{hypothesis-free} analysis.
+	\item The \emph{formula editor} allows to adjust the formula w.r.t. a hypothesis or conducting a \emph{hypothesis-free} analysis.
 	%%
@@
 %%
-On entering the text are, a context modal displays all data set variables, as well as the available operators and their function.
+On formula input, a context modal displays all data set variables, as well as the available operators and their function.
 %%
@@
 %%
-The visualization and interaction with the \emph{Regression Cube} is the heart of the prototype.
-%%
-Results from the statistical processors are uploaded into the visualization by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.
+The visualization and interaction with the \emph{Regression Cube} is the prototype core.
+%%
+Results from the statistical processors are uploaded into the visualization slice by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.
 
@@
 %% INTERACTION
-Rotating with the cube is restricted to y-axis, preserving the mental map to position individual features.
+Rotating with the cube is restricted to the y-axis, preserving the mental map to position individual features.
 %%
@@
 %%
-Allowing more degrees of freedom was confusing to our users and also does not add value to the visualization.
+Allowing more degrees of freedom was confusing to our users and also did not add value to the visualization.
 %%
@@
 %%
-Using web technologies for prototyping allows for:
-\begin{itemize}
-	\item easy exchange of software, without the need of installing anything other than a modern web browser,
-	\item fast iterative adaption of method changes, without the need of installing updates,
-	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
-\end{itemize}
+% Using web technologies for prototyping allows for:
+% \begin{itemize}
+% 	\item easy exchange of software, without the need of installing anything other than a modern web browser,
+% 	\item fast iterative adaption of method changes, without the need of installing updates,
+% 	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
+% \end{itemize}
+% %%
+%%
+We rely on web-based technologies for our prototype.
 %%
@@
 %%
-The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is very popular in information visualization using vector graphics \cite{D3}.
+The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
 %%
@@
 %%
-They rely on the statistical programming language \texttt{R}\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}.
+They rely on the statistical programming language \texttt{R}.\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}
 %%
@@
 %%
-The source for the prototype is freely available at \texttt{Github}\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}.
+The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
 %%
@@
 
-Both data sets are unusual for epidemiological analysis regarding their feature extend.
+Both data sets are unusual for epidemiological analysis regarding their feature extent.
 %%
@@
 %%
+The herein used data sets comprise several hundred features.
+%%
 Our method focuses on data exploration and knowledge extraction and requires a wide scope of sociodemographic, medical and lifestyle features.
@@
 %%
-To assess the ability of a system to discover knowledge is hard to measure.
+To assess the ability of a system to discover knowledge is difficult to measure.
 %%
 Lam et al.
-\cite{Lam2012} propose for this purpose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
+\cite{Lam2012} propose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
 %%
@@
 %%
+\paragraph{Participants, Setup and Procedure.}
+%%
+%One major advantage of focusing on web-based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
+%%
+We conducted a web-based analysis by using an online-meeting software, which features voice chat as well as screen sharing.
+%%
+Starting an analysis using these techniques take about 5-10 minutes of setup time.
+
+The sessions started with an initial overview of the system, showcasing its features and functionality.
+%%
+After that the experts use the system on their own computers.
+%%
+The screen-sharing function was still used to observe the actions of the expert.
+%%
+All sessions were video recorded to be processed later on.
+%%
+
+%%
 We conducted a study with two participants, who also co-authored this publication.
 %%
-UN is a data scientist who analyzed the hepatic steatosis data set by extracting decision rules.
-%%
-KH is a radiologist (10 years of experience) with specialization in epidemiological research and is also responsible for the SHIP MRI acquisition.
-%%
-\paragraph{Participants, Setup and Procedure.}
-%%
-One major advantage of focusing on web-based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
-%%
-We conducted a web-based research by using an online-meeting software, which features voice chat as well as screen sharing.
-%%
-Starting an analysis using these techniques take about 5-10 minutes of setup time.
-
-The sessions started with an initial overview of the system, showcasing its features and functionality.
-%%
-After that the experts use the system on their own computers.
-%%
-The screen-sharing function was still used to observe the actions of the expert.
-%%
-All sessions were video recorded to be processed later on.
-%%
-
 Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.

### 2015-03-28 draft.tex
revised captions, filled gaps in Application section

@@
  \caption{
- (a) Overview visualization using a heatmap for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
+ (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
  %%
@@
  %%
- (b) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ heatmap visualizations.
- %%
- These represent the slices in our cube visualization.
- %%
- The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view, reducing the occlusion of other values.
+ (right) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ mosaic plot visualizations.
+ %%
+ These represent the slices in our cube.
+ %%
+ The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
  }
@@
 
-\subsection{Analysis Workflow}
+\subsection{Analysis Workflow} \label{sec:Workflow}
 %%
@@
  %%
- [\textbf{1.AF}] The analysis starts with a set of features given by the data set.
- %%
- Then, the user may specify a formula regarding to specify a hypothesis, or use a predefined formula to start a explorative analysis of the data set.
- %%
- [\textbf{2.SI}] The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
- %%
- [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula.
- %%
- By applying a difference view, regression cubes can be compared to each other.
- %%
- [\textbf{4.DD}] Details about features using the 2D-heatmap representation yields insights and hypothesis about feature relations.
+ % The analysis starts with a set of features given by the data set.
+ %%
+ [\textbf{1.AF}] The workflow starts by declaring formula regarding to specify a hypothesis, or use a predefined formula for a hypothesis-free analysis.
+ %%
+ [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
+ %%
+ The user has the option to either adjust the formula or to derive details on demand on a specific regression.
+ %%
+ [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or selecting a difference view.
+ %%
+ The latter is used to compare regression cubes.
+ %%
+ [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yields insights and hypothesis about feature relations.
  }
@@
  %%
- (c) 2D mosaic plot visualization of the selected slice.
- %%
- \com{TODO: Adapt formula to final image.}
- %%
- Context information is shown for the regression model of $Parenchym \sim Bronchitis + AlcoholLast30Days$, depicting details about the confidence intervals, coefficients and associated p-values.
+ (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
+ %%
+ %\com{TODO: Adapt formula to final image.}
+ %%
+ %Context information is shown for the regression model of $Parenchym \sim Bronchitis + AlcoholLast30Days$, depicting details about the confidence intervals, coefficients and associated p-values.
  %%
@@
 \paragraph{3D-Prism as Data Mini-Map.}
+\com{Add section about difference view for static target features.}
 %%
@@
  \caption{
- \com{Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
+ The front-end (left) is written in \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and uses different Javascript libraries, such as \emph{Angularjs}, \emph{Three.js} and D3js.
+ %%
+ The web-server back-end (right) is written in \emph{Nodejs}, hosting is carried out using \emph{Heroku}.
+ %%
+ \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{Regression Cubes}.
+ %%
+ Additional statistical back-ends can be attached to the system to decrease the computation time.
+% The number of attached machines to the system 
+%    the  the  Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
  }
@@
  \caption{
- \com{Screenshots of Cubes from the Evaluation}
+ The analysis of the numerical and dichotomized target feature depicting liver-fat values yields similar results (left).
+ %%
+ In (a), hotspots for somatometrical features with high correlations were found.
+ %%
+ High correlations were also found for features depicting \emph{hepatic steatosis} (b).
+ %%
+ A high correlations between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was found during the analysis using the 2D mosaic plot.
+ %%
+ \com{Caption Breast Fat Data}
  }
@@
 \subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
-%%
-We refer to each analysis step regarding to their belonging in the VA-Mantra (Subsec.~\ref{sec:Workflow}).
-%%
-The analysis goal is reproducing results based \cite{Niemann2014}.
+\com{Add R$^2$ values}
+%%
+We refer to each analysis step regarding to their belonging in the VA-Mantra (Sec.~\ref{sec:Workflow}).
+%%
+The analysis goal is reproducing results based on \cite{Niemann2014}.
 %%
@@
 %%
-The \emph{SI} step using the 3D cube representations hotspots at the end of the cube (Fig.~\ref{fig:application} a).
+The \emph{SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
 %%
@@
 %%
+Our method was well received among the domain experts.
+%%
+The novel way of modeling expert knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
+%%
+%TI commented on the
+
 \textbf{Extracted Hypotheses have to investigated further.}
-\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
+%%
+Agreeing with TI's feedback, each finding and hypotheses has to be confirmed using dedicated statistical analysis.
+%%
+This incorporated investigating all potential confounding features as well as looking for outlier subjects.
+%%
+TI commented on the possibility of adding even more regression types to model different correlation types to customize the cube even more.
+%%
+%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality of fit measures, such as the $R^2$.
+%%
+%\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
 %%
@@
 %%
-\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
+Discretization reduces the information space and introduces bias into the data and is therefore avoided in epidemiological research when possible.
+%%
+In contrast to many data mining algorithms, our method allows to use the concurrent analysis of heterogeneous data types.
+%%
+Investigations of the the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results differing in detail.%, but differences  %differing in detail.
+%%
+The overall explanatory power towards the numerical feature was lower.
+%Our investigation regarding the hepatic steatosis data set confirms differences in results
+%%
+%\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
 %%
@@
 \textbf{Attention steering is crucial.}
-\com{Bring important events to the users attention, otherwise it might be overlooked!}
-
-\textbf{Support for different subject groups.}
-\com{Include this?}
-%%
-\com{Feedback from the evaluation goes here.}
-\com{Time-aspect critical, interactive analysis requires for fast response. The method needs to be speeded up.}
+%%
+Important events have to be highlighted in overview visualizations to steer the users attention to interesting parts of the data.
+%%
+Poor guidance lead to potentially overlooked relationships.
+%%
+We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose.
+%%
+Foe example, formulas with a static target feature (e.g.
+$Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z ray.
+%%
+
+%This allows for fast parsing through the data
+
+%In order to benefit most from the human pattern recognition, interesting events have to highlighted in order to steer the users attention to important parts of the data.
+%Using overview visualizations is associated with important 
+%\com{Bring important events to the users attention, otherwise it might be overlooked!}
+
+%\textbf{Support for different subject groups.}
+%\com{Include this?}
+%%
+%\com{Feedback from the evaluation goes here.}
+%\com{Time-aspect critical, interactive analysis requires for fast response. The method needs to be speeded up.}
 

### 2015-03-29 draft.tex
revised using Sylvias notes; shortened related work

@@
 %%
-%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
+With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract
@@
 %%
-Furthermore, they can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
-%%
-During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.
+Furthermore, the insights can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
+%%
+During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then statistically analyzed.
 
@@
 %%
-Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
+%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
+Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and required subject count.
 %%
@@
 %%
-An alternative data-driven approach is described as follows.
-
-\paragraph{Data Driven Hypothesis-Generation-based Analysis.}
-We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
-%%
-Their methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
+
+In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step.
+%%
+The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
 %%
 Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
+%%
+In this work, we focus on a similar approach to deriving insight and even new hypotheses through the data rather than using it just for confirmatory analysis.
+%An alternative data-driven approach is described as follows.
+%%
+% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
+% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
+% %%
+% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
+% %%
+% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
 
@@
 %%
-\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
-%%
-They yield a highly heterogenous and incomplete information space.
+There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
+%%
+%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
+We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
 %%
@@
 %\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
-\\\\
+
 %\paragraph{Data Types.}
@@
 %%
-More discussion of data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
-
+More discussion of cohort study data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
+%%
 %\paragraph{Image Data.}
@@
 % Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
-
-\paragraph{Confounding Features.}
-Features influencing the exposure as well as the outcome of an analysis are called confounders.
-%%
-The analysis model has to be adjusted by normalizing all included features towards the confounding feature.
+\\\\
+%\paragraph{Confounding Features.}
+Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
+%Confounding features have to be considered in epidemiological analysis, as they influence both the exposure as well as the outcome of an analysis.
+%%
+%Features influencing the exposure as well as the outcome of an analysis are called confounders.
+%%
+The analysis model has to be adjusted by normalizing all included features towards the confounder.
 %%
@@
 %%
+A example of a regression formula would be $Kidney\_Disorder ~ Smoking + Obesity$.
 The most used regression operators comprise:
@@
 \begin{itemize}
-	\item \fbox{$+,-$} inclusion/exclusion of the variable,
+	\item $+,-$ inclusion/exclusion of the variable,
 	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
@@
 The type of the target feature restricts the regression type.
-
+%%
+Different regression types are available, we focus on the following:
+%%
 \paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
@@
 %%
-Logistic models as opposed to linear models do not allow for extracting an $R^2$ goodness of fit value.
+Logistic models as opposed to linear models do not allow for extracting an $R^2$ quality of fit value.
 %%
@@
 %%
-We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
+%We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
 %%
@@
 %%
-Comparison of models is also included by mapping model values on line-style (e.g. solid for $model_a$ and dashed for $model_b$).
-%%
-This limits the method to comparing a few models at once.
-%%
-The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
+Comparison of models is also included by mapping model values on line-style. %(e.g. solid for $model_a$ and dashed for $model_b$).
+%%
+This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
+%%
+%The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
 %%
@@
 %%
-They also focus on metric features, while we process also categorical data.
-%%
-We share the matrix-arrangement of results depicting relationships between independent features in a model.
+They also focus on metric features. %, while we process also categorical data.
+%%
+%We share the matrix-arrangement of results depicting relationships between independent features in a model.
 
@@
 %%
-They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
-%%
-Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
-%%
-We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}, meaning that the system has to adapt to different requirements in different analysis steps.
-%%
-We implement this by supporting both hypothesis-based and hypothesis-free analysis.
+%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
+%%
+%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
+%%
+We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
+%%
+%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
 
@@
 %%
-They use \emph{Pearsson's r} to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
-%%
-We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
+They use \emph{Pearsson's} $r$ to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
+%%
+%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
 %%
@@
 Shneiderman et al.
-\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data, e.g.
+\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data.
+%, e.g.
 using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
 %%
-They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
-%%
-We focus on the latter.
+%They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
+%%
+%We focus on the latter.
 %%
 %Shneiderman and colleagues also present challenges, 
-The presented challenges in this application comprise custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
+The presented challenges for \emph{public health information visualization} comprising custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
 %%
@@
 %We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
-
+%%
 %%
@@
 %%
-We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
-
+%We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
+%%
 %%
@@
 %%
-We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
-
+%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
 %%
@@
 %%
-The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for small epidemiological data sets.
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for \emph{small} epidemiological data sets.
 %%
 Dai et al.
-\cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
+\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
 %%
@@
 %%
-We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
-%%
-
+%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
+%%
 \paragraph{Statistical Analysis.}
 %%
-Statistical analysis is an inherent part of Visual Analysis.
-%%
-Here we present work showing a strong focus on statistical analysis in this context.
+%Statistical analysis is an inherent part of Visual Analysis.
+%%
+%Here we present work showing a strong focus on statistical analysis in this context.
 %% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
@@
 %%
-The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.
+%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.
 
@@
 %%
-These subsets are then analyzed in detail, following a drilling-down approach.
-%%
-They incorporate scatter plot matrices (\texttt{SPLOM´S}) of quality metrics describing the features.
-%%
-We incorporate the idea of abstracting features for comparison.
+%These subsets are then analyzed in detail, following a drilling-down approach.
+%%
+They analyzing subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM´S}) of quality metrics. %describing the features.
+%%
+%We incorporate the idea of abstracting features for comparison.
 %%
@@
 %%
-Therefore, we utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
-%%
-The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
+We utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
+%%
+%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
 %%
@@
 The mosaic plot presented in \cite{Klemm2014VIS} only highlights pairwise correlation between features without a target disease.
-
+%%
 %% TODO: Check DLPD for correct page numbers
@@
  %%
- The $R^2$ values extracted from the regression formulas depict the goodness of fit and are mapped to color saturation.
+ The $R^2$ values extracted from the regression formulas depict the quality of fit and are mapped to color saturation.
  %%
@@
 %%
-For the \emph{Regression Cube} analysis we are interested in the goodness of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
+For the \emph{Regression Cube} analysis we are interested in the quality of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
 %%

### 2015-03-29 draft.tex

@@
 %%
-A qualitative analysis with three domain experts was conducted on a breast fat data set.
+A qualitative analysis was conducted on a breast fat data set.
 %%
@@
 %%
-With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
+With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract
@@
 %Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
-Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and required subject count.
+Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
 %%
@@
 %%
-Explorative analyses and overview visualizations of the data set as presented by Klemm et al.
-\cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
+Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
 %%
@@
 %	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
-	\item The expansion and adaption of our method to any data with our open and web-based system.
+	\item The expansion and adaption of our method to any data with our open web-based system.
 \end{itemize}
 %%
-The term Regression Cube has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
+The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
 %%
@@
 %%
-He emphasized the need to use data to derive hypotheses, which can then be tested again.
+He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
 %%
@@
 %%
+% der Satz kann raus. 1. Teil ist wirklich "common sense", 2. Teil wird im n√§chsten Satz mit "...closest to ours" nochmal genannt.
 %We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
@@
 \\\\
-The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
+The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} are closest related to ours regarding different aspects.
 %% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
@@
 %%
-Their main goal is to assess the influence of independent features towards the target pairwise.
-%%
-This is achieved by using a plot matrix displaying models as contours towards the target feature.
-%%
-They also incorporate 3D visualizations for each pairwise combination, but mainly because it is popular with the engineers in their target domains.
-%%
-Linked views of model deviations allow to select outlier residuals.
-%%
-Comparison of models is also included by mapping model values on line-style. %(e.g. solid for $model_a$ and dashed for $model_b$).
+Their main goal is to assess the pairwise influence of independent features towards the target feature using a plot matrix displaying models as contours.
+%%
+They also incorporate 3D visualizations for each pairwise combination, but mainly because its popularity with their target domain engineers.
+%%
+Linked views of model deviations allow to select outliers.
 %%
@@
 %%
-%The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
-%%
-The main difference is their focus on analyzing one complex model in detail, yielding in extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
-%%
-They also focus on metric features. %, while we process also categorical data.
+The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
+%%
+They also focus on metric features, while we process categorical data as well.
 %%
@@
 %% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
-We use the term Regression Cube to refer to a comparative 3D representation of multiple regression models.
-%%
-Chan et al. \cite{Chan} use the same term to describe an extension of the 2D-Scatterplot representation of a linear regression models (incorporating solely metric features) to a 3D-Cube.
+We use the term Regression Cube for a comparative 3D representation of multiple regression models.
+%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
+Chan et al. \cite{Chan} use the same term to describe an extension of the 2D Scatterplot representation of a linear regression models (incorporating solely metric features) to a 3D Cube.
 %%
@@
 %%
-\com{Does this suffice as reason on why we use the same term "Regression Cube" for different things?}
-
-Zhang et al. \cite{Zhang2014} is closest to our work regarding the application of visual analysis on cohort study data.
-%%
-They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} and its manipulation using operations - \emph{views} and \emph{analytics}.
-%%
-They use the system to find longitudinal pathways for diseases using subject health records.
-%%
-Sub-groups are created automatically by applying \emph{analytics} and expert-guided \emph{views}.
+
+Zhang et al. \cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
+%%
+They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
+%%
+They use the system to find longitudinal pathways for diseases using health records.
+%%
+Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
 %%
@@
 
-Angelelli et al. \cite{Angelelli} visualize both image-derived an non image data using cube data structures with focus on comparison and knowledge extraction.
-%%
-Their proposed engine represents multiple heterogenous cohort study data sets as multiple normalized data cubes to eliminate data redundancy between them and perform runtime measure aggregation when measures of the different cubes have to be combined.
+Angelelli et al. \cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
+%%
+Their proposed engine represents multiple heterogenous cohort study data sets as multiple normalized data cubes.
+%%
+Thus, data redundancy can be avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
 %%
@@
 %%
-We employ the idea of fast conducting large scale correlation analysis between many features to derive insight into the data.
+Also, we aim at a fast large scale correlation analysis between many features to derive insight into the data.
 
@@
 %Shneiderman and colleagues also present challenges, 
-The presented challenges for \emph{public health information visualization} comprising custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
+The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
 %%
@@
 %%
-In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
-%%
-They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
+
+%%
+In their survey on interactive information visualizations for health record data, Rind et al. \cite{Rind} identified future work addressing the usability of such frameworks.
 %%
@@
 %%
+We follow their recommendation to design an open system available to everyone and applicable to various data sets.
 %%
@@
 %%
-They derive hypotheses using their high variation average message density algorithms, steering the users focus on particular events.
+They derive hypotheses using their high variation average message density algorithms, steering the user's focus to particular events.
 %%
@@
 %%
-They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select data.
+They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select the data.
 %%
@@
 %%
-Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+Generalized Pairs plots (\texttt{GPLOM√Ç¬¥S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
 %%
@@
 %%
-A \emph{concept map} summarized features related to a specified disease.
+A \emph{concept map} summarizes features related to a specified disease.
 %%
@@
 %% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
-Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
-%%
-Their research agenda acts as guideline for our design, comprising applications using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
+Bertini et al. \cite{Bertini} present an overview over quality metrics describing high dimensional data.
+%%
+Their research agenda acts as guideline for our design, comprising applications using domain experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
 %%
 %% TODO: Use this one?
-"Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich ändert, wenn man per Zufall die Populationen verändert, um zu schauen wie zuverlässig diese Statistik ist.
+"Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich √É¬§ndert, wenn man per Zufall die Populationen ver√É¬§ndert, um zu schauen wie zuverl√É¬§ssig diese Statistik ist.
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
-Ahmadi et al.
-\cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described using by their most reliable linear regression model.
+Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described by their most reliable linear regression model.
 %%
@@
 %%
-They focus on presenting a algebraic for efficient regression-model calculation to find the best fit for a subspace.
+They focus on presenting an algebraic for efficient regression-model calculation to find the best fit for a subspace.
 %%
@@
 %%
-They analyzing subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM´S}) of quality metrics. %describing the features.
+They analyzing subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM¬´S}) of quality metrics. %describing the features.
 %%
@@
 %% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
-Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with a interactive data mining tools.
-%%
-They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
+Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with interactive data mining tools.
+%%
+They extracted association rules that serve as basis for a proof-of-concept evaluation.
 %%
@@
 %%
-%\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
+%\com{Gla√É¬üer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
 
 \paragraph{Prior Work.}
-In recent publications, we analyzed the healthy aging process of the lumbar spine.
+In our prior work, we analyzed the healthy aging process of the lumbar spine.
 %%
@@
 
-We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (see \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert-input as well as clustering methods \cite{Klemm2014VIS}.
-%%
-Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image-and non-image information.
-%%
-The popularity of the mosaic plot depicting \emph{Cram\'{e}r's V} contingency values among our domain experts was the inspiration for this work.
-%%
-The mosaic plot presented in \cite{Klemm2014VIS} only highlights pairwise correlation between features without a target disease.
-%%
+We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (recall \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods \cite{Klemm2014VIS}.
+%%
+Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image and non-image information.
+%% 
+% erst die Erkl√§rung, dann die Folge daraus =)
+The mosaic plot presented in \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
+%%
+Its great popularity among our domain experts was the inspiration for this work.
+%%
+
 %% TODO: Check DLPD for correct page numbers
-While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
+While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
 %%
@@
 %%
-We concluded that the model quality is not enough to characterize back pain.
+We concluded that the model quality is insufficient to characterize back pain.
 \\\\
@@
 %%
-A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
-%%
-\subsection{Cube Description using Regression Formula Notation}
-%%
-Expert knowledge is introduced into regression analyses using the regression formulas.
+% unified -> singular [Steffi?]
+A regression model is based on expert knowledge, there is no singular rule how to apply them to a given set of features. Thus, they have to be applied with care.
+%%
+% Verben in ‹berschriften immer groﬂ
+\subsection{Cube Description Using Regression Formula Notation}
+%%
+Expert knowledge is integrated into regression analysis by the regression formulas.
 %%
@@
 
-Since we want to use the regression analyses with an overview visualization, we are are interested in all possible combinations of (two or more) independent features describing a target.
+Since we want to associate the regression analyses with an overview visualization, we are interested in all possible combinations of (two or more) independent features describing a target.
 %%
@@
 %%
-In a data set with $n$ (e.g. 100) features, the regression formula
+In a data set with $n$ (e.g., 100) features, the regression formula
 \begin{equation}
@@
 \end{equation}
-would yield $n^2$ (10.000) regression models, describing all possible combinations of two features in the data describing the target $Cancer$.
-%%
-The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, because it uses the same notation. 
+would yield $n^2$ (10,000) regression models, describing all possible combinations of two features in the data describing the target feature  $Cancer$.
+%%
+The major advantage of this notation is that it comes natural to anyone familiar with regression analysis since it is the standard way of expression 
 %%
@@
 %%
-The number of calculated regression models increases exponentially for each dynamic variable added.
-%%
-If we assume a data set with 100 features with the formula $Z \sim X + Y$ we calculate 1,000,000 regression models.
-%%
-When each regression takes about 50~ms calculation, we have to wait roughly 14 hours for the calculation to complete.
+The number of calculated regression models exponentially increases for each dynamic variable added.
+%%
+If we assume a data set with 100 features and the formula $Z \sim X + Y$, we obtain 1,000,000 regression models.
+%%
+% hours -> h da ms ja auch benutzt wird.
+When each regression takes about 50~ms calculation, we have to roughly wait 14~h for the calculation to complete.
 %%
@@
 
-\subsection{Target-Variable-dependent Dimension Reduction}
+% -Dependent & -Based in ‹berschriften groﬂ, sicherheitshalber nochmal Steffi fragen
+% oder: Dimension Reduction Depending on the Target Variable
+\subsection{Target-Variable-Dependent Dimension Reduction}
 % \com{Uli, Can you please proofread this?}
@@
 % %%
-In epidemiological studies, manifold recordings lead to an abundance of features.
-%%
-Generally many of them exhibit low or none correlation towards the target feature.
+In epidemiological studies, manifold recordings lead to an abundance of features and thus a high-dimensional feature space.
+%%
+In general, many of them exhibit low or none correlation at all towards the target feature.
 %%
@@
 %%
-The Correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in F.
+% nochmal checken wo $F$ eingef¸hrt wird, hier oder im ¸bern‰chsten Satz?
+The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
 %%
@@
 %%
-Beginning with an empty set of features $F$ the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when adding any feature does not increase the merit.
-%%
-When for example the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
-%%
-They, however, correlate strongly with each other.
-%%
-The CFS algorithm would then select the feature which has the largest explanatory power and discards the other dimensions.
+% Ggf. wurde $F$ schon oben eingef¸hrt, TODO
+Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.
+%%
+For example, if the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
+%%
+However, they strongly correlate with each other.
+%%
+The CFS algorithm will select the feature which has the largest explanatory power and discards the other dimensions.
 
@@
 %%
-In our epidemiological data we observed a usual number of 10 to 20 features.
-%%
-For features, such as age or gender, which affect most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}) the number is significantly larger (about half the features in the whole data set).
-%%
-The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
-%%
-\com{Uli, is this correct this way?}
-%%
-
-Now we have a method to derive the interesting regression models in a reasonable time span.
+In our epidemiological data, we usually observed a number of 10 to 30 features.
+%For features, which affect the most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}), e.g., age or gender,  the number is significantly larger (about half the features in the whole data set).
+%%
+The number of selected features using the CFS algorithm reflects their information entropy on the target.
+%%
+A large list of features is an expression of low correlation to the target feature.
+%%
+%The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
+%%
+
+With this method, we are able to derive the interesting regression models in a reasonable time span.
 %%
@@
 
-\subsection{Abstracting Regression Results using $R^2$}
+\subsection{Abstracting Regression Results Using $R^2$}
 \begin{figure}[htb]
@@
  \caption{
- (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes feature $Age$.
- %%
- The $R^2$ values extracted from the regression formulas depict the quality of fit and are mapped to color saturation.
- %%
- A saturated color shows a strong correlation.
- %%
- (right) Since $Z$ assumes all features $n$ as given by the formula, it yields $n$ mosaic plot visualizations.
+ (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $Age$.
+ %%
+ The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
+ %%
+ (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
  %%
@@
 %% TODO: Cite another paper here!
-As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded using color on a mosaic plot.
+As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
 %%
@@
 %%
-For the \emph{Regression Cube} analysis we are interested in the quality of fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
+For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
 %%
@@
 %%
-This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g. $Age \sim X + Y$).
+This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
 %%
@@
 %%
-The visualization of $R^2$ values derived from different regression cubes (e.g. $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
-%%
-Therefore, the $R^2$ results of different regression methods are encoded using different colors (e.g. blue for linear regression and red for logistic regression).
-%%
-This way, the cube can easily be extended using other regression types.
-%%
-For cubes having one fixed target feature, such as $Cancer \sim X + Y + Z$ no such encodings are required and the $z$ dimension can be compared directly.
+The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%%
+% e.g. wenn man es w‰hlen kann
+% i.e. wenn es so ist
+Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
+%%
+% Adverb immer vors Hauptverb oder Satzende
+Thus, the cube can be easily extended using other regression types.
+%%
+For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
 %%
 \\\\
-Our goal is to create an overview visualization for a data set, but on the other hand we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
+Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
 %%
@@
 %%
-Different analysis approaches require different starting points using the \emph{Regression Cube}.
+The different analysis approaches require different starting points using the \emph{Regression Cube}.
 
@@
  \caption{
- Workflow of the analysis using \emph{Regression Cubes}.
+ Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
  %%
@@
  %%
- [\textbf{1.AF}] The workflow starts by declaring formula regarding to specify a hypothesis, or use a predefined formula for a hypothesis-free analysis.
+ [\textbf{1.AF}] The workflow starts by declaring a formula regarding to specify a hypothesis, or use a predefined formula for a hypothesis-free analysis.
  %%
@@
 %%
-The analysis workflow follows the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:
+% ‹bergang ist recht holprig gewesen
+Our Regression Cube models is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:
 
@@
 %%
-Observations in the 2D-mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
+Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
 %%
@@
 %%
-Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
-%%
-To support \emph{hypothesis-free} analysis we provide a default formula:
+Using the operators, the dynamic variables and the dataset features, many different assumptions can be expressed.
+%%
+To support \emph{hypothesis-free} analysis, we provide a default formula:
 
@@
 %%
-This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
+This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which features are of interest.
 %%
@@
 
-Hypotheses about the data are reflected in either how dynamic variables are related using the regression operators.
-%%
-Also, static features can be added for each regression formula.
+% warum war da ein "either", sollte der Satz etwas anderes ausdr¸cken?
+Hypotheses about the data are easily built up by relating dynamic variables with the regression operators.
+%%
+Furthermore, static features can be added for each regression formula.
 %%
@@
 %%
-$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
+$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
 %%
@@
 %%
-This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result.
+This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combination influence the result.
 %%
@@
 %%
-One cube (formula) acts as reference.
+One cube (extracted with one formula) acts as reference.
 %%
@@
 %%
-It can for example be utilized for comparing the influence of one feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
+For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
 %%
@@
 \end{figure}
-%%
-We designed our system with openness and ease of use in mind.
-%%
-Using open formats as input interfaces does not restrict the application towards non-epidemiological data sets.
+%% scientific english: kurz und langweilig ;)
+We designed our system to be freely accessible and easy to use.
+With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
+%%
 %%
@@
 \paragraph{3D-Prism as Data Mini-Map.}
-\com{Add section about difference view for static target features.}
 %%
@@
 %%
-It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
+%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
 %%
 The displayed prism shows values above the matrix diagonal.
+%%
+For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System} b).
+%%
+Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
+%%
+For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
+%%
+This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
 %%
@@
 \paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
-\com{Add Figure reference!}
-%%
-The 2D mosaic visualization shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
+%%
+The 2D mosaic visualization (Fig.~\ref{fig:System} c) shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
 %%
@@
 %%
-Domain expert for the breast fat data set analysis is KH, a clinician (10 years of experience) with focus on epidemiological research.
+Domain expert for the breast fat data set analysis is \emph{KH}, a clinician (10 years of experience) with focus on epidemiological research.
 %%
@@
 
-\emph{TI}, a statistician with focus on epidemiology (\com{TODO} years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
+\emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
+%%
+\subsection{The Hepatic Steatosis Data Set}
+We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
+%%
+The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
+%%
+Liver fat concentrations of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
+%%
+The data set contains labels for 578 participants for which the MRI data was available to that point.
+%%
+The MRI scans for each subject were introduced in SHIP-2, so the target feature is present for just one of three study waves.
+
+%%
+Apart from the target feature, the data set contains 199 features, comprising socio-demographic features (gender, age), consumption behavior (e.g.
+alcohol or tobacco), laboratory data (e.g.
+sera concentrations), and two features depicting the liver ultrasound.
+%%
+%The acquisition wave is denoted using the appendix; 85 for S0 (SHIP-0), 55 for S1, 55 for S2 and 10 time-independent SNPs (DNA base pairs).
+%%
+The acquisition wave is denoted using the appendix; \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features for \emph{s1} and 55 for \emph{s2}, alongside with 10 time-independent SNPs (DNA base pairs).
+%%
+
+In \cite{Niemann2014}, the authors show different class distribution of liver fat concentrations of women and men.
+%%
+For women a association between age and liver fat was identified.
+%Also, for women they identify that there is a association between age and liver fat concentration.
+%%
+An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
+%%
+Based on these observations, we perform our analysis on three sub-populations: \emph{females}, \emph{males} and \emph{females} older than 52 years.
 %%
@@
 
-The data comprises 1.186 subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
-%%
-It contains 231 dimensions, holding information about
-\begin{itemize}
-	\item \emph{somatometric features}, e.g., body size, weight or BMI, 
-	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume, 
-	\item \emph{personal history}, e.g., occupation, marital status,
-	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
-	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
-	\item \emph{mammography features} derived from the image data.
-\end{itemize}
-%%
-The latter were derived using a level set based segmentation \cite{Ivanovska2014} of MRI image data for each subject.
-%%
-\com{Check this with Mrs. Hegenscheid, probably most of them were segmented by hand!}
+The data comprises 1.186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
+%%
+It contains 231 dimensions, holding information about somatometric features (e.g.
+body size or weight) consumption behavior (e.g.
+alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women specific features (e.g., number of born children or contraception type) as well as mammography features.
+% It contains 231 dimensions, holding information about
+% \begin{itemize}
+% 	\item \emph{somatometric features}, e.g., body size, weight or BMI,
+% 	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume,
+% 	\item \emph{personal history}, e.g., occupation, marital status,
+% 	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
+% 	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
+% 	\item \emph{mammography features} derived from the image data.
+% \end{itemize}
+%%
+The latter were derived using manual segmentation of MRI image data for each subject carried out by radiologists.
+%%
 %%
@@
 %%
-This avoids the need of defining a separate data dictionary file for the \emph{Regression Cube}, translating the feature names.
-%%
-All male subjects were removed as their data does not contribute to the analysis.
-%%
-\subsection{The Hepatic Steatosis Data Set}
-We employ a subset of the data set which was used in \cite{Niemann2014} to identify predictive features with respect to the reversible disorder hepatic steatosis (fatty liver).
-%%
-The binary target feature is derived from the liver fat concentration which was measured with Magnetic Resonance Imaging (MRI).
-%%
-In particular, participants with a liver fat concentration of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
-%%
-The data set contains labels for 578 participants for which the MRI data was available to that point.
-%%
-Note that MRI was performed only in SHIP-2, so the target feature is present for just one of three study waves.
-
-%%
-Apart from the target feature, the data set contains 199 features extracted from participants' questionnaire answers and medical tests.
-%%
-They are features on socio-demographics (sex, age, etc.), features on consumption behavior (e.g.
-alcohol and cigarettes), SNPs, features extracted from laboratory data (e.g.
-sera concentrations), and two features on the results of the liver ultrasound.
-%%
-The names of 85 features have the suffix \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features exhibit the suffix \emph{s1} and 55 features have the suffix \emph{s2}, alongside 10 time-independent SNPs.
-%%
-
-In \cite{Niemann2014}, the authors show that women and men exhibit differences in the class distribution of the liver fat concentration.
-%%
-Also, for women they identify that there is a association between age and liver fat concentration.
-%%
-They compute an appropriate cut-off value of 52 years for which the class distribution is most homogeneous within the resulting subsets.
-%%
-Based on this observations, we perform our analysis on three sub-populations, \emph{females}, \emph{males} and \emph{females} older than 52 years.
-%%
-\subsection{Case 1: Hypothesis-free Analysis of the Breast Cancer Data Set}
-%%
-%%
-The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
-%%
-Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
-%%
-
-\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
-\com{Add R$^2$ values}
+This avoids the need of defining a separate data dictionary file for translating the feature names.
+%%
+%All male subjects were removed as their data does not contribute to the analysis.
+%%
+\subsection{Case 1: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
 %%
@@
 %%
-Therefore, \emph{UN} started the \textbf{AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
-%%
-The \emph{SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
-%%
-\textbf{ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
+Therefore, \emph{UN} started the \textbf{1.AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
+%%
+The \textbf{2.SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
+%%
+\textbf{3.ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{4.DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (used renal retention parameter) and uric acid (among others used as gout and diabetes risk factors) magnitudes.
 %%
@@
 %%
+A slight influence of age on the target feature could be observed for women ($R^2$ of $0.09$ for females compared to $0.02$ for males).
+%%
 Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
@@
 Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
-
+%%
 % %%
@@
 %%
-Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of XX for somatometric features as opposed to XX).
+Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
 %%
@@
 %%
+\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
+%%
+The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
+%%
+Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
+%%
 \subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
@@
 %%
-The novel way of modeling expert knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
+For the first time, they were able to derive a overview visualization custom tailored to underlying assumptions.
+%%
+In this context, \emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
+%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
 %%
@@
 
-\textbf{Extracted Hypotheses have to investigated further.}
-%%
-Agreeing with TI's feedback, each finding and hypotheses has to be confirmed using dedicated statistical analysis.
-%%
-This incorporated investigating all potential confounding features as well as looking for outlier subjects.
-%%
-TI commented on the possibility of adding even more regression types to model different correlation types to customize the cube even more.
-%%
-%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality of fit measures, such as the $R^2$.
+\textbf{Extracted Hypotheses have to be investigated further.}
+%%
+Agreeing with \emph{TI}'s feedback, each finding and hypothesis has to be confirmed using a dedicated statistical analysis.
+%%
+A accompanying search for correlations potentially highlighting confounders can be carried out using our method.
+%%
+%This incorporates detailed investigation of potential confounder features as well as looking for outlier subjects.
+%%
+\emph{TI} commented on the possibility of adding more regression types to model different correlation types.% to customize the cube even more.
+%%
+%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality-of-fit measures, such as the $R^2$.
 %%
@@
 
-\textbf{Overview Visualizations are preferred over Black-box Methods.}
-%%
-While the classical epidemiological workflow (see Sec.~\ref{sec:Background}) represents the foundation of the research in this area, explorative analysis based on the data gains importance.
-%%
-Incorporating data mining algorithms with statistically established and overview visualizations allow for reasoning about risk factors using underlying epidemiological data sets.
-%%
-One major advantage we observed is the confidence of an expert in the results he or she actually observe.
-%%
-Their participation in the analysis using human pattern-detection and expert knowledge is for experts preferred over results derived using a black-box method.
-%%
-By observing expected correlations matching the expert knowledge strengthens the confidence in the method and subsequently in the hypothesis generated from unexpected relationships.
+\textbf{Overview Visualizations are Preferred over Black-box Methods.}
+%%
+Explorative analysis based on the data gains importance in epidemiology as the data set complexity increases.
+%While the classical epidemiological workflow represents the foundation of the research in this area, explorative analysis based on the data gains importance.
+Results from automatic `black-box' methods, such as data mining algorithm are more often obscure to the expert.
+%%
+Findings and hypotheses derived through overview visualizations, however, are met with more confidence, because the user actually observed the behavior himself.
+%%
+The participation and steering of the analysis using human pattern-detection and expert knowledge is preferred.
+%%
+%Incorporating data mining algorithms with overview visualizations of statistical measures allow for reasoning about risk factors using underlying epidemiological data sets rather than predefined hypotheses.
+%%
+%We  a higher confidence of an expert in a result he or she actually observed in the data rather 
+%One major advantage we observed is the confidence of an expert in the results he or she actually observe.
+%%
+Observing \emph{expected} correlations, which match the expert knowledge, strengthens the confidence in the method and, subsequently, in the hypothesis generated from unanticipated relationships.
 %%
@@
 %%
-Investigations of the the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results differing in detail.%, but differences  %differing in detail.
-%%
-The overall explanatory power towards the numerical feature was lower.
+Investigations of the the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results differing in detail. %, but differences  %differing in detail.
+%%
+The overall explanatory power towards the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results towards the expected result.
+%%
 %Our investigation regarding the hepatic steatosis data set confirms differences in results
@@
 %%
-Poor guidance lead to potentially overlooked relationships.
-%%
-We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose.
-%%
-Foe example, formulas with a static target feature (e.g.
-$Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z ray.
+Poor guidance potentially leads to overlooked relationships.
+%%
+We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
+%%
+%Foe example, formulas with a static target feature (e.g.
+$Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z-ray.
 %%
@@
 Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern.
-The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
+The University of Greifswald is a member of the √¢¬Ä¬òCentre of Knowledge Interchange√¢¬Ä¬ô program of the Siemens AG.
 This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.}

### 2015-03-29 draft.tex
revised related work

@@
 %%
-% der Satz kann raus. 1. Teil ist wirklich "common sense", 2. Teil wird im n√§chsten Satz mit "...closest to ours" nochmal genannt.
 %We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
@@
 %%
-Generalized Pairs plots (\texttt{GPLOM√Ç¬¥S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
 %%
@@
 %%
-Their research agenda acts as guideline for our design, comprising applications using domain experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
+Their research agenda acts as guideline for our design, comprising perceptual tuning towards human pattern recognition of important aspects, scalability between different data set sizes and application evaluation with domain experts.
 %%
@@
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
-Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described by their most reliable linear regression model.
+Ahmadi et al.
+\cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional data into subspaces, which are then described by their most reliable linear regression model.
 %%
@@
 %%
-They focus on presenting an algebraic for efficient regression-model calculation to find the best fit for a subspace.
+They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
 %%
@@
 %%
-They analyzing subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM¬´S}) of quality metrics. %describing the features.
+They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOMÂ«S}) of quality metrics. %describing the features.
 %%
@@
 %% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
-Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with interactive data mining tools.
+Niemann et al. \cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
 %%
@@
 %%
-In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{Evolution Features}) that describe latent temporal information across the study waves.
+In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
 %%
@@
 %%
-% Verben in ‹berschriften immer groﬂ
 \subsection{Cube Description Using Regression Formula Notation}
@@
 %\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction.}
-
-% -Dependent & -Based in ‹berschriften groﬂ, sicherheitshalber nochmal Steffi fragen
-% oder: Dimension Reduction Depending on the Target Variable
+%%
 \subsection{Target-Variable-Dependent Dimension Reduction}

### 2015-03-29 draft.tex
minor revisions

@@
 %%
-Statistical programs, such as \texttt{SPSS}, are used to analyze the data using the classical sequential epidemiological workflow.
+Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
 Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
@@
 %%
-Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
+Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
 %%
@@
 %%
-The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
+The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical images.
 %%
@@
 %\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
-
+%%
 %\paragraph{Data Types.}
 %%
-Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
-%%
-The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
-%%
-The data heterogeneity has to be taken into account when the analysis method is chosen.
+Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
+%%
+%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
+%%
+%The data heterogeneity has to be taken into account when the analysis method is chosen.
 %%
@@
 %%
-Another important confounder is \emph{Gender}.
+Another important confounder is \emph{gender}.
 %%
@@
 %%
-A example of a regression formula would be $Kidney\_Disorder ~ Smoking + Obesity$.
+A example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
 The most used regression operators comprise:
@@
 %%
-\emph{Nagelkerke $R^2$} cannot be compared to $R^2$ values extracted from linear regression model.
+\emph{Nagelkerke $R^2$} behave different than $R^2$ values extracted from linear regression model.
+%%
+Comparisons have to be handled with care.
 
@@
 %%
-However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study ("\emph{method loyality}").
+However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study. % ("\emph{method loyality}").
 %%
@@
 %%
-The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target variables.
+The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
 %%
@@
 %%
-The major advantage of this notation is that it comes natural to anyone familiar with regression analysis since it is the standard way of expression 
+The major advantage of this notation is that it comes natural to anyone familiar with regression analysis since it is the standard way of expression.
 %%

### 2015-03-31 draft.tex
cleaned up comments

@@
 %% OnlineID. Otherwise, you may safely leave it at ``0''.
-%% TODO: Blind Review: Replace with online ID
 \onlineid{175}
@@
 %% Paper title.
-%% TODO: Replace towards with regarding?
 
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernard Preim}
+\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernhard Preim}
 \authorfooter{
@@
 \item
- Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,niemann,preim\}@ovgu.de
+ Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,uli.niemann,preim\}@ovgu.de
 \item
@@
 %%
-%These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
-%%
 Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
@@
 %%
-This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
-%%
-A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
+This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization that reveals insights into the data by highlighting relationships.
+%%
+A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
 %%
@@
 %%
-We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
+We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
 %%
@@
 %%
-Furthermore, the insights can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
-%%
-During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then statistically analyzed.
+Furthermore, the insights can be used to derive recommendations regarding a healthy lifestyle or to provide information about widespread diseases.
+%%
+During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then are statistically analyzed.
 
@@
 %%
-These data may be quantified based on user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
+These data may be quantified based on user-defined landmarks, which describe attributes like the shape, volume or diameter of a structure.
 %%
@@
 %%
-%Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
 Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
 %%
-% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
 Due to the amount of data and only limited overview techniques, possible correlations may be missed.
 %%
-Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
+Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature and mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
 %%
@@
 %%
-This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
+This is achieved by providing template regression formulas, which are applied to all potential feature combinations.
 %%
@@
 %%
-Difference views between regression formulas allow to assess the influences of individual variables on the process.
+Difference views between regression formulas allow to assess the influences of individual features on the process.
 
@@
 \begin{itemize}
+	%%
 	\item An overview visualization technique representing feature interactions using target features.
+	%%
 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
+	%%
 	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
-	\item The analysis of confounding variables by providing comparison views between different formula results.
-%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
+	%%
+	\item The analysis of confounding features by providing comparison views between different formula results.
+	%%
 	\item The expansion and adaption of our method to any data with our open web-based system.
+	%%
 \end{itemize}
 %%
-The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
+The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
 %%
@@
 %%
-In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
+In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
 %%
@@
 %%
-Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
-%%
-As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
+Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
+%%
+As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
 \begin{itemize}
 	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
-	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
+	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
 	\item Statisticians assess the association of the derived features towards the investigated disease.
@@
 
-In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step.
+In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step.
 %%
@@
 %%
-In this work, we focus on a similar approach to deriving insight and even new hypotheses through the data rather than using it just for confirmatory analysis.
-%An alternative data-driven approach is described as follows.
-%%
-% \paragraph{Data Driven Hypothesis-Generation-based Analysis.}
-% We already described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the variable listing step \cite{Klemm2014VIS}.
-% %%
-% The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
-% %%
-% Employing \emph{hypotheses generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
-
+In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
+%%
 \subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
 %%
-%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
-%%
 There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
 %%
-%We focus on cohort studies are a major epidemiological tool for gathering epidemiological features.
 We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
@@
 %%
-The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical images.
+The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
 %%
@@
 %%
-Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
+Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
 %%
@@
 %%
-%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
-%%
-%\paragraph{Data Types.}
-%%
 Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
 %%
-%The data space comprises continuous variables (somatometric variables, such as body weight) as well as categorical variables (e.g. graduation type).
-%%
-%The data heterogeneity has to be taken into account when the analysis method is chosen.
-%%
-Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
+Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
 %%
@@
 %%
-We incorporate image-derived features, but do not put focus on analyzing image data.
-%%
-More discussion of cohort study data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.
-%%
-%\paragraph{Image Data.}
-% \com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
-% %%
-% Since the Rotterdam study, many modern cohort studies include medical image data.
-% %%
-% For ethical reasons, the imaging modalities usually not include ionizing radiation.
-% %%
-% The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
-% %%
-% These data are hard to analyze as they require segmentations to outline structures of interest.
-% %%
-% This process is prone to inter- and intra-observer variability when carried out manually.
-% %%
-% Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
-% %%
-% Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).
+We incorporate image-derived features, but do not focus on analyzing image data.
+%%
+More discussions of cohort study data types and characteristics can be found in \cite{Preim2014, Toennies2015}.
+%%
 \\\\
-%\paragraph{Confounding Features.}
 Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
-%Confounding features have to be considered in epidemiological analysis, as they influence both the exposure as well as the outcome of an analysis.
-%%
-%Features influencing the exposure as well as the outcome of an analysis are called confounders.
 %%
@@
 %%
-Regression analysis is the most important statistical tool when analyzing epidemiological data.
-%%
-The presented work is based on regression analysis as well.
+Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
 %%
@@
 %%
-Different metrics, such as the weightings itself and associated $p$ values describe the resulting function (the \emph{model}).
+Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
 %%
@@
 %%
-The value is in the range [0,1], where 1 encodes a perfect fit.
+The value is in the range $[0, 1]$, where 1 encodes a perfect fit.
 
@@
 Dependent \sim Independent_1 + ... + Independent_n
-%Dependent \sim \sum_{i=1}^n Independent_i
 \label{eq:RegressionNotation}
@@
 %%
-A example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
-The most used regression operators comprise:
+An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
+The most commonly used regression operators comprise:
 %%
 \begin{itemize}
-	\item $+,-$ inclusion/exclusion of the variable,
+	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
 	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
-	\item $*$ inclusion the variables as well as their interactions (e.g. $x*y$)
+	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
 	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
@@
 %%
-The type of the target feature restricts the regression type.
+The class of the target feature restricts the regression type.
 %%
@@
 %%
-Logistic models as opposed to linear models do not allow for extracting an $R^2$ quality of fit value.
+Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
 %%
@@
 %%
-\emph{Nagelkerke $R^2$} behave different than $R^2$ values extracted from linear regression model.
+\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
 %%
 Comparisons have to be handled with care.
-
-%\com{Regression cubes are an important tool for checking statistical resilience.}
-%\com{Describe how regression cubes work (including notation).}
-%\com{Move this to Regression Cube part?}
-
+%%
 \subsection{The Study of Health in Pomerania (SHIP)}
@@
 %%
-However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study. % ("\emph{method loyality}").
-%%
-%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
-%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
-%
-%They slightly differ in clinical focus, acquired data and epidemiological research questions.
-%
-%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
-%
-%Data are collected without focus on a group of diseases.
-%
-%This allows to query the data regarding many diseases and conditions.
-%
-%
+However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
+%%
 \section{Prior and Related Work}
@@
 %%
-%We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
-%%
 \\\\
-The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} are closest related to ours regarding different aspects.
+The work of Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan}, Angelelli et al.~\cite{Angelelli} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
 %% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
-Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
+Akin to our approach, Piringer et al.~\cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
 %%
@@
 %%
-They also incorporate 3D visualizations for each pairwise combination, but mainly because its popularity with their target domain engineers.
+They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
 %%
@@
 %%
-The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
+The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models in terms of \emph{different} features.
 %%
@@
 %%
-%We share the matrix-arrangement of results depicting relationships between independent features in a model.
 
 %% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
-We use the term Regression Cube for a comparative 3D representation of multiple regression models.
-%% einheitlich 2D-XXX oder 2D XXX, meist ist es 2D XXX
-Chan et al. \cite{Chan} use the same term to describe an extension of the 2D Scatterplot representation of a linear regression models (incorporating solely metric features) to a 3D Cube.
+We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
+%%
+Chan et al.~\cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
 %%
@@
 %%
-Similar to \cite{Piringer}, they focus on highlighting details of the included models, rather than comparing models consisting of different features.
+Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
 %%
@@
 
-Zhang et al. \cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
+Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
 %%
@@
 
-Angelelli et al. \cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
+Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
 %%
@@
 %%--
-Shneiderman et al.
-\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data.
+Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data.
 %, e.g.
@@
 %Shneiderman and colleagues also present challenges,
-The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
+The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
 %%
@@
 %%
-%%
-
-%%
-In their survey on interactive information visualizations for health record data, Rind et al. \cite{Rind} identified future work addressing the usability of such frameworks.
+
+In their survey on interactive information visualizations for health record data, Rind et al.~\cite{Rind} identified future work addressing the usability of such frameworks.
 %%
@@
 %%
-Schreck and Keim \cite{Schreck} presented a method for analyzing social media data with focus on characterization and cause of an epidemic.
+Schreck and Keim \cite{Schreck} present a method for analyzing social media data with focus on characterization and cause of an epidemic.
 %%
@@
 %% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
-Steenwijk et al. \cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
+Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
 %%
@@
 %%
-Dai et al.
-\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
+Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
 %%
@@
 %%
-Time-dependent epidemiological data is visualized by Chui et al.
-\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
+Time-dependent epidemiological data is visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
 %%
@@
 %% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
-Bertini et al. \cite{Bertini} present an overview over quality metrics describing high dimensional data.
-%%
-Their research agenda acts as guideline for our design, comprising perceptual tuning towards human pattern recognition of important aspects, scalability between different data set sizes and application evaluation with domain experts.
-%%
-%% TODO: Use this one?
-"Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich √É¬§ndert, wenn man per Zufall die Populationen ver√É¬§ndert, um zu schauen wie zuverl√É¬§ssig diese Statistik ist.
+Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
+%%
+Their research agenda acts as guideline for our design, comprising perceptual tuning towards human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
-Ahmadi et al.
-\cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional data into subspaces, which are then described by their most reliable linear regression model.
+Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
 %%
@@
 %% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
-Albuquerque et al. \cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
+Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
 %%
@@
 %%
-They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOMÂ«S}) of quality metrics. %describing the features.
+They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM'S}) of quality metrics. %describing the features.
 %%
@@
 %%
-Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
-%%
-The approach was applied to a cognitive aging studies, where new hypotheses could be derived.
-
-%Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
-%\com{Uli, can you please complete this, I do not have the pdf available!}
+Turkay et al.~\cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
+%%
+The approach was applied to a cognitive aging study, where new hypotheses could be derived.
+
 %% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
-Niemann et al. \cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
-%%
-They extracted association rules that serve as basis for a proof-of-concept evaluation.
-%%
-In \cite{Niemann2015}, Niemann et al. improve on classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
-%%
-We try to reproduce their results and investigate their findings further.
-%%
-%\com{Gla√É¬üer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}
-
+Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
+%%
+They extract classification rules that serve as basis for our proof-of-concept tests.
+%%
+In \cite{Niemann2015}, Niemann et al. improve the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
+%%
+We try to reproduce results and investigate findings presented in \cite{Niemann2014} further.
+%%
 \paragraph{Prior Work.}
@@
 %%
-We utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
+We utilized semi-automatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
 %%
@@
 %%
-% erst die Erkl√§rung, dann die Folge daraus =)
 The mosaic plot presented in \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
@@
 
-%% TODO: Check DLPD for correct page numbers
 While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
@@
 The difference of the presented approach towards our previous and related work is twofold:
+%%
 \begin{enumerate}
-	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
-	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
+	%%
+	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+	%%
+	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
+	%%
 \end{enumerate}
@@
 \section{Regression Cube Analysis of Cohort Study Data}
-% TODO: Do we need a preprocessing section?
-%\com{Add Data Preprocessing Section.}
 %%
@@
 %%
-Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
+Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
 %%
@@
 %%
-% unified -> singular [Steffi?]
-A regression model is based on expert knowledge, there is no singular rule how to apply them to a given set of features. Thus, they have to be applied with care.
+A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
 %%
@@
 As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
-%%
-%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
 %%
@@
 %%
-The major advantage of this notation is that it comes natural to anyone familiar with regression analysis since it is the standard way of expression.
-%%
-%This allows for a fast adaptation and in the epidemiological application domain.
+The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
 %%
@@
 \begin{itemize}
-	\item $Z \sim X + Y$ calculates all combinations of two features towards all possible target features.
-	\item $Cancer \sim X + Y + BodyWeight$ includes the $BodyWeight$ feature into all regression models as feature.
-	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features towards the target.
+	%%
+	\item $Z \sim X + Y$ calculates all combinations of two features w.r.t. all possible target features.
+	%%
+	\item $Cancer \sim X + Y + BodyWeight$ includes the $BodyWeight$ feature into all regression models as feature with $Cancer$ as target.
+	%%
+	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features towards the $Cancer$ target.
+	%%
 \end{itemize}
@@
 %%
-% hours -> h da ms ja auch benutzt wird.
-When each regression takes about 50~ms calculation, we have to roughly wait 14~h for the calculation to complete.
+When each regression takes about 50~ms calculation, we roughly have to wait 14~h for the calculation to complete.
 %%
@@
 %%
-%\com{Regression formulas are described using standard regression formula notation by introducing variables.}
-%%
-%\com{Target Group: Physicians/Epidemiologists with basic statistical background to understand regression formulas and Interaction.}
+An approach for this is presented in the following section.
 %%
 \subsection{Target-Variable-Dependent Dimension Reduction}
-% \com{Uli, Can you please proofread this?}
-% %%
 % The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
@@
 %%
-In general, many of them exhibit low or none correlation at all towards the target feature.
+In general, many of them exhibit a low or no correlation at all in view of the target feature.
 %%
@@
 %%
-% nochmal checken wo $F$ eingef¸hrt wird, hier oder im ¸bern‰chsten Satz?
-The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
+The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$, which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
 %%
@@
 %%
-% Ggf. wurde $F$ schon oben eingef¸hrt, TODO
-Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.
+Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.
 %%
@@
 %%
-The CFS algorithm will select the feature which has the largest explanatory power and discards the other dimensions.
+The CFS algorithm will select the feature, which has the largest explanatory power and discards the other dimensions.
 
@@
 In our epidemiological data, we usually observed a number of 10 to 30 features.
-%For features, which affect the most other features (Recall "Confounder" in Section~\ref{sec:EpidemiologicalData}), e.g., age or gender,  the number is significantly larger (about half the features in the whole data set).
 %%
@@
 %%
-%The number of features extracted is not directly associated with the overall number, but rather on their information entropy towards each other.
-%%
-
-With this method, we are able to derive the interesting regression models in a reasonable time span.
+
+With this method, we are able to derive the interesting regression models in a reasonable time span (seconds to minutes instead of hours).
 %%
@@
  \centering
- \includegraphics[width=3.5in]{figures/cube}
+ \includegraphics[width=1.0\linewidth]{figures/cube}
  \caption{
- (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $Age$.
+ (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
  %%
@@
  %%
- (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
+ (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
  %%
@@
  %%
- The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
+ The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
  }
@@
 The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
-%% TODO: Cite another paper here!
+%%
 As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
@@
 %%
-Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
-%%
-We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
+Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
+%%
+We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
 %%
@@
 %%
-Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
-%%
-$R^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
+Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube}b).
+%%
+$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
 %%
@@
 %%
-% e.g. wenn man es w‰hlen kann
-% i.e. wenn es so ist
 Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
 %%
-% Adverb immer vors Hauptverb oder Satzende
 Thus, the cube can be easily extended using other regression types.
@@
 \\\\
+%%
 Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
@@
  \centering
- \includegraphics[width=3.0in]{figures/workflow}
+ \includegraphics[width=1.0\linewidth]{figures/workflow}
  \caption{
@@
  %%
- % The analysis starts with a set of features given by the data set.
- %%
- [\textbf{1.AF}] The workflow starts by declaring a formula regarding to specify a hypothesis, or use a predefined formula for a hypothesis-free analysis.
+ [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
  %%
@@
  %%
- [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or selecting a difference view.
+ [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
  %%
@@
  %%
- [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yields insights and hypothesis about feature relations.
+ [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
  }
@@
 %%
-%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
-%\com{Workflow for different analysis approaches.}
-%\com{0. Select the Variables.}
-%\com{1. Select the Formulas}
-%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
-%\com{b) Analysis without hypothesis with confounder (z ~ age)}
-%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
-%\com{d) Use Difference View}
-%\com{d) Clarify using confounder (Var ~ x)}
-%\com{3. Find interesting pane.}
-%\com{4. Get Details on Demand.}
-%%
-% ‹bergang ist recht holprig gewesen
-Our Regression Cube models is well-suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:
+Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
 
@@
 
-\textbf{Show the Important} [\textbf{2.SI}]. The 3D-cube visualization acts as an overview over the whole data set.
-%%
-Regression models with large $R^2$ values can be spotted here fast, steering the users attention to the respective slice.
+\textbf{Show the Important} [\textbf{2.SI}]. The 3D cube visualization acts as an overview over the whole data set.
+%%
+Here, regression models with large $R^2$ values can be spotted fast, steering the user's attention to the respective slice.
 
@@
 Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
-in a hover-modal on a currently selected data point).
+in a hover modal on a currently selected data point).
 \\\\
-We use the squared bracket abbreviation for each step to denote the affiliation the system design section later on.
-%%
-As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
+We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
+%%
+As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
 %%
@@
 %%
-This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
+This can be carried out either to refine the current formula based on observations, or to create a new regression cube for a difference view.
 %%
@@
 %%
-Using the operators, the dynamic variables and the dataset features, many different assumptions can be expressed.
-%%
-To support \emph{hypothesis-free} analysis, we provide a default formula:
+Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
+%%
+To support the \emph{hypothesis-free} analysis, we provide a default formula:
 
@@
 
-% warum war da ein "either", sollte der Satz etwas anderes ausdr¸cken?
 Hypotheses about the data are easily built up by relating dynamic variables with the regression operators.
@@
 %%
-$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
+$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
 %%
@@
 %%
-This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combination influence the result.
-%%
-Therefore, the cube incorporates this interaction for all $X$,$Y$ and $Z$ values as independent features.
-%%
-
-$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
+This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with `$:$') being relevant for the target feature, but it is not clear how other feature combinations influence the result.
+%%
+Therefore, the cube incorporates this interaction for all $X$, $Y$ and $Z$ values as independent features.
+%%
+
+$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect that age has in view of the target $Cancer$ feature.
 %%
@@
 %%
-Regression Cubes can be compared by creating difference views.
+\emph{Regression Cubes} can be compared by creating difference views.
 %%
@@
 For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
-%%
-% \paragraph{Formula Types.}
-% %%
-% Input formulas reflect \emph{hypotheses} about the data.
-% %%
-% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
-% %%
-% Here we show just a few examples on how formulas can serve different purposes.
-%
-% $Z \sim X + Y$ is the default formula.
-% %%
-% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
-% %%
-% Each slice represents a different target feature.
-% %%
-% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.
 
@@
  \centering
- \includegraphics[width=3.5in]{figures/system}
+ \includegraphics[width=1.0\linewidth]{figures/system}
  \caption{
@@
  %%
- %\com{TODO: Adapt formula to final image.}
- %%
- %Context information is shown for the regression model of $Parenchym \sim Bronchitis + AlcoholLast30Days$, depicting details about the confidence intervals, coefficients and associated p-values.
- %%
- %\com{Screenshot of the cube visualization embedded in the framework}
- %\com{Also include active tooltip}
  }
@@
 \end{figure}
-%% scientific english: kurz und langweilig ;)
+%%
 We designed our system to be freely accessible and easy to use.
@@
 %%
-%%
-The focus lies on creating an overview visualization and gaining insight into relationships into the data, which triggers further analyses (maybe with other (statistical) tools).
+The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
 %%
@@
 \\\\
-% \subsection{Design and Visualization Techniques}
 %%
 \noindent Design choices and spaces are restricted by the underlying technologies and devices.
-%% TODO: Umschreiben
-Resting collaboration and exchange about data and methods on web-based technologies shows promising results, as there is no set-up time involved and domain experts can use the methods from any computer connected with the web.
-%%
-Web technology is by design based on a client-server architecture, making it easy to outsource computationally heavy tasks on server-clusters and transferring results to the client device.
-%%
-The design space spanned by web-technologies is different to standard WIMP-applications: right click menus, modal windows and menu bars are not established UI components in this context.
-%%
-Therefore, we have to adapt our design respecting the affordances of web pages.
+%%
+The exchange of data and methods using web-based technologies offers many advantages.
+%%
+There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
+%%
+By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
+%%
+The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
+%%
+Therefore, we have to adapt our design respecting the conventions of web pages.
 %%
@@
 This allows for a fast learning of the individual system parts.
-%%
-%More importantly, it allows the user to focus all mental resources on the analysis tasks, rather than wrangling with configuring the system to serve the current task or hypothesis.
 %%
@@
 	%%
-	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file [1.AF].
-	%%
-	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube [2.SI].
-	%%
-	\item The \emph{formula editor} allows to adjust the formula w.r.t. a hypothesis or conducting a \emph{hypothesis-free} analysis.
-	%%
-	It also allows to select a reference formula for creating a difference-cube [3.ZF].
+	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
+	%%
+	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
+	%%
+	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
+	%%
+	It also allows to select a reference formula for creating a difference cube [1.AF, 3.ZF].
 	%%
 \end{itemize}
-%%\com{Focus on simplification and ease of use, reducing the number of UI elements as much as possible (Statistical Visualizations often tend to put all information and settings at once to the user, which is overwhelming for the user).}
-\paragraph{File-Upload and Classification [1.AF].}
+%%
+\paragraph{File Upload and Classification [1.AF].}
 %%
@@
 %%
-Each line after that represents one subject (row) and their feature manifestations.
-
-\textbf{Classification.}
+Each line after that represents one subject (row) and its feature manifestations.
+
+\textbf{Encoding via CSV Files.}
 %%
@@
 %%
-We, however, need to ensure the correct variable type classification and have to enforce some basic standards.
+However, we need to ensure the correct variable type classification and have to enforce some basic standards.
 %%
@@
 %%
-This seems obvious, but in fact many cohort study data sets encode categorical features using ID-values, which are denoted in a data dictionary.
+Although this seems obvious, many cohort study data sets encode categorical features using ID values that are denoted in a data dictionary.
 %%
@@
 %%
-Security issues are raised by uploading data into an online service, such as our prototype.
+Security issues are raised by uploading data into an online service such as our prototype.
 %%
@@
 %%
-%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
-%%
-We apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
-%% TODO: Implement this
+We calculate a SHA-256 hash to derive the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
+%%
 Data sets are deleted from the server after closing a session.
@@
 %%
-Entering a formula is provided via text input.
-%%
-On formula input, a context modal displays all data set variables, as well as the available operators and their function.
+Entering a formula is facilitated via text input.
+%%
+On formula input, a context modal displays all data set features as well as the available operators and their function.
 %%
@@
 %%
-Auto-completing input variables also simplifies the approach and also works as spell-check of variable names.
+Auto-completing input features also simplifies the approach and works as spell check of feature names.
 %%
@@
 %%
-These are then divided by the number of available statistical back-end processors available, driving a \emph{cloud-computing}-based approach.
+These are then divided by the number of available statistical back-end processors, driving a \emph{cloud-computing}-based approach.
 %%
@@
 %%
-In praxis, data transmission and differences in machine specifications always influence the speed.
-
-%%
-\textbf{Difference-Cube.}
+In practice, data transmission and differences in machine specifications always influence the speed.
+
+%%
+\textbf{Difference Cube.}
 Adding a formula also adds it to the reference selection for a difference cube.
@@
 %%
-%%If the user selects a reference formula, the absolute difference of $R^2$ values between the cube described by the reference and the currently active formula are calculated and shown.
-%%
-
-%\com{Dynamic Type Checking, Auto-Completion Analysis Trigger, Comparison between Cubes, triggering of cube calculation, parallelization of formula calculation}
-
 \subsection{Regression Cube Visualization [2.SI].}
@@
 %%
-Results from the statistical processors are uploaded into the visualization slice by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.
-
-\paragraph{Usage of a Regression Prism for information reduction.}
+Results from the statistical processors are uploaded into the visualization slice by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in progress.
+
+\paragraph{Usage of a Regression Prism for Information Reduction.}
 %%
@@
 %%
-Therefore we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
+Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
 %%
@@
 %%
-Along the diagonal, $X$ and $Y$ assume the same feature, $Z ~ X + Y$ turns into $Z ~ X$ because the regression automatically ignores doublings.
+Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
 %%
@@
 
-\paragraph{3D-Prism as Data Mini-Map.}
-%%
-The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
-%%
-It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
-%%
-%It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
+\paragraph{3D Prism as Data Mini-Map.}
+%%
+The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
+%%
+It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
 %%
@@
 %%
-For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System} b).
+For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System}b).
 %%
@@
 %%
-%% TODO: Add reference for disadvantages of 3D information visualizations!
 \paragraph{Tackling the disadvantages of 3D information visualization.}
-3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
+%%
+3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
 %%
@@
 %%
-Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
-%%
-It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
-%%
-Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
+Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
+%%
+It also creates a sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
+%%
+Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
 %%
@@
 %%
-The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
+The cube is always oriented according to the 2D representation, allowing for an easy mental combination of the two representations.
 %%
@@
 %%
-We provide also a zoom functionality using the mouse wheel input.
-%%
-%\\\\
+Also we provide a zoom functionality using the mouse wheel input.
+%%
 \paragraph{Cube Slice Selection [3.ZF].}
@@
 	%%
-	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
-	%%
-	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
-	%%
-	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
+	In medical volume data renderings, slicing views are very common to view details on a selected plane in the scene.
+	%%
+	We employ this technique for selecting cube slices (e.g., by moving a plane via vertical mouse input while pressing the right mouse button).
+	%%
+	We, however, still display the whole 3D object instead of cutting away information towards the slice position.
 	%%
@@
 %%
-The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
-%%
-the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
-%%
-%It also allows for a fast slicing through the cube and assessing the information.
-%%
-%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
-%\\\\
+The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
+%%
+The values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
+%%
 \paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
 %%
-The 2D mosaic visualization (Fig.~\ref{fig:System} c) shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
-%%
-To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
+The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
+%%
+To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation-based feature selection.
 %%
@@
 %%
-By hovering over a data entry in the plot, a tooltip model displays detailed information about a models coefficients, associated $p$ values and confidence intervals.
-%%
-
-%\com{Simplicity of interface to allow for steep learning curve.}
-%\com{Visualization as skewed cube to reduce visual clutter and remove double entries.}
-%\com{Cube acts as visual mini map to give a global impression over the data.}
-%\com{current pane shown using a heat map visualization.}
-%\com{details on demand for selected regression formula.}
-%\com{comparison view using reference cubes.}
+By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values and confidence intervals.
+%%
 
@@
  \centering
- \includegraphics[width=3.5in]{figures/implementation}
+ \includegraphics[width=1.0\linewidth]{figures/implementation}
  \caption{
- The front-end (left) is written in \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and uses different Javascript libraries, such as \emph{Angularjs}, \emph{Three.js} and D3js.
- %%
- The web-server back-end (right) is written in \emph{Nodejs}, hosting is carried out using \emph{Heroku}.
+ The front-end (left) is realized with \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and different Javascript libraries, such as \emph{Angular.js}, \emph{Three.js} and D3.js.
+ %%
+ The web-server back-end (right) is written using \emph{Node.js}, hosting is carried out using \emph{Heroku}.
  %%
@@
  Additional statistical back-ends can be attached to the system to decrease the computation time.
-% The number of attached machines to the system 
-%    the  the  Describe the Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend.}
+ %%
  }
@@
 %%
-% Using web technologies for prototyping allows for:
-% \begin{itemize}
-% 	\item easy exchange of software, without the need of installing anything other than a modern web browser,
-% 	\item fast iterative adaption of method changes, without the need of installing updates,
-% 	\item outsourcing computation-heavy tasks to server-clusters, which do not have to be in local proximity.
-% \end{itemize}
-% %%
-%%
 We rely on web-based technologies for our prototype.
@@
 %%
-\texttt{Angular.js} is forcing developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
-%%
-The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of UI elements with proper stylings.
+It forces developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
+%%
+The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements with proper stylings.
 %%
@@
 %%
-We use a shader-based solution by rendering the cube as a sprite based particle system, allowing to customize color and opacity of every data point.
+Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
 %%
@@
 %%
-The first one is the web-server, which is written in \texttt{Javascript} using \texttt{NodeJS}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
+The first one is the web-server, which is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
 %%
@@
 %%
-\texttt{OpenCPU} is a \texttt{R} package and provides a API for accessing it via HTTP calls \cite{Ooms}.
+\texttt{OpenCPU} is an \texttt{R} package and provides an API for accessing it via HTTP calls \cite{Ooms}.
 %%
@@
 %%
-The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Nodejs Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
+The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Node.js Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
 %%
@@
 %%
-The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku homepage. 
-%%
-%The front-end source as well as the \texttt{Nodejs} backend can be found here: .
-%%
-% \com{Use VAST'14 as Guideline}
-% \com{Cloud computing}
-% \com{Volume Rendering vs Particles}
-% \com{Speed is merely a matter of available server machines due to parallelization process.}
-% \com{Security enabled by hashing files.}
-% \com{Bootstrap/Angular/D3 Frontend, Node backend, R/OCPU Backend (Fig.~\ref{fig:Implementation})}
-
+The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku app.
+%%
 \section{Application} \label{application}
@@
  \centering
- \includegraphics[width=7.0in]{figures/application}
+ \includegraphics[width=1.0\linewidth]{figures/application}
  \caption{
- The analysis of the numerical and dichotomized target feature depicting liver-fat values yields similar results (left).
+ The analysis of the numerical and dichotomized target feature depicting liver fat values yields similar results (left).
  %%
@@
  %%
- A high correlations between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was found during the analysis using the 2D mosaic plot.
- %%
- \com{Caption Breast Fat Data}
+ A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D mosaic plot.
+ %%
+ The hypothesis-free analysis of the breast fat data set (right) w.r.t.
+the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
  }
@@
 %%
-The breast fat data set is the foundation for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
+The breast fat data set is the basis for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
 %%
@@
 %%
-Usually, only a few features depicting a hypotheses are compiled into a data set to assess them using statistical tools.
+Usually, only a few features depicting a hypothesis are compiled into a data set to assess them using statistical tools.
 %%
@@
 %%
-To assess the ability of a system to discover knowledge is difficult to measure.
-%%
-Lam et al.
-\cite{Lam2012} propose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
+The ability of a system to discover knowledge is difficult to measure.
+%%
+Lam et al.~\cite{Lam2012} propose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
 %%
@@
 %%
+We employ \emph{VDAR} for analyzing our system.
+%%
 \paragraph{Participants, Setup and Procedure.}
 %%
-%One major advantage of focusing on web-based research was the possibility to bridge the geographical gap towards the epidemiological domain experts.
-%%
 We conducted a web-based analysis by using an online-meeting software, which features voice chat as well as screen sharing.
 %%
-Starting an analysis using these techniques take about 5-10 minutes of setup time.
+Starting an analysis using these techniques takes about 5-10 minutes of setup time.
 
@@
 %%
-After that the experts use the system on their own computers.
+Afterwards, the experts use the system on their own computers.
 %%
@@
 %%
-We conducted a study with two participants, who also co-authored this publication.
-%%
-Domain expert for the breast fat data set analysis is \emph{KH}, a clinician (10 years of experience) with focus on epidemiological research.
-%%
-\emph{KH} is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
-
-The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data with.
-%%
-
-\emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
+We conducted the analysis with three participants.
+%%
+\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set analysis.
+%%
+She is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
+%%
+The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data.
+%%
+The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
 %%
@@
 %%
-Liver fat concentrations of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class.
+Liver fat concentrations of no more than 10\% are mapped to the `negative' class; values greater than 10\% are mapped to the `positive' class to indicate absence or presence of the disease.
 %%
@@
 %%
-The MRI scans for each subject were introduced in SHIP-2, so the target feature is present for just one of three study waves.
-
-%%
-Apart from the target feature, the data set contains 199 features, comprising socio-demographic features (gender, age), consumption behavior (e.g.
-alcohol or tobacco), laboratory data (e.g.
-sera concentrations), and two features depicting the liver ultrasound.
-%%
-%The acquisition wave is denoted using the appendix; 85 for S0 (SHIP-0), 55 for S1, 55 for S2 and 10 time-independent SNPs (DNA base pairs).
-%%
-The acquisition wave is denoted using the appendix; \emph{s0} meaning that their values have been recorded in SHIP-0 (first study moment), 50 features for \emph{s1} and 55 for \emph{s2}, alongside with 10 time-independent SNPs (DNA base pairs).
-%%
-
-In \cite{Niemann2014}, the authors show different class distribution of liver fat concentrations of women and men.
-%%
-For women a association between age and liver fat was identified.
-%Also, for women they identify that there is a association between age and liver fat concentration.
+The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present the last study wave.
+
+%%
+Apart from the target feature, the data set contains 199 features, comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol or tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
+%%
+The acquisition wave is denoted using the appendix; 85 features with appendix \emph{s0} denote their affiliation to SHIP-0 (first study moment), 50 features for \emph{s1} and 55 for \emph{s2}, alongside with 10 time-independent SNPs (DNA base pairs).
+%%
+
+%%
+In \cite{Niemann2014}, the authors show different class distributions of liver fat concentrations of women and men.
+%%
+For women, an association between age and liver fat was identified.
 %%
@@
 %%
-Based on these observations, we perform our analysis on three sub-populations: \emph{females}, \emph{males} and \emph{females} older than 52 years.
+Based on these observations, we perform our analysis on three populations: \emph{males}, \emph{females (all ages)} and \emph{females older than 52 years}.
 %%
 \subsection{The Breast Fat Data Set}
-The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
+The breast fat data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
 %%
@@
 
-The data comprises 1.186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
-%%
-It contains 231 dimensions, holding information about somatometric features (e.g.
-body size or weight) consumption behavior (e.g.
-alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women specific features (e.g., number of born children or contraception type) as well as mammography features.
-% It contains 231 dimensions, holding information about
-% \begin{itemize}
-% 	\item \emph{somatometric features}, e.g., body size, weight or BMI,
-% 	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume,
-% 	\item \emph{personal history}, e.g., occupation, marital status,
-% 	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
-% 	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
-% 	\item \emph{mammography features} derived from the image data.
-% \end{itemize}
-%%
-The latter were derived using manual segmentation of MRI image data for each subject carried out by radiologists.
-%%
-%%
-%\com{Describe Parenchym tissue and its relation to breast cancer.}
-%\com{Describe how the data was acquired and preprocessed}
+The data comprises 1,186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
+%%
+It contains 231 dimensions, holding information about somatometric features (e.g., body size or weight) consumption behavior (e.g., alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women-specific features (e.g., number of born children or contraception type) as well as mammography features (e.g., fat content or parenchyma tissue proportion to volume).
+%%
+The latter were derived from MRI data for each subject, which was manually segmented by radiologists.
+%%
 
@@
 %%
-All data related to the mammography was stored in a additional file.
+All features related to the mammography attributes were stored in an additional file.
 %%
@@
 %%
-All features were renamed to be expressive, e.g., \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
+All features were renamed to be self-explaining, e.g., \emph{chro\_09a} is now denoted as \emph{Disease\_Osteoporosis}.
 %%
@@
 %%
-%All male subjects were removed as their data does not contribute to the analysis.
-%%
-\subsection{Case 1: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
-%%
-We refer to each analysis step regarding to their belonging in the VA-Mantra (Sec.~\ref{sec:Workflow}).
-%%
-The analysis goal is reproducing results based on \cite{Niemann2014}.
-%%
-Therefore, \emph{UN} started the \textbf{1.AF} step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
-%%
-The \textbf{2.SI} step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application} a).
-%%
-\textbf{3.ZF} by slicing through the cube using the mouse input to inspect the hotspots using the mosaic plot (\textbf{4.DD}) revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as creatinine (used renal retention parameter) and uric acid (among others used as gout and diabetes risk factors) magnitudes.
-%%
-Similar results present for analyzing the \emph{female} groups.
-%%
-\emph{UN} could reproduce most results, some features features exhibit lower correlations though, e.g., creatinine magnitudes.
-%%
-A slight influence of age on the target feature could be observed for women ($R^2$ of $0.09$ for females compared to $0.02$ for males).
-%%
-Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g. \emph{aspartate aminotransferase}.
-%%
-Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
-%%
-% %%
-% The goal at the beginning of the analysis was to reproduce results based on the prior analysis.
-% %%
-% Subjects with a fatty liver could be identified using somatometric features, hepatic steatosis indicator-features as well as creatinine (\com{important for kidney}) and uric acid (\com{gout, diabetes}) magnitudes.
-% %%
-% The analysis was started with the target feature \emph{liver fat content} (\emph{mrt\_liverfat\_s2}), which was derived from the MRI data analysis.
-% %%
-% \emph{UN} therefore started the calculation using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subects.
-% %%
-% %%
-% %%
-% The data set containing all \emph{males} showed broad consensus with the expected results.
-% %%
-% The 3D-cube showed highlighted features at the end of the cube (Fig.~\ref{fig:application} a).
-% %%
-% Using the mouse-slicing allowed for fast navigation to the respective features.
-% %%
-% They include indicators hepatic steatosis ($R^2$:XX) as well as somatometric features, such as \emph{body mass index} (\emph{BMI}) and waist circumference.
-% %%
-% Feature correlations with \emph{gamma-glutamyl transpeptidase} (\emph{GGT}, indicating damaged livers if elevated) were found, but with low $R^2$ of XX.
+\subsection{Case 1: Hypothesis-Driven Analysis of the Hepatic Steatosis Data Set}
+%%
+We refer to each analysis step with regard to its belonging in the VA-Mantra (recall Sec.~\ref{sec:Workflow}).
+%%
+The analysis goal was reproducing results with our visual analysis framework that are in accordance to the data mining-based results presented in \cite{Niemann2014}.
+%%
+Therefore, \emph{UN} started the [\textbf{1.AF}] step using the dichotomized MRI fat liver concentration and the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.
+%%
+The [\textbf{2.SI}] step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application}a).
+%%
+The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the cube using the mouse input to inspect the hotspots.
+%%
+Analyzing the mosaic plot [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
+%%
+Similar results were present for analyzing the \emph{female} groups.
+%%
+\emph{UN} could reproduce most results, some features exhibit lower correlations though, e.g., \emph{creatinine} magnitudes.
+%%
+A slight influence of \emph{age} on the target feature could be observed for women ($R^2$ of $0.09$ for females compared to $0.02$ for males).
+%% TODO Feedback of Henry!
+Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g., \emph{aspartate aminotransferase}.
 %
-% %%
-% UN analyzed both \emph{female} groups in the same manner, concluding with the same findings regarding \emph{somatometric} features as well as \emph{hepatic steatosis} indicators.
-% %%
-% Using the mosaic plot, relationships incorporating enzyme used for the diagnosis of liver dysfunctions were found, e.g. aspartate aminotransferase.
-% %%
-%
-% For most features, UN found at least tendencies using our approach.
-% %%
-% Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.
-%
-% %Regression cube calculations were carried out for
-
-\textbf{Analysis of non-discretized target feature.}
-%%
-Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
-%%
-The 3D-cube representation already showed generally lower $R^2$ values, but since the analysis is now based on linear regression, the $R^2$ values can not be directly compared.
-%%
-Correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
-%%
-We conclude that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.
-%%
-
-\textbf{Interleukin-6 correlation with liver fat.}
-%%
-During the analysis, one hotspot was always observed in the \emph{SI} and \emph{ZF} incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
+%%
+Due to the difference between our regression model approach and the decision tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features is not expected.
+%%
+
+%%
+\textbf{Analysis of Non-discretized Target Feature.}
+%%
+Since our method can assess numerical target features, the analysis was conducted again for the non-dichotomized target using the same formula.
+%%
+The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
+%%
+The correlation hotspots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
+%%
+We assume that the bias introduced by dichotomizing the fat liver content enforces the findings of liver diseases, while using the numerical features is less expressive.
+
+\textbf{Interleukin-6 Correlation With Liver Fat.}
+%%
+During the analysis, one hotspot was always observable in the [\textbf{2.SI}] and [\textbf{3.ZF}] steps, incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values ($R^2$ of $0.8$, see Fig.~\ref{fig:Application}b).
 %%
@@
 %%
-In the literature, relations between liver cancer \cite{He2013} and also chronic liver diseases \cite{Streetz2003}.
-%%
-Strong effects of \emph{IL-6} with hepatic steatosis were also prior described with mice \cite{Hong2004}.
+The literature described relations between \emph{IL-6} and liver cancer \cite{He2013} as well as chronic liver diseases \cite{Streetz2003}.
+%%
+For mice, strong effects of \emph{IL-6} with hepatic steatosis were described \cite{Hong2004}.
 %%
@@
 %%
-The analysis goal is to find relationships on the breast cancer data using features derived from the mammography analysis.
-%%
-Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
+The analysis aims to find relationships on the breast fat data using mammography analysis features.
+%%
+Relationships between the share parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
+%%
+The [\textbf{1.AF}] was started by \emph{KH} using the default formula for hypothesis-free analysis ($Z \sim X + Y$).
+%%
+At first, she was interested in correlations with the \emph{parenchyma tissue} percentage, which was selected through the drop-down for the z-axis [\textbf{2.SI}].
+%%
+She observed strong correlations with \emph{age}, \emph{body fat percentage}, \emph{hip} and \emph{waist circumference} as well as \emph{menstrual period} or \emph{pregnancy status} as expected (Fig.~\ref{fig:Application} right).
+%%
+Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
+%%
+\emph{Age} is a strong influencing factor, as breast tissue and subsequently the parenchyma tissue degrades over time.
+%%
+\emph{KH} proceeded using [\textbf{3.ZF}] and [\textbf{4.DD}] to check for relationships for different target features, such as current \emph{hormone replacement therapy}, \emph{BI-RADS} (classification of the mammography findings) as well as different diseases, such as \emph{diabetes} or \emph{gout}.
+%%
+She observed relationships matching her expectations and expert knowledge.
+%%
+One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
+%%
+\emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
+%%
+
+%%
+\textbf{Detailed Breast Parenchyma Analysis.}
+%%
+The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
+%%
+Using the 3D cube, \emph{KH} observed several hotspots [\textbf{2.SI}].
+%%
+Navigating to them using the slicing facility of the 3D visualization [\textbf{3.ZF}] highlighted features of high influence, such as image-derived features, as \emph{glandular tissue density} and \emph{parenchyma segmentation} metrics.
+%%
+Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. its strong influence to the parenchyma tissue.
+%%
+A surprising finding was the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
+%%
+The [\textbf{4.DD}] analysis, however, showed only 8 subjects with this disease.
+%%
+Too few subjects impose the risk of a biased finding.
+%%
+The correlation was noted and will be further investigated in the near future using an extensive data set.
+%%
+Lastly, \emph{KH} assessed the influence of contraception-related features, such as use of \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
+%%
+Other consumption behavior features, such as \emph{alcohol intake} also yield no elevated $R^2$ values.
+%%
+\emph{KH} remarked that these features are suspected to have an impact on the parenchyma tissue, but they are less reliable, since they are self-reported.
 %%
@@
 %%
-Our method was well received among the domain experts.
-%%
-For the first time, they were able to derive a overview visualization custom tailored to underlying assumptions.
-%%
-In this context, \emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
-%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
-%%
-%TI commented on the
-
-\textbf{Extracted Hypotheses have to be investigated further.}
+The presented method was well received among the domain experts.
+%%
+For the first time, they were able to derive an overview visualization custom-tailored to underlying assumptions.
+%%
+\emph{KH} noted the ease of use, which "\emph{converts data sets into a feasible form}".
+%%
+She highlighted the efficiency of combining fast target feature selection with visually highlighting interesting results, enabling rapid analysis cycles.
+%%
+To get nearly similar results, she had to spend hours using \texttt{SPSS} and potentially missed interesting hotspots during this process.
+%%
+\emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
+%%
+\textbf{Extracted Hypotheses Have to be Investigated Further.}
+%%
+We map results of complex statistical results into comprehensive visualizations.
 %%
@@
 %%
-A accompanying search for correlations potentially highlighting confounders can be carried out using our method.
-%%
-%This incorporates detailed investigation of potential confounder features as well as looking for outlier subjects.
+An accompanying search for correlations potentially highlighting confounders can be carried out using our method.
+%%
+Statistical validation of an epidemiological end result still has to be carried out by statisticians using their respective tools.
 %%
@@
 %%
-%TI hinted the neccissity of different regression types, which also model non-linear relationships, but do not comprise of comparable quality-of-fit measures, such as the $R^2$.
-%%
-%\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
-%%
-
+
+%%
 \textbf{Overview Visualizations are Preferred over Black-box Methods.}
 %%
-Explorative analysis based on the data gains importance in epidemiology as the data set complexity increases.
-%While the classical epidemiological workflow represents the foundation of the research in this area, explorative analysis based on the data gains importance.
-Results from automatic `black-box' methods, such as data mining algorithm are more often obscure to the expert.
-%%
-Findings and hypotheses derived through overview visualizations, however, are met with more confidence, because the user actually observed the behavior himself.
-%%
-The participation and steering of the analysis using human pattern-detection and expert knowledge is preferred.
-%%
-%Incorporating data mining algorithms with overview visualizations of statistical measures allow for reasoning about risk factors using underlying epidemiological data sets rather than predefined hypotheses.
-%%
-%We  a higher confidence of an expert in a result he or she actually observed in the data rather 
-%One major advantage we observed is the confidence of an expert in the results he or she actually observe.
-%%
-Observing \emph{expected} correlations, which match the expert knowledge, strengthens the confidence in the method and, subsequently, in the hypothesis generated from unanticipated relationships.
-%%
-%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms.
-People believe more in what they see, rather than in what they are presented by an algorithm.
-Regression cubes incorporate this by showing known associations as well as new ones.
-Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}
-
-%%
-\textbf{Using non-discretized features reduces information bias.}
-%%
-Discretization reduces the information space and introduces bias into the data and is therefore avoided in epidemiological research when possible.
+Explorative analysis based on the data gains importance in epidemiology with increasing data set complexity.
+%%
+Results from automatic `black-box' methods, such as data mining algorithms are more often obscure to the expert.
+%%
+Findings and hypotheses derived through overview visualizations, however, are met with more confidence, because the users actually observed the behavior themselves.
+%%
+The participation and steering of the analysis using human pattern detection and expert knowledge is preferred.
+%%
+Observing \emph{expected} correlations matching the expert knowledge strengthens the confidence in the method and, subsequently, in the hypotheses generated from unanticipated relationships.
+%%
+
+%%
+\textbf{Using Non-discretized Features Reduces Information Bias.}
+%%
+Discretization reduces the information space and introduces bias into the data and is therefore avoided in epidemiological research whenever possible.
 %%
@@
 %%
-Investigations of the the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results differing in detail. %, but differences  %differing in detail.
+Investigations of the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results but slightly differed when it comes to details.
 %%
@@
 %%
-%Our investigation regarding the hepatic steatosis data set confirms differences in results
-%%
-%\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
-%%
-
-\textbf{Attention steering is crucial.}
-%%
-Important events have to be highlighted in overview visualizations to steer the users attention to interesting parts of the data.
+
+%%
+\textbf{Attention Steering is Crucial.}
+%%
+Important events have to be highlighted in overview visualizations to direct the user's attention to interesting parts of the data.
 %%
@@
 %%
-%Foe example, formulas with a static target feature (e.g.
-$Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z-ray.
-%%
-
-%This allows for fast parsing through the data
-
-%In order to benefit most from the human pattern recognition, interesting events have to highlighted in order to steer the users attention to important parts of the data.
-%Using overview visualizations is associated with important 
-%\com{Bring important events to the users attention, otherwise it might be overlooked!}
-
-%\textbf{Support for different subject groups.}
-%\com{Include this?}
-%%
-%\com{Feedback from the evaluation goes here.}
-%\com{Time-aspect critical, interactive analysis requires for fast response. The method needs to be speeded up.}
-
-\section{Summary and Conclusion}
+\section{Summary and Outlook}
 %%
@@
 %%
-Model expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
+Model expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
 %%
@@
 %%
-Details on demand for each model allow for detailed assessment of regression models.
+Details on demand for each model allow for a detailed assessment of regression models.
 %%
@@
 %%
-We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
-%%
-
-We published all associated code and provide an analysis platform open to heterogenous data types.
-%%
-We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
-
-% \com{Rewrite, focus on precise definition on what was done and why it is awesome.}
-% We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
-% %%
-% \emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
-% %%
-% Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
-% %%
-% These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
-% %%
-% Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
-% %%
-% The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
-% %%
-% All code associated to this project is also open source.
-% %%
-% Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
-% %%
-%
-% As future work, we want to introduce more regression models to the data set, which model different correlation types.
-% %%
-% We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
-% %%
-%
-% The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
-% %%
-% The observed relationship are now subject of detailed statistical analyses.
-% %%
-% Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
-% %%
-% We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
-%%
-%%\paragraph{Future Work.}
-%%\com{Implementing of regression analysis on the graphics card.}
-%%
-% \begin{small}
+We also want to extend the cube to time-dependent data by expanding the difference cube approach.
+%%
+
+We published all associated code and provide a freely accessible analysis platform open to heterogenous data types.
+%%
+We want to support opening up knowledge discovery to allow a diverse group of domain experts to derive insight into their data and support the notion of open science.
+%%
+\begin{small}
 	%\acknowledgments{Omitted due to blind review}
@@
 Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern.
-The University of Greifswald is a member of the √¢¬Ä¬òCentre of Knowledge Interchange√¢¬Ä¬ô program of the Siemens AG.
-This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.}
-% \end{small}
+The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
+This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.
+This work was supported by the federal state of Saxony-Anhalt under grant number `I 60' within the Forschungscampus STIMULATE.}
+\end{small}
 \clearpage

### 2015-03-31 draft.tex
final revision

@@
 	%%
-	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
+	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
 	%%
@@
 %%
-Medical status features or lifestyle factors are often of \emph{dichotomous} (binary) type.
+Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
 %%
@@
 %%
-\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
+\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
 %%
@@
 %%
-The value is in the range $[0, 1]$, where 1 encodes a perfect fit.
+The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
 
@@
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-In 1977, Tukey already stated that data is too often only analyzed using confirmatory data analysis \cite{Tukey}.
+In 1977, Tukey already stated that data is too often analyzed solely using confirmatory data analysis \cite{Tukey}.
 %%
@@
 %%
-They use the system to find longitudinal pathways for diseases using health records.
+They use the system to find longitudinal pathways for diseases on the basis of health records.
 %%
@@
 %%
-Their proposed engine represents multiple heterogenous cohort study data sets as multiple normalized data cubes.
-%%
-Thus, data redundancy can be avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
+Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
+%%
+Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
 %%
@@
 %%
-Also, we aim at a fast large scale correlation analysis between many features to derive insight into the data.
+In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
 
@@
 %Shneiderman and colleagues also present challenges,
-The presented challenges in this application comprise the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
+The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
 %%
@@
 %%
-They incorporate linked views using scatter plots, bar charts, parallel coordinates and time plots to display and select the data.
+They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
 %%
@@
 %%
-The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter and are therefore suitable for \emph{small} epidemiological data sets.
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
+%%
+They are therefore suitable for \emph{small} epidemiological data sets.
 %%
@@
 %%
-We try to reproduce results and investigate findings presented in \cite{Niemann2014} further.
+We try to reproduce their results and investigate findings presented in \cite{Niemann2014} further.
 %%
@@
 %%
-We utilized semi-automatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
+We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
 %%
@@
 
-We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (recall \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods \cite{Klemm2014VIS}.
+We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (recall Sec.
+\ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods \cite{Klemm2014VIS}.
 %%
@@
 %%
-Expert knowledge is integrated into regression analysis by the regression formulas.
+Expert knowledge is integrated into the analysis using regression formulas.
 %%
@@
 %%
-When each regression takes about 50~ms calculation, we roughly have to wait 14~h for the calculation to complete.
+When each regression takes about 50~ms of calculation time, we roughly have to wait 14~h for the calculation to complete.
 %%
@@
 %%
-The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$, which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
+%The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$, which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
+The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$, which is the ratio between the average feature-class and feature-feature dependencies in the feature set $F$.
 %%
@@
 %%
-Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
+Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
 %%
@@
 %%
-This can be carried out either to refine the current formula based on observations, or to create a new regression cube for a difference view.
+This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
 %%
@@
 %%
-On formula input, a context modal displays all data set features as well as the available operators and their function.
+On formula input, a context panel displays all data set features as well as the available operators and their function.
 %%
@@
 %%
-We aim to minimize these problems as much as possible.
+We aim to minimize these problems.
 %%
@@
 %%
-Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
+Overlapping is still an issue, but this way greatly reduced in its effect to the readability of the visualization.
 %%
@@
 %%
-Dimension reduction using the target restricts the analysis to most important features.
+Dimension reduction using the target restricts the analysis to the most important features.
 %%
@@
 %%
-Model expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
+Modeling expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
 %%

### 2015-06-15 draft.tex
sorted reviews, minor changes

@@
 \item
- Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,uli.niemann,preim\}@ovgu.de
+ Paul Klemm, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,uli.niemann,preim\}@ovgu.de
 \item
  Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
+\item
+ Kai Lawonn is with Otto-von-Guericke University Magdeburg, Germany and Delft University of Technology, Delft, Netherlands. E-mail: lawonn@ovgu.de
 }

### 2015-06-18 draft.tex
covered metrics feedback from the reviews

@@
 %%
+\add{
+%%
+The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models.
+%%
+A small AIC value indicates a higher quality model.
+%%
+}
 $R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
@@
 %%
-The \emph{dependent} variable has to be of a continuous type.
-%%
+The \emph{dependent} variable has to be of \emph{continuous type}.
+%%
+\add{
+%%
+Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
+%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
+%%
+The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to them being not considered at all.
+%%
+}
 
@@
 
-\subsection{Abstracting Regression Results Using $R^2$}
+%\subsection{Abstracting Regression Results Using $R^2$}
+\subsection{Abstracting Regression Results}
 \begin{figure}[htb]
@@
  %%
- The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
+ The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
  %%
@@
  %%
- The $R^2$ value of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
+ The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
  }
@@
 %%
-The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
+\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
 %%
@@
 %%
-As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
+As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{and adjusted $R^2$} value are metrics allowing for this kind of assessment.
+%%
+\add{
+%%
+Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
+%%
+Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
+%%
+To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
+%%
+%The \emph{Regression Cube} can also be rendered using other descriptive features presented in \ref{sec:RegressionAnalysis} albeit our focus on $R^2$ values.
+%%
+%Domain experts with sound background on regression analysis can investigate the the cube with emphasis on specific aspects of the model.
+%%
+}
 %%
@@
 %%
+\add{
+%%
+$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
+%%
+$AIC$ values are not restricted to a normalized scale.
+%%
+They have to be normalized in order to map it onto color saturation.
+%%
+%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
+The resulting scale may be distorted by outliers derived from poor regression models.
+%%
+To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
+%%
+Outliers can be cut off to emphasize ranges of interest.
+%%
+}
 
@@
 %%
-$R^2$ values of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
+\add{The selected metric (per default set to $R^2$)} of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
 %%
@@
 
+\paragraph{Selecting and Scaling the Descriptive Regression Metric.}
+\add{
+%%
+The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
+%%
+The default selection is $R^2$.
+%%
+$AIC$ displays model quality.
+%%
+\emph{Adjusted} $R^2$ values are only available for linear regression, logistic regression results are represented via $R^2$ values in this mode.
+%%
+As they are visually distinguished using color, confusions are avoided.
+%%
+The transfer function of the color intensity (2D) and opacity (3D) can be adapted using a slider input.
+%%
+This additionally allows to filter models with desired features, such as only very high $R^2$ values.
+}
+
 \paragraph{3D Prism as Data Mini-Map.}
@@
 %%
-By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values and confidence intervals.
-%%
+By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-scores and AIC values}.
+%%
+\add{
+%%
+It also contains a scatter plot of the \emph{model residuals}, which shows the difference between the observed data points with the fitted values.
+%%
+Epidemiologists use such plots to validate models w.r.t. the model assumptions, such as homogeneity, normality, and independence \cite{GraphicModelAssessment}.
+}
 
@@
 \begin{small}
-	%\acknowledgments{Omitted due to blind review}
-\acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no.
+\acknowledgments{
+%%
+SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no.
 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania.
+%%
 Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern.
+%%
 The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
+%%
 This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics.
-This work was supported by the federal state of Saxony-Anhalt under grant number `I 60' within the Forschungscampus STIMULATE.}
+%%
+Sylvia Gla{\ss}er and Kai Lawonn are funded by the Federal Ministry of Education and Research within the Forschungs-campus STIMULATE under grant number '13GW0095A'.
+}
 \end{small}

### 2015-06-18 draft.tex
supported design choices with expert feedback

@@
 \newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
-\newcommand{\add}[1]{\textcolor{blue}{\uline{#1}}}
-\newcommand{\com}[1]{\textcolor{orange}{\uline{#1}}}
+\newcommand{\add}[1]{\textcolor{blue}{#1}}
+\newcommand{\com}[1]{\textcolor{red}{\underline{#1}}}
+\newcommand{\design}[1]{\textcolor{orange}{#1}}
 
@@
 \section{Regression Cube Analysis of Cohort Study Data}
-%%
-The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
-%%
-Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
-%%
-Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
-%%
-In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
-%%
-These target features often indicate the presence of the investigated disease.
-%%
-As described in Section \ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing these relationships.
+\com{TODO: Put into separate subsection?}
+%%
+\design{
+%%
+The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
+%%
+The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
+%%
+It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
+%%
+In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
+%%
+That is the focus of this work.
+%%
+% These target features often indicate the presence of the investigated disease.
+%%
+The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
+%%
+For explorative analysis, they preferred an approach, which highlights associations w.r.t.
+various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
+%%
+Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
+%%
+This aspect is considered later in the Section~\ref{sec:SystemDesign}.
+%%
+%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
+}
+%%
+%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
+%%
+%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
+%%
+%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
+%%
+%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
+%%
+%These target features often indicate the presence of the investigated disease.
+%%
+
+As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
 %%
@@
 %%
+% \design{
+% %%
+% Key requirements based on the epidemiological application domain on the Regression Cube are:
+% %%
+% \begin{itemize}
+% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
+% 	\item Allow for modeling assumptions regarding the target feature.
+% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
+% \end{itemize}
+% %%
+% }
 \subsection{Cube Description Using Regression Formula Notation}
@@
 %%
-\add{
-%%
-Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
+% \add{
+% %%
+% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
+% %%
+% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
+% %%
+% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
+% }
+%%
+\paragraph{2D (Slice) View.}
+%%
+Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
+%%
+We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
+%%
+This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
+%%
+\design{
+%%
+Based on expert feedback, the amount of features used to compare regression models was extended.
 %%
@@
 %%
-To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
-%%
-%The \emph{Regression Cube} can also be rendered using other descriptive features presented in \ref{sec:RegressionAnalysis} albeit our focus on $R^2$ values.
-%%
-%Domain experts with sound background on regression analysis can investigate the the cube with emphasis on specific aspects of the model.
-%%
-}
-%%
-\paragraph{2D (Slice) View.}
-%%
-Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
-%%
-We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
-%%
-This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
+}
 %%
@@
 }
+\design{
+%%
+To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
+%%
+}
 
@@
 %%
-The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
-%%
-Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
+\design{
+%%
+Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%%
+Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
+%%
+}
+%%
+%The visualization of descriptive metrics derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%%
+%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
 %%
@@
 %%
-Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
+\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
 
@@
 %%
-We designed our system to be freely accessible and easy to use.
+%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
+%%
+We designed our system to be openly accessible and easy to use.
+%%
 With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
@@
 %%
-\noindent Design choices and spaces are restricted by the underlying technologies and devices.
-%%
-The exchange of data and methods using web-based technologies offers many advantages.
+\design{
+%A key requirement was little set up time to reduce the effort on the domain expert side, as they are mostly practicing clinicians with small time-frames for epidemiological research.
+The design choice of using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
+}
+%Design choices and spaces are restricted by the underlying technologies and devices.
+%%
+%Therefore, the exchange of data and methods using web-based technologies offers many advantages.
 %%
@@
 %%
-The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
-%%
-Therefore, we have to adapt our design respecting the conventions of web pages.
+%The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
+%%
+%Therefore, we have to adapt our design respecting the conventions of web pages.
 %%
@@
 %%
-The \emph{Regression Cube} design focuses on a clean interface, reducing the amount of user-interface elements as much as possible.
+The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
 %%
@@
 \emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
+%%
+
 %%

### 2015-06-19 draft.tex
added reasoning on Support Design Decisions with Expert Feedback

@@
 \section{Regression Cube Analysis of Cohort Study Data}
-\com{TODO: Put into separate subsection?}
+\design{
+This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
+}
+\subsection{Iterative Design Based on Expert Feedback}
 %%
@@
 %%
-Expert knowledge is integrated into the analysis using regression formulas.
+\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
+%Expert knowledge is integrated into the analysis using regression formulas.
 %%
@@
 %%
-Based on expert feedback, the amount of features used to compare regression models was extended.
+Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
 %%
@@
 %%
-Input formulas reflect \emph{hypotheses} about the data.
+\design{
+%%
+Early analysis sessions yielded two approaches on analyzing the data.
+%%
+The classic approach is the \emph{hypothesis-based} approach, where the expert already has a good understanding of the data and potential associations (e.g.
+reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
+%%
+The \emph{hypothesis-free} approach is well suited for deriving new insight into the data, such as identifying cofounding features or potential targets (e.g.
+deriving risk factors for breast cancer associated features).
+%%
+}
+%%
+%Input formulas reflect \emph{hypotheses} about the data.
+\emph{Hypotheses} about the data are reflected using input formulas.
 %%
@@
 \paragraph{Cube Comparison.}
+\design{
+%%
+Cube comparisons were introduced later in the project, as epidemiologists with focus on statistics pointed out, that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
+%%
+}
 %%
@@
 %%
-For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
+For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).
 
@@
 The design choice of using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
+%%
+Since they are usually practicing clinicians, they usually have little time to wrangle software.
+%%
+A web-based approach has no set-up time besides loading up the data set and can be carried out with any computer connected with the web.
+%%
+Feedback of domain experts even can be implemented directly in analysis sessions and just requires a page reload.
+%%
+By providing a service a website it has a much larger chance of being tested and potentially adapted by a wide range of experts.
+%%
 }
@@
 %%
-There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
+%There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
 %%
@@
 %%
-The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
+\design{
+%%
+Complex system, which take a lot of training and experience to work with are not suitable for the task of deriving overview visualizations for epidemiologists, as they have little time slots available for their research.
+%%
+If the system is not easy to learn, it will not be adapted.
+%%
+}
+%%
+%The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
+Therefore, the \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
 %%
@@
 	%%
-	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
+	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
 	%%
@@
 %%
-Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
+Popular analytics tools, such as WEKA \cite{WEKA}, owe part of their success to their support of open file types.
 %%
@@
 \paragraph{Selecting and Scaling the Descriptive Regression Metric.}
+\design{
+%%
+As the feedback made apparent that other features are of interest for analyzing regression models too, UI elements for controlling them were introduced to the system.
+%%
+}
 \add{
@@
 \paragraph{3D Prism as Data Mini-Map.}
-%%
-The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
+\design{
+%%
+In early prototype versions, the 3D prism acted as starting point for the data analysis without the implementation of a separate 2D view.
+%%
+Slices were shown using cutaway planes.
+%%
+This approach was not popular with our domain experts, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
+%%
+}
+%%
+The 3D cube representation \design{was redesigned to} act as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
 %%
 It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
+%%
+\design{The distinction between overview and details and demand using two different representations was well received with our domain experts.}
 %%

### 2015-06-19 draft.tex
covered a bunch of reviewer comments

@@
 \newcommand{\design}[1]{\textcolor{orange}{#1}}
+\newcommand{\magenta}[1]{\textcolor{magenta}{#1}}
+% white, black, red, green, blue, cyan, magenta, yellow.
 
@@
 Our contributions are:
+
+\design{
 \begin{itemize}
 	%%
-	\item An overview visualization technique representing feature interactions using target features.
-	%%
-	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
-	%%
-	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
-	%%
-	\item The analysis of confounding features by providing comparison views between different formula results.
-	%%
-	\item The expansion and adaption of our method to any data with our open web-based system.
+  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
+	%%
+  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
+	%%
+  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand.
+	%%
+  \item An openly available web-based system, which incorporates a open data format to be applicable for data of other application domains.
 	%%
 \end{itemize}
-%%
-The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
+}
+% \begin{itemize}
+% 	%%
+% 	\item An overview visualization technique representing feature interactions using target features.
+% 	%%
+% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
+% 	%%
+% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
+% 	%%
+% 	\item The analysis of confounding features by providing comparison views between different formula results.
+% 	%%
+% 	\item The expansion and adaption of our method to any data with our open web-based system.
+% 	%%
+% \end{itemize}
+%%
+\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
 %%
@@
 %%
-\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
+\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
 Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
@@
 As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
-\begin{itemize}
+\begin{enumerate}
 	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
@@
 	\item Statisticians assess the association of the derived features towards the investigated disease.
-\end{itemize}
+\end{enumerate}
 %%
@@
 \section{Regression Cube Analysis of Cohort Study Data}
-\design{
-This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
-}
+\magenta{
+The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.
+%%
+Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.
+%%
+Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as SPSS or STATA.
+}
+% \design{
+% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
+% }
 \subsection{Iterative Design Based on Expert Feedback}
@@
 %%
-The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
+The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses with other (statistical) tools.
+%%
+\magenta{This is is, however, out of the scope of this work.}
 %%
@@
 %%
-By providing a service a website it has a much larger chance of being tested and potentially adapted by a wide range of experts.
+By providing a service a website it has a much larger chance of being tested and potentially adapted by a broad user base.
 %%
@@
 The method was well received by our clinical partners, triggering detailed investigations of the findings.
+
+\magenta{
+%One limitation of the proposed method is that the regression metrics always only capture a part of the underlying model.
+%%
+%The analyst has to keep the aspects of the respective metric in mind to avoid false conclusions.
+%%
+The analysis is limited to three dynamic variables represented the cube dimensions.
+%%
+Investigating more dynamic variables can be achieved by projecting the resulting hyper-cube into a three-dimensional representation.
+%%
+This, however, increases the cognitive load and complexity of the analysis substantially and needs to be accompanied by techniques, which simplify this approach.
+%%
+Static features can be added at will using the formula input without increasing the complexity of the visualization.
+}
 

### 2015-06-20 draft.tex
covered security issues

@@
 %%
+%Concerns regarding security issues the web-based approach were raised by domain experts due to the 
 Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
 %%
-This aspect is considered later in the Section~\ref{sec:SystemDesign}.
+Technical measures taken to enable a secure transfer and storage are described later in the Section~\ref{sec:SystemDesign}.
 %%
@@
 %%
+\magenta{
+%%
+This architecture is also prone to security issues, such as the storage of confidential data, especially in the epidemiological context.
+%%
+Therefore, we have to incorporate technical measures to ensure a secure workflow.
+%%
+}
+%%
 %The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.

### 2015-06-20 draft.tex
added justification for color, size endcoding and slicing

@@
 %%
-We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
+We can apply the same mosaic plot mapping by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
 %%
@@
 \add{The selected metric (per default set to $R^2$)} of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
+%%
+\magenta{Object size is not used to encode information because it would result in a cluttered view.}
 %%
@@
 %%
-For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System}b).
+For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes the absolute regression metric values (Fig.~\ref{fig:System}b).
 %%
@@
 	We, however, still display the whole 3D object instead of cutting away information towards the slice position.
+	%%
+	\design{Early prototypes only provided this method to select a slice of interest, which was inefficient when the user was looking for a specific slice.
+	%%
+	Hence, an additional method was implemented.
+	%%
+	}
 	%%

### 2015-06-20 draft.tex
replaced all occurences of towards which sounded odd

@@
 %%
-Although there is a strong demand for overview visualizations of a whole data set towards a target feature, no suitable tool is available for epidemiological researchers.
+Although there is a strong demand for overview visualizations of a whole data set w.r.t. a target feature, no suitable tool is available for epidemiological researchers.
 
@@
 %%
-A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
-%%
-Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
+The \emph{Regression Cube}, a novel 3D visual encoding, acts as an overview of the whole data set.
+%%
+The Cube shows all combinations of two to three independent features w.r.t. a specific target disease.
+%%
+%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
+%%
+Slicing through the \emph{Regression Cube} allows for the detailed analysis of the underlying relationships.
 %%
@@
 %%
-We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
+We were able to derive new hypotheses about relations between breast fat density and breast lesions with breast cancer.
 %%
@@
 %%
-We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
+We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t.
+a specific disease or hypothesis.
 %%
@@
 	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
-	\item Statisticians assess the association of the derived features towards the investigated disease.
+	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
 \end{enumerate}
 %%
-Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
+Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
 %%
@@
 %%
-Subject data are collected either towards a specific disease or with the widest range possible.
-%%
-The latter allows the data set to be assessed towards different diseases.
+Subject data are collected either regarding a specific disease or with the widest range possible.
+%%
+The latter allows the data set to be assessed w.r.t. different diseases.
 %%
@@
 %%
-The analysis model has to be adjusted by normalizing all included features towards the confounder.
+The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
 %%
@@
 %%
-Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
+Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features and yields lower values, when features with little entropy w.r.t.
+the target feature are included.
 %Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
@@
 %%
-Their main goal is to assess the pairwise influence of independent features towards the target feature using a plot matrix displaying models as contours.
+Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
 %%
@@
 %%
-They use \emph{Pearsson's} $r$ to depict relationships towards target features and employ list views and scatter plots to visualize and rank them.
+They use \emph{Pearsson's} $r$ to depict relationships with target features and employ list views and scatter plots to visualize and rank them.
 %%
@@
 %%
-Their research agenda acts as guideline for our design, comprising perceptual tuning towards human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
+Their research agenda acts as guideline for our design, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
@@
 
-While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
+While the presented techniques were well received by the epidemiological experts, the explanatory power w.r.t.
+back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
 %%
@@
 \\\\
-The difference of the presented approach towards our previous and related work is twofold:
+The difference of the presented approach in comparison with our previous and related work is twofold:
 %%
@@
 %%
-In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
+In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power on the target feature, often indicate the presence of the investigated disease.
 %%
@@
 	%%
-	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features towards the $Cancer$ target.
+	\item $Cancer \sim X + Y + Z$ calculates all combinations of three features w.r.t. the $Cancer$ target.
 	%%
@@
 %%
-The dependency of a set of features utilizes the entropy-based information gain to measure the explanatory power towards the target feature.
+The dependency of a set of features utilizes the entropy-based information gain to measure the explanatory power w.r.t. the target feature.
 %%
@@
 %%
-For example, if the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
+For example, if the \emph{body weight} has a strong explanatory power on the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
 %%
@@
 %%
-This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which features are of interest.
+This cube represents all possible combinations of two independent features w.r.t. all features in the data set, since we do not know which features are of interest.
 %%
@@
 %%
-All combinations of three independent features towards the target are analyzed through the cube.
+All combinations of three independent features with the target are analyzed through the cube.
 
@@
 %%
-The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
-%%
-For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).
+The absolute difference in $R^2$ values with the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
+%%
+For example, it can be utilized for comparing the influence of a single feature on the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).
 
@@
 %%
-It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
+It serves as a function similar to a mini-map, guiding the attention to points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
 %%
@@
 %%
-In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
+In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate to different slices of interest.
 %%
@@
 	%%
-	We, however, still display the whole 3D object instead of cutting away information towards the slice position.
+	We, however, still display the whole 3D object instead of cutting away information down to the slice position.
 	%%
@@
 %%
-The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
+The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice, creating a optical equivalence with the 3D cube.
 %%
@@
 %%
-The breast fat data set is the basis for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
+The breast fat data set is the basis for an explorative analysis w.r.t. the influencing parameters of the parenchyma tissue of the female breast.
 %%
@@
 %%
-One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
+One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} w.r.t. \emph{spiral contraception} ($R^2$ of $0.77$).
 %%
@@
 %%
-The overall explanatory power towards the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results towards the expected result.
+The overall explanatory power on the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results w.r.t.
+the expected result.
 %%

### 2015-06-20 draft.tex
simplified large compound sentences

@@
 %%
-This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization that reveals insights into the data by highlighting relationships.
+This approach combines all \emph{independent features} with respect to a \emph{target feature}.
+%%
+It provides a visualization that reveals insights into the data by highlighting relationships.
 %%
@@
 %%
-During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then are statistically analyzed.
+During the standard workflow, physicians transform observations into hypotheses.
+%%
+The hypotheses are depicted using epidemiological features and are then statistically analyzed.
 
@@
 %%
-Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature and mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
+Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
+%%
+They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
 %%
@@
 \subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
+%%
 Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
@@
 Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
+%%
 Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
@@
 %%
-However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
+However, this should be avoided, since it reduces the information space.
+%%
+It also introduces a new information bias, as assumptions are modeled through the discretization.
 %%
@@
 %%
-\paragraph{Linear Regression for Continuous Target.} The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
+\paragraph{Linear Regression for Continuous Target.}
+%%
+The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
 %%
@@
 %%
-Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features and yields lower values, when features with little entropy w.r.t.
-the target feature are included.
+Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
+%%
+It yields lower values, when features with little entropy w.r.t. the target feature are included.
 %Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
@@
 %%
-The main difference is their focus on analyzing one complex model in detail, yielding extensive plots, while we analyze a large amount of simpler models in terms of \emph{different} features.
+The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
+%%
+We analyze a large amount of simpler models in terms of \emph{different} features.
 %%
@@
 %%
-Their research agenda acts as guideline for our design, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
+Their research agenda acts as guideline for our design.
+%%
+Namely, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
@@
 
-We defined an Interactive Visual Analysis workflow for image-centric cohort study data, by extending the feature selection step (recall Sec.
-\ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods \cite{Klemm2014VIS}.
+We defined an Interactive Visual Analysis workflow for image-centric cohort study data \cite{Klemm2014VIS}. 
+%%
+We extended the feature selection step (recall Sec.
+\ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
 %%
@@
 
-While the presented techniques were well received by the epidemiological experts, the explanatory power w.r.t.
-back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
+While the presented techniques were well received by the epidemiological experts, the explanatory power w.r.t. back pain was limited.
+%%
+It yielded an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
 %%
@@
 %%
-Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as SPSS or STATA.
+Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as \texttt{SPSS}.
 }
@@
 %%
-In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power on the target feature, often indicate the presence of the investigated disease.
+In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power on the target feature.
+%%
+This target often indicates the presence of the investigated disease.
 %%
@@
 %%
-For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
+For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model. 
+%%
+This allows to infer about the predictive quality of the independent features included in the model.
 %%
@@
 %%
-Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
+Our goal is to create an overview visualization for a data set.
+%%
+We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
 %%
@@
 %%
-The classic approach is the \emph{hypothesis-based} approach, where the expert already has a good understanding of the data and potential associations (e.g.
+The classic approach is \emph{hypothesis-based}, where the expert already knows the data and potential associations (e.g.
 reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
 %%
-The \emph{hypothesis-free} approach is well suited for deriving new insight into the data, such as identifying cofounding features or potential targets (e.g.
+The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g.
 deriving risk factors for breast cancer associated features).
@@
 %%
-Cube comparisons were introduced later in the project, as epidemiologists with focus on statistics pointed out, that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
+Cube comparisons were introduced later in the project.
+%%
+Epidemiologists with focus on statistics pointed out, that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
 %%
@@
 %%
-By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
+By design, web technology is based on a client-server architecture.
+%%
+It allows for outsourcing computationally heavy tasks on server clusters and transferring results to the client device.
 %%
@@
 %%
-Complex system, which take a lot of training and experience to work with are not suitable for the task of deriving overview visualizations for epidemiologists, as they have little time slots available for their research.
+Epidemiologists have little time slots available for their research.
+%%
+Therefore, complex systems, which take a lot of training and experience to work with are not suitable for the task of deriving overview visualizations.
 %%
@@
 %%
-Results from the statistical processors are uploaded into the visualization slice by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in progress.
+Results from the statistical processors are uploaded into the visualization slice by slice.
+%%
+This allows the assessment of the data as soon as parts of the calculations are finished while the rest is still in progress.
 
@@
 %%
-As the feedback made apparent that other features are of interest for analyzing regression models too, UI elements for controlling them were introduced to the system.
+The feedback made apparent that other features are of interest for analyzing regression models too.
+%%
+Hence, UI elements for controlling them were introduced to the system.
 %%
@@
 %%
-The 3D cube representation \design{was redesigned to} act as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
-%%
-It serves as a function similar to a mini-map, guiding the attention to points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
+The 3D cube representation \design{was redesigned to} act as an overview over the whole data set. 
+%%
+Its purpose is not to derive detailed information about data points.
+%%
+It serves as a function similar to a mini-map, guiding the attention to points of interest in the data.
+%%
+It also gives context information about adjacent data values when using the 2D mosaic plot.
 %%
@@
 %%
-3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
+3D information visualizations are often criticized for introducing occlusions and interaction problems.
+%%
+These are often not balanced out by the advantages of using the third dimension for visual mapping.
 %%
@@
 %%
-The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice, creating a optical equivalence with the 3D cube.
+The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
+%%
+It creates a optical equivalence with the 3D cube.
 %%
@@
 %%
-The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
+The 3D cube representation already showed lower $R^2$ values in general.
+%%
+However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
 %%

### 2015-06-20 draft.tex
citations are now treated as gramatically invisible

@@
 \newcommand{\add}[1]{\textcolor{blue}{#1}}
-\newcommand{\com}[1]{\textcolor{red}{\underline{#1}}}
+\newcommand{\com}[1]{\textcolor{red}{#1}}
 \newcommand{\design}[1]{\textcolor{orange}{#1}}
@@
 %%
+\com{TODO: Reviewer suggested deletion of this sentence, but starting with subsection right after the section is ugly.}
+%%
 \subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
@@
 
-In \cite{Klemm2014VIS} we described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step.
+We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
 %%
@@
 
-Medical image data is also acquired and analyzed in modern cohort studies.
+Medical image data are also acquired and analyzed in modern cohort studies.
 %%
@@
 %%
-More discussions of cohort study data types and characteristics can be found in \cite{Preim2014, Toennies2015}.
+More discussions of cohort study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
 %%
@@
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-In 1977, Tukey already stated that data is too often analyzed solely using confirmatory data analysis \cite{Tukey}.
+In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
 %%
@@
 %%
-Similar to \cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
+Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
 %%
@@
 %%
-Time-dependent epidemiological data is visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
+Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
 %%
@@
 %%
-In \cite{Niemann2015}, Niemann et al. improve the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
-%%
-We try to reproduce their results and investigate findings presented in \cite{Niemann2014} further.
+Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
+%%
+We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
 %%
@@
 %%
-The mosaic plot presented in \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
+The mosaic plot \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
 %%
@@
  \caption{
- Different workflow types of the analysis using \emph{Regression Cubes} (based on~\cite{Keim}).
+ Different workflow types of the analysis using \emph{Regression Cubes}.
  %%
@@
 \subsection{The Hepatic Steatosis Data Set}
-We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
+We employ the data set used by Niemann et al.~\cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
 %%
@@
 %%
-In \cite{Niemann2014}, the authors show different class distributions of liver fat concentrations of women and men.
+Niemann et al.~\cite{Niemann2014} show different class distributions of liver fat concentrations of women and men.
 %%
@@
 %%
-The analysis goal was reproducing results with our visual analysis framework that are in accordance to the data mining-based results presented in \cite{Niemann2014}.
+The analysis goal was reproducing results with our visual analysis framework that are in accordance to the data mining-based results presented by Niemann et al.~\cite{Niemann2014}.
 %%
@@
 %% TODO Feedback of Henry!
-Relationships not described in \cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g., \emph{aspartate aminotransferase}.
+Relationships not described by Niemann et al.~\cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g., \emph{aspartate aminotransferase}.
 %
 %%
-Due to the difference between our regression model approach and the decision tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features is not expected.
+Due to the difference between our regression model approach and the decision tree approach presented by Niemann et al.~\cite{Niemann2014}, a complete matching set of correlating features is not expected.
 %%

### 2015-06-20 draft.tex
covered orderings as well as advantages/disadvanteges of the mini map

@@
 %%
-Their research agenda acts as guideline for our design.
+Their research agenda acts as a guideline for our design.
 %%
@@
 %%
+\magenta{Variables are ordered the same way in the 2D mosaic plot and 3D cube to preserve the mental model and make them visually analogous.}
+
 \paragraph{Tackling the disadvantages of 3D information visualization.}
@@
 %%
-The hepatic steatosis data set is analyzed by \emph{UN}, a data scientists responsible for prior analysis of the data.
+The hepatic steatosis data set is analyzed by \emph{UN}, a data scientist responsible for prior analysis of the data.
 %%

### 2015-06-20 draft.tex
added limitations for CFS

@@
 %%
+\magenta{
+The tradeoff of the CFS algorithm is the potential removal of features of interest for the domain expert.
+%%
+This problem is handled in the next subsection as part of the 3D representation of the regression results.
+}
+%%
 
@@
 %%
+\magenta{
+As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
+%%
+To tackle this problem, the $z$ dimension of the cube contains all features of the data set, allowing to assess their influence on the models.
+%%
+The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
+}
+%%
 \\\\
@@
 Each line after that represents one subject (row) and its feature manifestations.
+%%
+\magenta{Using a check-box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.}
 

### 2015-06-20 draft.tex
added further feedback from Reviewer 3, fixed cube figure

@@
 %%
+\design{
+Early designs rendered the whole cube, which made it hard to assess the position of the plane.
+%%
+%Since the regression metrics are mirrored along the diagonal, we decided to use this space to support better identification of the current slice.
+Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
+}
+%%
 The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
 %%
-The values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
+The \magenta{regression metric values} are projected on this plane to provide an overlapping-free view.
+%%
+This allows for a easier to identify the current slice.
 %%

### 2015-06-20 draft.tex
added paper requrested from reviewer

@@
 \section{Prior and Related Work}
+\com{TODO: Sort!}
+\magenta{
+Sedlmair et al.~\cite{Sedlmair} present a conceptual framework of parameter space analysis.
+%%
+It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
+%%
+The abstract description allows to sort approaches to find works with similar problems to derive new design ideas and evaluate ideas.
+%%
+Based on their input parameter classification, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
+%%
+We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
+}
+
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4

### 2015-06-21 draft.tex
added missing papers

@@
 
+\magenta{
+Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups, which fit best.
+%%
+The data space is visualized using a scatter plot matrix.
+%%
+Linear models are calculated by defining dependent and independent features.
+%%
+The model view allows for assessing different models by color coding distances to the line-of-fit.
+%%
+The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
+%%
+The paper is focused on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
+}
+
+\magenta{
+M\"{u}hlbacher et al.~\cite{Muehlbacher} provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
+%%
+A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
+%%
+The visualization allows to assess different model complexities.
+%%
+Existing regression models can be validated and compared using 3D views and 2D slice-views.
+%%
+We employ the same steps by assessing single-relationships with the target and providing a overview over combinations.
+%%
+M\"{u}hlebacher et al.
+focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models using color coded regression metrics.
+%%
+We also take categorical targets and regression formula inputs into account.
+%%
+Existing regression models can be validated and compared using 3D views and 2D slice-views.
+}
+
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4

### 2015-06-21 draft.tex
references related work in the Cover Letter

@@
 
+\magenta{
+Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
+%%
+Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
+%%
+A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+%%
+Dendograms are used to select different brain regions. 
+}
+
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4

### 2015-06-21 draft.tex
restructured and shortened related work

--- 2015-06-20 draft.tex
@@
 \section{Prior and Related Work}
-\com{TODO: Sort!}
+%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
+In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
+%%
+He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
+%%
+In this section we present prior and related work trying to achieve this goal.
+%%
+\\\\
+%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
+\paragraph{Parameter Space Analysis using Regression Models.}
+%%
 \magenta{
@@
 %%
-It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
-%%
-The abstract description allows to sort approaches to find works with similar problems to derive new design ideas and evaluate ideas.
-%%
-Based on their input parameter classification, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
+%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
+%%
+The taxonomy allows to sort approaches to find works with similar problems to derive new design ideas and evaluate ideas.
+%%
+Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
 %%
@@
 
-%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
-%%
-He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
-%%
-In this section we present prior and related work trying to achieve this goal.
-%%
-\\\\
-The work of Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan}, Angelelli et al.~\cite{Angelelli} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
+\magenta{
+%% A Partition-based Framework for Building and Validating Regression Models
+The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
+%%
+They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
+%%
+%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
+%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
+%%
+%The visualization allows to assess different model complexities.
+%%
+Existing regression models can be validated and compared using 3D views and 2D slice-views.
+%%
+We employ the same steps by preprocessing features using correlations with the target and providing a overview over combinations.
+%%
+M\"{u}hlebacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
+%%
+%We also take categorical targets and regression formula inputs into account.
+%%
+}
+
 %% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
-Akin to our approach, Piringer et al.~\cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
+Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
 %%
@@
 %%
-They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
+%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
 %%
@@
 %%
-The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
-%%
-We analyze a large amount of simpler models in terms of \emph{different} features.
-%%
-They also focus on metric features, while we process categorical data as well.
-%%
-
+% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
+% %%
+% We analyze a large amount of simpler models in terms of \emph{different} features.
+% %%
+% They also focus on metric features, while we process categorical data as well.
+%%
+\magenta{
+%%
+Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
+%%
+The data space is visualized using a scatter plot matrix.
+%%
+Linear models are calculated by defining dependent and independent features.
+%%
+The model view allows for assessing different models by color coding distances to the line-of-fit.
+%%
+The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
+%%
+}
 %% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
@@
 %%
-In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
-%%
-We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
-%%
-
+%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
+%%
+%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
+%%
+Piringer et al.~\cite{Piringer} Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t.
+all combinations in the data set.
+%%
+Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
+%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
+% %%
+
+\paragraph{Visual Analysis of Epidemiological Data.}
+%%
 Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
@@
 
+\magenta{
+Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
+%%
+Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
+%%
+A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+%%
+Dendograms are used to select different brain regions. 
+}
+%%
 Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
@@
 In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
-
-\paragraph{Visual Analysis in Public Health.}
+%%
+%\paragraph{Visual Analysis in Public Health.}
 %%--
-Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data.
+% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data.
 %, e.g.
 using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
-%%
-%They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
-%%
-%We focus on the latter.
-%%
-%Shneiderman and colleagues also present challenges,
-The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
-%%
-Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
+% %%
+% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
+% %%
+% %We focus on the latter.
+% %%
+% %Shneiderman and colleagues also present challenges,
+% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
+% %%
+% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
 %%
@@
 
-In their survey on interactive information visualizations for health record data, Rind et al.~\cite{Rind} identified future work addressing the usability of such frameworks.
-%%
-%We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
-%%
-We follow their recommendation to design an open system available to everyone and applicable to various data sets.
-%%
-Schreck and Keim \cite{Schreck} present a method for analyzing social media data with focus on characterization and cause of an epidemic.
-%%
-They employ data aggregation techniques and use geographical data to map social media events, focusing on word clouds to derive insight into possible causes.
-%%
-They derive hypotheses using their high variation average message density algorithms, steering the user's focus to particular events.
+% In their survey on interactive information visualizations for health record data, Rind et al.~\cite{Rind} identified future work addressing the usability of such frameworks.
+% %%
+% %We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.
+% %%
+% We follow their recommendation to design an open system available to everyone and applicable to various data sets.
+% %%
+% Schreck and Keim \cite{Schreck} present a method for analyzing social media data with focus on characterization and cause of an epidemic.
+% %%
+% They employ data aggregation techniques and use geographical data to map social media events, focusing on word clouds to derive insight into possible causes.
+% %%
+% They derive hypotheses using their high variation average message density algorithms, steering the user's focus to particular events.
 %%
@@
 %%
-We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
+%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
 %%
@@
 %%
-
 We defined an Interactive Visual Analysis workflow for image-centric cohort study data \cite{Klemm2014VIS}. 
@@
 
-\textbf{Encoding via CSV Files.}
-%%
-Encoding variable types in CSV files is not standardized.
-%%
-However, we need to ensure the correct variable type classification and have to enforce some basic standards.
-%%
-All categorical values have to be enclosed by quotation marks.
-%%
-Continuous variables are denoted as digits without enclosing quotation marks.
-%%
-Although this seems obvious, many cohort study data sets encode categorical features using ID values that are denoted in a data dictionary.
-%%
-Variables with only two manifestations are classified as dichotomous, leading to three possible data types: numerical, categorical and categorical/dichotomous.
-%%
-Missing values are denoted by using no character at all, a whitespace, or an empty quotation mark encapsulated string.
+% \textbf{Encoding via CSV Files.}
+% %%
+% Encoding variable types in CSV files is not standardized.
+% %%
+% However, we need to ensure the correct variable type classification and have to enforce some basic standards.
+% %%
+% All categorical values have to be enclosed by quotation marks.
+% %%
+% Continuous variables are denoted as digits without enclosing quotation marks.
+% %%
+% Although this seems obvious, many cohort study data sets encode categorical features using ID values that are denoted in a data dictionary.
+% %%
+% Variables with only two manifestations are classified as dichotomous, leading to three possible data types: numerical, categorical and categorical/dichotomous.
+% %%
+% Missing values are denoted by using no character at all, a whitespace, or an empty quotation mark encapsulated string.
 

### 2015-06-21 draft.tex
restructured related work further
revised related work

@@
 %We implement this by supporting both hypothesis-based and hypothesis-free analysis.
-
-\magenta{
-Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
-%%
-Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
-%%
-A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
-%%
-Dendograms are used to select different brain regions. 
-}
 %%
@@
 %%
-Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
+%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
 %%
@@
 %%
+The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
+%%
+They are therefore suitable for \emph{small} epidemiological data sets.
+%%
+\magenta{
+Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
+%%
+Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
+%%
+A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+%%
+%Dendograms are used to select different brain regions. 
+Instead of applying several linked views, we focus on one view of abstracted data.
+}
+
 Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
 %%
-The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
-%%
-They are therefore suitable for \emph{small} epidemiological data sets.
-%%
 Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
@@
 Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
+%%
+\magenta{
+We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
+%%
+}
 %%

### 2015-06-21 draft.tex
cut several parts to shorten the paper

@@
 %%
-We designed our system to be openly accessible and easy to use.
+\noindent We designed our system to be openly accessible and easy to use.
 %%
@@
 %A key requirement was little set up time to reduce the effort on the domain expert side, as they are mostly practicing clinicians with small time-frames for epidemiological research.
-The design choice of using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
+Using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
 %%
@@
 We propose two ways to achieve this.
-\begin{itemize}
-	\item \emph{Applying the slicing metaphor from 3D volume data.}
+
+%\begin{itemize}
+	%\item 
+	\textbf{Applying the slicing metaphor from 3D volume data.}
 	%%
@@
 	%%
-	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
-	%%
-\end{itemize}
+	%\item
+
+	\textbf{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
+	%%
+%\end{itemize}
 %%
@@
 The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
-%%
-\paragraph{Front-End.}
+
+%\paragraph{Front-End.}
 %%
@@
 %%
-We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
-%%
-Open source volume rendering methods are available but do not satisfy our requirements.
-%%
-Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
-%%
-Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
-%%
-It also is the fastest solution that we tested.
-
-\paragraph{Back-End.}
+% We experimented with different ways for achieving the cube representation, including volume rendering, cube primitives for each data point and shader-based solutions.
+% %%
+% Open source volume rendering methods are available but do not satisfy our requirements.
+% %%
+% Creating a cube primitive for each data point resulted in non-interactive frame-rates for data sets larger than 30 features (creating 30$^3$ cube primitives).
+% %%
+% Therefore, we decided to use a shader-based solution by rendering the cube as a sprite-based particle system, allowing to customize color and opacity of every data point.
+% %%
+% It also is the fastest solution that we tested.
+
+%\paragraph{Back-End.}
 Two server structures serve as back-end.
@@
 %%
-The back-end functions necessary for all cube calculations are provided via an \texttt{R} package.
-%%
-It uses multi-core optimization to use all machine CPUs to speed up the calculation process.
-%%
-The server workload balances are managed by the front-end code.
-
-\paragraph{Access and Source.}
+% The back-end functions necessary for all cube calculations are provided via an \texttt{R} package.
+% %%
+% It uses multi-core optimization to use all machine CPUs to speed up the calculation process.
+% %%
+% The server workload balances are managed by the front-end code.
+
+%\paragraph{Access and Source.}
 %%

### 2015-06-21 draft.tex
changed title

@@
 
-\title{Regression Cube Analysis of Cohort Study Data}
+\title{Regression Cube Analysis of Population Study Data}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Regression Cube Analysis of Cohort Study Data}
+\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Regression Cube Analysis of Population Study Data}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 %\com{Problem}.
-Epidemiological studies comprise heterogeneous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
+Epidemiological studies comprise heterogeneous data about a subject group to define disease-specific risk factors.
 %%
@@
 %%
-We focus on cohort studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
+We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
 %%
@@
 
-Medical image data are also acquired and analyzed in modern cohort studies.
+Medical image data are also acquired and analyzed in modern population studies.
 %%
@@
 %%
-More discussions of cohort study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
+More discussions of population study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
 %%
@@
 %%
-It also influences the general body condition and thereby almost all features acquired through cohort studies.
+It also influences the general body condition and thereby almost all features acquired through population studies.
 %%
@@
 %%
-Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
+Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
 %%
@@
 %%
-We defined an Interactive Visual Analysis workflow for image-centric cohort study data \cite{Klemm2014VIS}. 
+We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
 %%
@@
 %%
-Information visualizations of cohort study features were augmented with extracted image data, yielding linked views with both image and non-image information.
+Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
 %%
@@
 %%
-\section{Regression Cube Analysis of Cohort Study Data}
+\section{Regression Cube Analysis of Population Study Data}
 \magenta{
@@
 %%
-We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
+We presented a technique for knowledge discovery in population study data sets with user-defined target features.
 %%

### 2015-06-22 draft.tex
replaced R² value appearances

@@
 %%
-As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{and adjusted $R^2$} value are metrics allowing for this kind of assessment.
+As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
 %%
@@
 %%
-Here, regression models with large $R^2$ values can be spotted fast, steering the user's attention to the respective slice.
+Here, regression models with large regression metric values can be spotted fast, steering the user's attention to the respective slice.
 
@@
 
-$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect that age has in view of the target $Cancer$ feature.
+$Cancer \sim X + Y + Z$ subtracted with the regression metric from $Cancer \sim Age$ excludes the confounding effect that age has in view of the target $Cancer$ feature.
 %%
@@
 %%
-The absolute difference in $R^2$ values with the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
+The absolute difference in the regression metric values with the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
 %%
@@
 %%
-Since all cells in the cube are represented using $R^2$ values, a difference cube is calculated by denoting the absolute difference of $R^2$ for each cell.
+Since all cells in the cube are represented using regression metric values, a difference cube is calculated by denoting the absolute difference of regression metric for each cell.
 %%
@@
 %%
-For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
+For such formulas, the 3D view encodes every data element as absolute difference between its regression metric values and the global mean along the z-axis.
 %%
@@
 %%
-Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
+Since the regression metric (e.g. $R^2$) values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
 %%

### 2015-06-22 draft.tex
minor changes

@@
 	%%
-  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand.
-	%%
-  \item An openly available web-based system, which incorporates a open data format to be applicable for data of other application domains.
+  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
+	%%
+  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
 	%%
@@
 This section covers the epidemiological workflow and requirements.
-%%
-\com{TODO: Reviewer suggested deletion of this sentence, but starting with subsection right after the section is ugly.}
 %%
@@
 Outliers can be cut off to emphasize ranges of interest.
+%%
+Small $AIC$ values indicate a good model.
+%%
+Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
 %%

### 2015-06-22 draft.tex
revised using Henrys feedback

@@
 
-An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
+%An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
+An important epidemiological tool for deriving such features are \emph{cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
 %%
@@
 
-Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
+Testing features for association with diseases using regression models is one of the most important epidemiological tools.
 %%
@@
 %%
-Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
+Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
 %%
@@
 \begin{enumerate}
-	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
+	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
+	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
 	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
@@
 %%
-The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
+%The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG.
 %%

### 2015-06-22 draft.tex
added Bernhards revision of the cover letter
renamed heat map in heatmaps

@@
 %%
-These data contain information (\emph{features}) about a subject's lifestyle, medical condition as well as medical image data.
+These data contain information (\emph{features}) about a subject's lifestyle, medical status as well as medical image data.
 %%
@@
 %%
-The mosaic plot \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
+The 2D heatmap \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
 %%
@@
 %%
-The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
+The idea emerged from analysis sessions of a previous project, which contained a 2D heatmap showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
 %%
@@
  \caption{
- (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
+ (a) Overview visualization using a 2D heatmap for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
  %%
@@
  %%
- (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
+ (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heatmap visualizations.
  %%
@@
 %%
-As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
+As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heatmap.
 %%
@@
 %%
-We can apply the same mosaic plot mapping by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
+We can apply the same 2D heatmap by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
 %%
@@
  %%
- [\textbf{4.DD}] Details about features using the 2D mosaic plot representation yield insights and hypotheses about feature relations.
+ [\textbf{4.DD}] Details about features using the 2D heatmap representation yield insights and hypotheses about feature relations.
  }
@@
 
-\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D mosaic plot of the slice.
+\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heatmap of the slice.
 
@@
 %%
-Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
+Observations in the 2D heatmap or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
 %%
@@
  %%
- (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
+ (c) 2D heatmap visualization of the selected slice for feature \emph{Pain/Discomfort}.
  %%
@@
 	%%
-	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
+	\item The \emph{cube visualization} consisting of the 2D heatmap as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
 	%%
@@
 %%
-Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the mosaic plot matrix.
+Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heatmap matrix.
 %%
@@
 %%
-It also gives context information about adjacent data values when using the 2D mosaic plot.
+It also gives context information about adjacent data values when using the 2D heatmap.
 %%
@@
 %%
-\magenta{Variables are ordered the same way in the 2D mosaic plot and 3D cube to preserve the mental model and make them visually analogous.}
+\magenta{Variables are ordered the same way in the 2D heatmap and 3D cube to preserve the mental model and make them visually analogous.}
 
@@
 %%
-\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
-%%
-The 2D mosaic visualization (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
+\paragraph{2D heatmap Plot Slice Visualization [4.DD].}
+%%
+The 2D heatmap (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
 %%
@@
 %%
-The 2D mosaic plot is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
+The 2D heatmap is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
 %%
@@
  %%
- A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D mosaic plot.
+ A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D heatmap.
  %%
@@
 %%
-Analyzing the mosaic plot [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
+Analyzing the 2D heatmap [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
 %%
@@
 %%
-These can then be analyzed further using the 2D mosaic plot for each cube slice.
+These can then be analyzed further using the 2D heatmap for each cube slice.
 %%

### 2015-06-22 draft.tex
reverted last change

@@
 %%
-The 2D heatmap \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
+The 2D heat map \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
 %%
@@
 %%
-The idea emerged from analysis sessions of a previous project, which contained a 2D heatmap showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
+The idea emerged from analysis sessions of a previous project, which contained a 2D heat map showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
 %%
@@
  \caption{
- (a) Overview visualization using a 2D heatmap for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
+ (a) Overview visualization using a 2D heat map for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
  %%
@@
  %%
- (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heatmap visualizations.
+ (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
  %%
@@
 %%
-As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heatmap.
+As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
 %%
@@
 %%
-We can apply the same 2D heatmap by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
+We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
 %%
@@
  %%
- [\textbf{4.DD}] Details about features using the 2D heatmap representation yield insights and hypotheses about feature relations.
+ [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
  }
@@
 
-\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heatmap of the slice.
+\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heat map of the slice.
 
@@
 %%
-Observations in the 2D heatmap or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
+Observations in the 2D heat map or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
 %%
@@
  %%
- (c) 2D heatmap visualization of the selected slice for feature \emph{Pain/Discomfort}.
+ (c) 2D heat map visualization of the selected slice for feature \emph{Pain/Discomfort}.
  %%
@@
 	%%
-	\item The \emph{cube visualization} consisting of the 2D heatmap as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
+	\item The \emph{cube visualization} consisting of the 2D heat map as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
 	%%
@@
 %%
-Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heatmap matrix.
+Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heat map matrix.
 %%
@@
 %%
-It also gives context information about adjacent data values when using the 2D heatmap.
+It also gives context information about adjacent data values when using the 2D heat map.
 %%
@@
 %%
-\magenta{Variables are ordered the same way in the 2D heatmap and 3D cube to preserve the mental model and make them visually analogous.}
+\magenta{Variables are ordered the same way in the 2D heat map and 3D cube to preserve the mental model and make them visually analogous.}
 
@@
 %%
-\paragraph{2D heatmap Plot Slice Visualization [4.DD].}
-%%
-The 2D heatmap (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
+\paragraph{2D Heat Map Plot Slice Visualization [4.DD].}
+%%
+The 2D heat map (Fig.~\ref{fig:System}c) shows all values below the matrix diagonal of the current slice.
 %%
@@
 %%
-The 2D heatmap is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
+The 2D heat map is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
 %%
@@
  %%
- A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D heatmap.
+ A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D heat map.
  %%
@@
 %%
-Analyzing the 2D heatmap [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
+Analyzing the 2D heat map [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
 %%
@@
 %%
-These can then be analyzed further using the 2D heatmap for each cube slice.
+These can then be analyzed further using the 2D heat map for each cube slice.
 %%

### 2015-06-22 draft.tex
replaced occurences of 'cube'

@@
 
-\title{Regression Cube Analysis of Population Study Data}
+\title{3D Regression Heat Map Analysis of Population Study Data}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: Regression Cube Analysis of Population Study Data}
+\shortauthortitle{Klemm \MakeLowercase{\textit{et al.}}: 3D Regression Heat Map Analysis of Population Study Data}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 %%
-The \emph{Regression Cube}, a novel 3D visual encoding, acts as an overview of the whole data set.
-%%
-The Cube shows all combinations of two to three independent features w.r.t. a specific target disease.
-%%
-%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
-%%
-Slicing through the \emph{Regression Cube} allows for the detailed analysis of the underlying relationships.
+The \emph{3D Regression Heat Map}, a novel 3D visual encoding, acts as an overview of the whole data set.
+%%
+It shows all combinations of two to three independent features w.r.t. a specific target disease.
+%%
+%A 3D visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{3D Regression Heat Map}.
+%%
+Slicing through the \emph{3D Regression Heat Map} allows for the detailed analysis of the underlying relationships.
 %%
@@
 %%
-We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
+We applied our \emph{3D Regression Heat Map} method to a hepatic steatosis data set to reproduce results from a data mining-driven analysis.
 %%
@@
 %%
-With the \emph{Regression Cube}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
+With the \emph{3D Regression Heat Map}, we present a visual overview of epidemiological data that allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
 } % end of abstract
@@
 %%
-\noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
-%%
-They extend 2D scatter plots to display detailed information about a regression model.
-%%
-In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
-%%
+% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
+% %%
+% They extend 2D scatter plots to display detailed information about a regression model.
+% %%
+% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
+% %%
 \section{Epidemiological Background} \label{sec:Background}
@@
 %% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
-We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
-%%
-Chan et al.~\cite{Chan} use the same term to describe an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
+%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
+%%
+Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
 %%
@@
 %%
-\section{Regression Cube Analysis of Population Study Data}
+\section{3D Regression Heat Map Analysis of Population Study Data}
 \magenta{
-The \emph{Regression Cube} is designed to provide overview visualizations to support hypothesis generation.
+The \emph{3D Regression Heat Map} is designed to provide overview visualizations to support hypothesis generation.
 %%
@@
 % \design{
-% This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
+% This section covers the design of the 3D Regression Heat Map as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
 % }
@@
 %%
-The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
+The \emph{3D Regression Heat Map} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
 %%
@@
 %%
-%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
+%The basic idea of our \emph{3D Regression Heat Map} is to provide an overview visualization of large cohort study data sets towards target features.
 %%
@@
 % %%
-% Key requirements based on the epidemiological application domain on the Regression Cube are:
+% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
 % %%
@@
 % }
-\subsection{Cube Description Using Regression Formula Notation}
+\subsection{Regression Heat Map Description Using Regression Formula Notation}
 %%
@@
 %%
-Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{Regression Cube} representation.
+Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{3D Regression Heat Map} representation.
 %%
@@
  %%
- These represent the slices in our cube.
+ These represent the slices in our \emph{3D Regression Heat Map}.
  %%
@@
 %%
-For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model.
+For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
 %%
@@
 % %%
-% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
+% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
 % %%
@@
 % %%
-% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
+% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
 % }
@@
 %%
-Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
+Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
 %%
@@
 %%
-To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
-%%
-}
-
-\paragraph{3D (Cube) View.}
-%%
-Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube}b).
-%%
-\add{The selected metric (per default set to $R^2$)} of each cube entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
+To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
+%%
+}
+
+\paragraph{3D View.}
+%%
+Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Cube}b).
+%%
+\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
 %%
@@
 %%
-%The visualization of descriptive metrics derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
 %%
@@
 %%
-Thus, the cube can be easily extended using other regression types.
-%%
-For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
+Thus, the visualization can be easily extended using other regression types.
+%%
+For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
 %%
@@
 %%
-To tackle this problem, the $z$ dimension of the cube contains all features of the data set, allowing to assess their influence on the models.
+To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
 %%
@@
 %%
-The different analysis approaches require different starting points using the \emph{Regression Cube}.
+The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.
 
@@
  \caption{
- Different workflow types of the analysis using \emph{Regression Cubes}.
+ Different workflow types of the analysis using \emph{3D Regression Heat Maps}.
  %%
@@
  %%
- [\textbf{2.SI}] The \emph{Regression Cube} is then visualized.
+ [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
  %%
@@
  %%
- The latter is used to compare regression cubes.
+ The latter is used to compare 3D Regression Heat Maps.
  %%
@@
 %%
-\noindent Our \emph{Regression Cube} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
+\noindent Our \emph{3D Regression Heat Map} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
 
 \textbf{Analyze First} [\textbf{1.AF}].
-Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
-
-\textbf{Show the Important} [\textbf{2.SI}]. The 3D cube visualization acts as an overview over the whole data set.
+Choosing an initial regression formula triggers the \emph{3D Regression Heat Map} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
+
+\textbf{Show the Important} [\textbf{2.SI}]. The 3D visualization acts as an overview over the whole data set.
 %%
@@
 %%
-This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
+This can be carried out either to refine the current formula based on observations, or to create a new \emph{3D Regression Heat Map} for a difference view.
 %%
@@
 %%
-This cube represents all possible combinations of two independent features w.r.t. all features in the data set, since we do not know which features are of interest.
+It represents all possible combinations of two independent features w.r.t. all features in the data set, since we do not know which features are of interest.
 %%
@@
 %%
-All combinations of three independent features with the target are analyzed through the cube.
+All combinations of three independent features with the target are analyzed through this \emph{3D Regression Heat Map}.
 
@@
 %%
-Therefore, the cube incorporates this interaction for all $X$, $Y$ and $Z$ values as independent features.
+Therefore, this interaction is incorporated for all $X$, $Y$ and $Z$ values as independent features.
 %%
@@
 %%
-This is achieved through cube comparison.
-%%
-
-\paragraph{Cube Comparison.}
+This is achieved through \emph{3D Regression Heat Map} comparison.
+%%
+
+\paragraph{3D Regression Heat Map Comparison.}
 \design{
 %%
-Cube comparisons were introduced later in the project.
+Comparisons were introduced later in the project.
 %%
@@
 %%
-\emph{Regression Cubes} can be compared by creating difference views.
-%%
-One cube (extracted with one formula) acts as reference.
-%%
-The absolute difference in the regression metric values with the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
+\emph{3D Regression Heat Maps} can be compared by creating difference views.
+%%
+One formula acts as reference.
+%%
+The absolute difference in the regression metric values with the second formula is calculated. %yielding the difference containing only the differences between the two formulas.
 %%
@@
  %%
- Breast fat data set loaded into the \emph{Regression Cube} prototype.
+ Breast fat data set loaded into the \emph{3D Regression Heat Map} prototype.
  %%
@@
  %%
- (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
+ (b) 3D heat map visualization, showing values above the matrix diagonal as overview.
  %%
@@
 %%
-%The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
-Therefore, the \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
+%The \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
+Therefore, the \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
 %%
@@
 	%%
-	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
-	%%
-	\item The \emph{cube visualization} consisting of the 2D heat map as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
-	%%
-	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
-	%%
-	It also allows to select a reference formula for creating a difference cube [1.AF, 3.ZF].
+	\item The \emph{File Upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
+	%%
+	\item The \emph{Regression Heat Map Visualization} consisting of the 2D heat map as well as a 3D representation of all regression models \add{with facilities to change the represented regression metric and its range} [2.SI].
+	%%
+	\item The \emph{Formula Editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
+	%%
+	It also allows to select a reference formula for creating difference models [1.AF, 3.ZF].
 	%%
@@
 %%
-Confirming a formula triggers the cube calculation, which is preceded by determining all required formulas.
+Confirming a formula triggers the \emph{Regression Heat Map} calculation, which is preceded by determining all required formulas.
 %%
@@
 %%
-\textbf{Difference Cube.}
-Adding a formula also adds it to the reference selection for a difference cube.
-%%
-Since all cells in the cube are represented using regression metric values, a difference cube is calculated by denoting the absolute difference of regression metric for each cell.
-%%
-\subsection{Regression Cube Visualization [2.SI].}
-%%
-The visualization and interaction with the \emph{Regression Cube} is the prototype core.
+\textbf{Difference Regression Heat Maps.}
+%%
+Adding a formula also adds it to the reference selection for a difference \emph{Regression Heat Map}.
+%%
+Since all cells in the heat map are represented using regression metric values, the difference is the absolute difference of regression metric for each cell.
+%%
+\subsection{3D Regression Heat Map Visualization [2.SI].}
+%%
+The visualization and interaction with the \emph{3D Regression Heat Map} is the prototype core.
 %%
@@
 %%
-The 3D cube representation \design{was redesigned to} act as an overview over the whole data set.
+The \emph{3D Regression Heat Map} representation \design{was redesigned to} act as an overview over the whole data set.
 %%
@@
 %%
-\magenta{Variables are ordered the same way in the 2D heat map and 3D cube to preserve the mental model and make them visually analogous.}
+\magenta{Variables are ordered the same way in the 2D and 3D heat map to preserve the mental model and make them visually analogous.}
 
@@
 %%
-Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
+Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse heat maps.
 %%
@@
 %% INTERACTION
-Rotating with the cube is restricted to the y-axis, preserving the mental map to position individual features.
-%%
-The cube is always oriented according to the 2D representation, allowing for an easy mental combination of the two representations.
+Rotating the 3D heat map is restricted to the y-axis, preserving the mental map to position individual features.
+%%
+The 3D heat map is always oriented according to the 2D representation, allowing for an easy mental combination of the two representations.
 %%
@@
 %%
-\paragraph{Cube Slice Selection [3.ZF].}
+\paragraph{3D Heat Map Slice Selection [3.ZF].}
 %%
@@
 	%%
-	We employ this technique for selecting cube slices (e.g., by moving a plane via vertical mouse input while pressing the right mouse button).
+	We employ this technique for selecting 3D heat map slices (e.g., by moving a plane via vertical mouse input while pressing the right mouse button).
 	%%
@@
 \design{
-Early designs rendered the whole cube, which made it hard to assess the position of the plane.
+Early designs rendered the whole \emph{3D Regression Heat Map}, which made it hard to assess the position of the plane.
 %%
 %Since the regression metrics are mirrored along the diagonal, we decided to use this space to support better identification of the current slice.
-Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
-}
-%%
-The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the currently selected plane.
+Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the heat map diagonal is used to display information about the currently selected plane.
+}
 %%
@@
 %%
-It creates a optical equivalence with the 3D cube.
+It creates a optical equivalence with the 3D heat map.
 %%
@@
 %%
-The free space above the matrix diagonal is used to display the 3D cube.
+The free space above the matrix diagonal is used to display the 3D heat map.
 %%
@@
  %%
- \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{Regression Cubes}.
+ \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{3D Regression Heat Maps}.
  %%
@@
 %%
-A running instance of the \emph{Regression Cube} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
+A running instance of the \emph{3D Regression Heat Map} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
 %%
@@
 %%
-Instructions and code on how to create a setup running the \emph{Regression Cube} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
+Instructions and code on how to create a setup running the \emph{3D Regression Heat Map} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
 %%
@@
 %%
-In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
+In this section, we describe how we applied the \emph{3D Regression Heat Map} to two epidemiological data sets.
 %%
@@
 %%
-The [\textbf{2.SI}] step using the 3D cube representations locates hotspots at the end of the cube (Fig.~\ref{fig:Application}a).
-%%
-The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the cube using the mouse input to inspect the hotspots.
+The [\textbf{2.SI}] step using the 3D heat map locates hotspots at the end of the heat map (Fig.~\ref{fig:Application}a).
+%%
+The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the 3D heat map using the mouse input to inspect the hotspots.
 %%
@@
 %%
-The 3D cube representation already showed lower $R^2$ values in general.
+The 3D heat map already showed lower $R^2$ values in general.
 %%
@@
 %%
-The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
-%%
-Using the 3D cube, \emph{KH} observed several hotspots [\textbf{2.SI}].
+The analysis was conducted by calculating the formula $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
+%%
+Using the 3D heat map, \emph{KH} observed several hotspots [\textbf{2.SI}].
 %%
@@
 %%
-We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
+We found the 3D heat map as supporting mini-map visualizations most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
 %%
@@
 %%
-A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
-%%
-These can then be analyzed further using the 2D heat map for each cube slice.
+A \emph{3D Regression Heat Map} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
+%%
+These can then be analyzed further using the 2D plot for each 3D heat map slice.
 %%
@@
 %%
-The analysis is limited to three dynamic variables represented the cube dimensions.
-%%
-Investigating more dynamic variables can be achieved by projecting the resulting hyper-cube into a three-dimensional representation.
+The analysis is limited to three dynamic variables representing the \emph{3D Regression Heat Map} dimensions.
+%%
+Investigating more dynamic variables can be achieved by projecting the resulting high-dimensional space into a three-dimensional representation.
 %%
@@
 %%
-We also want to extend the cube to time-dependent data by expanding the difference cube approach.
+We also want to extend the 3D heat map to time-dependent data by expanding the difference heat map approach.
 %%

### 2015-06-22 draft.tex
renamed cube figure

@@
  \centering
- \includegraphics[width=1.0\linewidth]{figures/cube}
+ \includegraphics[width=1.0\linewidth]{figures/heat_map}
  \caption{
- (a) Overview visualization using a 2D heat map for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
+ (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
  %%
@@
  }
-  \label{fig:Cube}
+  \label{fig:Heat_Map}
 \end{figure}
@@
 %%
-We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
+We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
 %%
@@
 %%
-Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Cube}b).
+Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Heat_Map}b).
 %%
@@
 %%
-Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heat map matrix.
+Figure~\ref{fig:Heat_Map} shows that all values are mirrored along the diagonal of the 2D heat map matrix.
 %%

### 2015-06-22 draft.tex
revised cube occurences in the cover letter; updated Section titles
added Bernhards revision of page 1

@@
 %%
-Although there is a strong demand for overview visualizations of a whole data set w.r.t. a target feature, no suitable tool is available for epidemiological researchers.
+%Although there is a strong demand for overview visualizations of a whole data set w.r.t. a target feature, no suitable tool is available for epidemiological researchers.
 
@@
 %%
-This approach combines all \emph{independent features} with respect to a \emph{target feature}.
+This approach combines all \emph{independent features} w.r.t. a \emph{target feature}.
 %%
@@
 %%
-It shows all combinations of two to three independent features w.r.t. a specific target disease.
+It shows all combinations of two to three independent features with a specific target disease.
 %%
@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{Interactive Visual Analysis, Epidemiology, Breast Cancer, Hepatic Steatosis}
+\keywords{Interactive Visual Analysis, regression, Epidemiology, Breast Cancer, Hepatic Steatosis}
 
@@
 %%
-Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
+Insights about risk factors allow to characterize disease-specific high-risk groups \cite{Fletcher2012}.
 %%
@@
 %%
-Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
-%%
-These data may be quantified based on user-defined landmarks, which describe attributes like the shape, volume or diameter of a structure.
-%%
-
-Testing features for association with diseases using regression models is one of the most important epidemiological tools.
-%%
-Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
-%%
-Due to the amount of data and only limited overview techniques, possible correlations may be missed.
+%Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
+%%
+%These data may be quantified based on user-defined landmarks, which describe attributes like the shape, volume or diameter of a structure.
+%%
+
+Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
+%%
+Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
+%%
+Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
 %%
@@
 %%
-This section covers the epidemiological workflow and requirements.
+This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
 %%
@@
 %%
-\paragraph{2D Heat Map Plot Slice Visualization [4.DD].}
+\paragraph{2D Heat Map Slice Visualization [4.DD].}
 %%

### 2015-06-23 draft.tex
second round on Bernhards revision; revised paragraphs

@@
 %%
-They indicate the per-subject chance of developing the disease.
+Derived absolute risks indicate the per-subject chance of developing the disease.
 %%
@@
 %%
-There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
+There are many data acquisition modalities available to epidemiologists. %, which are suitable for different analysis types.
 %%
@@
 %%
-Subject data are collected either regarding a specific disease or with the widest range possible.
+Subject data are collected either regarding specific diseases or with the widest range possible.
 %%
@@
 %%
-These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
+These features are gathered through different modalities, such as questionnaires, medical examinations or laboratory analyses.
 %%
@@
 %%
-It does not focus on a specific disease, making the data set open for many diseases.
+%It does not focus on a specific disease, making the data set open for many diseases.
 %%
@@
 %%
-\\\\
 %The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
-\paragraph{Parameter Space Analysis using Regression Models.}
+\paragraph{Parameter space analysis using regression models.}
 %%
 \magenta{
-Sedlmair et al.~\cite{Sedlmair} present a conceptual framework of parameter space analysis.
+Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
 %%
@@
 %%
-The taxonomy allows to sort approaches to find works with similar problems to derive new design ideas and evaluate ideas.
+Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
+%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
 %%
@@
 %%
-We employ the same steps by preprocessing features using correlations with the target and providing a overview over combinations.
+We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
 %%
@@
 %%
-Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
+Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
 %%
@@
 %%
-Piringer et al.~\cite{Piringer} Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t.
+Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t.
 all combinations in the data set.
@@
 
-\paragraph{Visual Analysis of Epidemiological Data.}
+\paragraph{Visual analysis of epidemiological data.}
 %%
@@
 %%
-They use \emph{Pearsson's} $r$ to depict relationships with target features and employ list views and scatter plots to visualize and rank them.
+They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
 %%
@@
 
-Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
 %%
@@
 %%
-\paragraph{Statistical Analysis.}
+\paragraph{Statistical analysis.}
 %%
@@
 %%
-Namely, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
+Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
@@
 %%
-The approach was applied to a cognitive aging study, where new hypotheses could be derived.
+%The approach was applied to a cognitive aging study, where new hypotheses could be derived.
 
@@
 %%
-\paragraph{Prior Work.}
+\paragraph{Prior work.}
 In our prior work, we analyzed the healthy aging process of the lumbar spine.
@@
 %%
-Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.
+Hence, it is associated with step 1 and 2 of the epidemiological workflow (recall Section~\ref{sec:EpidemiologicalWorkflow}).
 %%
@@
 %%
-That is the focus of this work.
+% That is the focus of this work.
 %%
@@
 %Concerns regarding security issues the web-based approach were raised by domain experts due to the
-Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
-%%
-Technical measures taken to enable a secure transfer and storage are described later in the Section~\ref{sec:SystemDesign}.
+Additionally, due to the sensitive nature of population study data, the data has to be handled securely.
+%%
+Technical measures to enable a secure transfer and storage are described in Section~\ref{sec:SystemDesign}.
 %%
@@
 %%
-The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
+This notation comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
 %%
@@
 %%
-For example, if the \emph{body weight} has a strong explanatory power on the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
+For example, if the \emph{body weight} has a strong explanatory power w.r.t. the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
 %%
@@
 %%
-\paragraph{2D (Slice) View.}
+\paragraph{2D (slice) view.}
 %%
@@
 
-\paragraph{3D View.}
+\paragraph{3D view.}
 %%
@@
 %%
-Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+Epidemiologists with background in statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
 %%
@@
 
-\textbf{Analyze First} [\textbf{1.AF}].
-Choosing an initial regression formula triggers the \emph{3D Regression Heat Map} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.
-
-\textbf{Show the Important} [\textbf{2.SI}]. The 3D visualization acts as an overview over the whole data set.
+\textbf{Analyze first} [\textbf{1.AF}].
+Choosing an initial regression formula triggers the \emph{3D Regression Heat Map} calculation, filtering the dimensions of the dependent feature through the CFS algorithm.
+
+\textbf{Show the important} [\textbf{2.SI}]. The 3D visualization acts as an overview over the whole data set.
 %%
@@
 
-\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heat map of the slice.
-
-\textbf{Details on Demand} [\textbf{4.DD}].
+\textbf{Zoom, filter and analyze further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D heat map of the slice.
+
+\textbf{Details on demand} [\textbf{4.DD}].
 Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
@@
 %%
-\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
+\paragraph{Hypothesis-free and hypothesis-based analysis.}
 %%
@@
 %%
-Early analysis sessions yielded two approaches on analyzing the data.
+Early analysis sessions yielded two approaches of analyzing the data.
 %%
@@
 %%
-The \emph{hypothesis-free} analysis is well suited for deriving new insight into the data, such as identifying confounding features or potential targets (e.g.
+The \emph{hypothesis-free} analysis allows to derive new insight into the data, such as identifying confounding features or potential targets (e.g.
 deriving risk factors for breast cancer associated features).
@@
 %%
-It is therefore suitable for an explorative analysis to give a general impression about relationships in the data set.
+It is therefore suitable for an exploratory analysis.
 
@@
 
-\paragraph{3D Regression Heat Map Comparison.}
+\paragraph{3D regression heat map comparison.}
 \design{
@@
 %%
-\magenta{This is is, however, out of the scope of this work.}
+\magenta{This is, however, out of the scope of this work.}
 %%
 Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
-%%
-\\\\
+
 %%
@@
 %%
-Feedback of domain experts even can be implemented directly in analysis sessions and just requires a page reload.
-%%
-By providing a service a website it has a much larger chance of being tested and potentially adapted by a broad user base.
+We can even implement small changes based on feedback of domain experts directly during analysis sessions. %and just requires a page reload.
+%%
+By providing a service using a website it has a much larger chance of being tested and potentially adapted by a broad user base.
 %%
@@
 %%
-Therefore, complex systems, which take a lot of training and experience to work with are not suitable for the task of deriving overview visualizations.
-%%
-If the system is not easy to learn, it will not be adapted.
-%%
-}
-%%
-%The \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
-Therefore, the \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
+Therefore, complex systems, which require substantial training to work with are not suitable. %for the task of deriving overview visualizations.
+%%
+%If the system is not easy to learn, it will not be adapted.
+%%
+}
+%%
+%The \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user interface elements as much as possible.
+Therefore, the \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user interface elements as much as possible.
 %%
@@
 %%
-\paragraph{File Upload and Classification [1.AF].}
+\paragraph{File upload and classification [\textbf{1.AF}].}
 %%
@@
 
-\textbf{Data Security.}
-%%
-Security issues are raised by uploading data into an online service such as our prototype.
+\textbf{Data security} issues are raised by uploading data into an online service such as our prototype.
 %%
@@
 
-\paragraph{Formula Editor [1.AF, 3.ZF].}
+\paragraph{Formula editor [\textbf{1.AF}, \textbf{3.ZF}].}
 %%
@@
 
-\textbf{Formula Validation and Calculation.} The formula is checked for validity directly on input.
+\textbf{Formula validation} is carried out directly on input.
 %%
@@
 %%
-Confirming a formula triggers the \emph{Regression Heat Map} calculation, which is preceded by determining all required formulas.
-%%
-These are then divided by the number of available statistical back-end processors, driving a \emph{cloud-computing}-based approach.
-%%
-In theory, the calculation duration is reduced by a factor of $0.5$ by every statistical processor.
+Confirming a formula triggers the \emph{Regression Heat Map} \textbf{calculation}, which is preceded by determining all required formulas.
+%%
+These are then divided by the number of available statistical back-end processors, driving a \emph{cloud computing}-based approach.
+%%
+In theory, the calculation duration is reduced by a factor of $2$ by every statistical processor.
 %%
@@
 %%
-\textbf{Difference Regression Heat Maps.}
-%%
-Adding a formula also adds it to the reference selection for a difference \emph{Regression Heat Map}.
+\textbf{Difference heat maps} can be generated for each formula added to the system. %also adds it to the reference selection for a difference \emph{Regression Heat Map}.
+Using a dropdown menu it can be selected as reference.
 %%
@@
 
-\paragraph{Usage of a Regression Prism for Information Reduction.}
+\paragraph{Usage of a regression prism for information reduction.}
 %%
@@
 
-\paragraph{Selecting and Scaling the Descriptive Regression Metric.}
+\paragraph{Selecting and scaling the descriptive regression metric.}
 \design{
@@
 
-\paragraph{3D Prism as Data Mini-Map.}
+\paragraph{3D prism as data mini-map.}
 \design{
@@
 %%
-\paragraph{3D Heat Map Slice Selection [3.ZF].}
+\paragraph{3D heat map slice selection [\textbf{3.ZF}].}
 %%
@@
 	%\item
-	\textbf{Applying the slicing metaphor from 3D volume data.}
+	\textbf{We apply the slicing metaphor from 3D volume data.}
 	%%
@@
 %%
-\paragraph{2D Heat Map Slice Visualization [4.DD].}
+\paragraph{2D heat map slice visualization [\textbf{4.DD}].}
 %%
@@
 %%
-\paragraph{Participants, Setup and Procedure.}
+\paragraph{Participants, setup and procedure.}
 %%
@@
 %%
-\textbf{Analysis of Non-discretized Target Feature.}
+\paragraph{Analysis of non-discretized target feature.}
 %%
@@
 
-\textbf{Interleukin-6 Correlation With Liver Fat.}
+\paragraph{Interleukin-6 correlation with liver fat.}
 %%
@@
 %%
-\textbf{Detailed Breast Parenchyma Analysis.}
+\paragraph{Detailed breast parenchyma analysis.}
 %%
@@
 %%
-\textbf{Extracted Hypotheses Have to be Investigated Further.}
+\paragraph{Extracted hypotheses have to be investigated further.}
 %%
@@
 %%
-\textbf{Overview Visualizations are Preferred over Black-box Methods.}
+\paragraph{Overview visualizations are preferred over black-box methods.}
 %%
@@
 %%
-\textbf{Using Non-discretized Features Reduces Information Bias.}
+\paragraph{Using non-discretized features reduces information bias.}
 %%
@@
 %%
-\textbf{Attention Steering is Crucial.}
+\paragraph{Attention steering is crucial.}
 %%

### 2015-06-23 draft.tex
added Bernhards revision

@@
 %%
-This approach was not popular with our domain experts, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
+This approach was not popular among epidemiologists, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
 %%
@@
 %%
-Its purpose is not to derive detailed information about data points.
+%Its purpose is not to derive detailed information about data points.
 %%
@@
 %%
-\design{The distinction between overview and details and demand using two different representations was well received with our domain experts.}
+\design{The distinction between overview and details on demand using two different representations was well received with our domain experts.}
 %%
@@
 %%
-It also creates a sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
+The visualization is sparse, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
 %%
@@
 %%
-Overlapping is still an issue, but this way greatly reduced in its effect to the readability of the visualization.
+Overlapping is still an issue, but greatly reduced in its effect to the readability of the visualization.
 %%
@@
 %% INTERACTION
-Rotating the 3D heat map is restricted to the y-axis, preserving the mental map to position individual features.
-%%
-The 3D heat map is always oriented according to the 2D representation, allowing for an easy mental combination of the two representations.
+Transformation of the 3D heat map is restricted to the y-axis (horizontal only), preserving the mental map to position individual features.
+%%
+The 3D heat map is always oriented according to the 2D representation, allowing for an easy mental combination of them.
 %%
@@
 	%%
-	In medical volume data renderings, slicing views are very common to view details on a selected plane in the scene.
+	In medical volume renderings, slicing views are common to view details on a selected plane in the scene.
 	%%
@@
 	%%
-	We, however, still display the whole 3D object instead of cutting away information down to the slice position.
+	We, however, still display the whole 3D object instead of cutting away information. %down to the slice position.
 	%%
@@
 
-	\textbf{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
+	\textbf{Selecting the slice using a dropdown menu} containing the feature names provides fast access to plane selections when the user already knows the slices of interest.
 	%%
@@
 %%
-The currently selected slice is denoted with a semi-transparent gray plane.
+The currently selected slice is displayed as a semi-transparent gray plane.
 %%
 \design{
-Early designs rendered the whole \emph{3D Regression Heat Map}, which made it hard to assess the position of the plane.
+Early prototypes rendered the whole \emph{3D Regression Heat Map}, which made it hard to assess the position of the plane.
 %%
@@
 %%
-It creates a optical equivalence with the 3D heat map.
+It creates an optical equivalence with the 3D heat map.
 %%
@@
 %%
-The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
-
+%The ongoing transition of open-science software into the web spawned numerous projects, making state-of-the-art algorithms available in this domain.
+%%
 %\paragraph{Front-End.}
@@
 %%
-It forces developers to write modularized code, which makes the components easier expandable while keeping the code maintainable by including unit tests.
-%%
-The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements with proper stylings.
+In is easily expandable by forcing developers to write modularized code.
+%It forces developers to write modularized code, which makes the components easier expandable. %while keeping the code maintainable by including unit tests.
+%%
+The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements.
+%with proper stylings.
 %%
@@
 %%
-The first one is the web-server, which is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
+The web-server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
 %%
@@
 
-The statistical processors yield the second structure.
+The statistical computations are performed on the second structure.
 %%
@@
 %%
-The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-cube-r-package}{\\\texttt{github.com/paulklemm/regression-cube-r-package}}}$^{,}$\footnote{Front-End and Node.js Webserver: \href{https://github.com/paulklemm/regression-cube-prototype}{\texttt{\\github.com/paulklemm/regression-cube-prototype}}}
+The source for the prototype is freely available at \texttt{Github}.\footnote{\texttt{R}-based back-end: \href{https://github.com/paulklemm/regression-heatmap-r-package}{\\\texttt{github.com/paulklemm/regression-heatmap-r-package}}}$^{,}$\footnote{Front-End and Node.js Webserver: \href{https://github.com/paulklemm/regression-heatmap-prototype}{\texttt{\\github.com/paulklemm/regression-heatmap-prototype}}}
 %%
@@
 %%
-In this section, we describe how we applied the \emph{3D Regression Heat Map} to two epidemiological data sets.
+In this section, we describe the application of the \emph{3D Regression Heat Map} to two epidemiological data sets.
 %%
@@
 %%
-\emph{VDAR} can be carried out using case studies using thinking-aloud techniques to comprehend the reasoning and thought process of the user.
+\emph{VDAR} can be carried out based on case studies using thinking-aloud techniques to comprehend the reasoning and thought process of the user.
 %%
@@
 %%
-\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set analysis.
+\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set.
 %%
@@
 %%
-The data set contains labels for 578 participants for which the MRI data was available to that point.
-%%
-The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present the last study wave.
-
-%%
-Apart from the target feature, the data set contains 199 features, comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol or tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
+The data set contains labels for 578 participants. %for which the MRI data was available to that point.
+%%
+The MRI scans for each subject are only available in SHIP-2. %, so the target feature is only present the last study wave.
+
+%%
+Apart from the target feature, the data set contains 199 features, comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol and tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
 %%
@@
 %%
-The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
+Breast density is denoted as the ratio between parenchyma and cellular connective tissue and has been shown to be associated with breast cancer.
 %%
@@
 
-The data comprises 1,186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
-%%
-It contains 231 dimensions, holding information about somatometric features (e.g., body size or weight) consumption behavior (e.g., alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women-specific features (e.g., number of born children or contraception type) as well as mammography features (e.g., fat content or parenchyma tissue proportion to volume).
+The data comprises 1,186 female subjects. %(368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
+%%
+It contains 231 dimensions, holding information about somatometric features (e.g., body size and weight) consumption behavior, personal and medical history (e.g., occupation and prior diseases), women-specific features (e.g., number of born children and contraception type) as well as mammography features (e.g., fat content and parenchyma tissue proportion to volume).
 %%
@@
 %%
-The [\textbf{2.SI}] step using the 3D heat map locates hotspots at the end of the heat map (Fig.~\ref{fig:Application}a).
+The [\textbf{2.SI}] step using the 3D heat map locates hotspots at the end of the heat map (Fig.~\ref{fig:Application} left).
 %%
@@
 %%
-The 3D heat map already showed lower $R^2$ values in general.
-%%
-However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
+The 3D heat map showed lower $R^2$ values in general.
+%%
+However, the analysis is now based on linear regression and the $R^2$ values cannot be compared directly.
 %%
@@
 %%
-Relationships between the share parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
+Relationships between the share of parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
 %%
@@
 %%
-We map results of complex statistical results into comprehensive visualizations.
+We map results of complex statistical computations into comprehensive visualizations.
 %%
@@
 %%
-Statistical validation of an epidemiological end result still has to be carried out by statisticians using their respective tools.
+Statistical validation of an epidemiological result still has to be carried out by statisticians using their respective tools.
 %%
@@
 %%
-Investigations of the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results but slightly differed when it comes to details.
-%%
-The overall explanatory power on the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results w.r.t.
-the expected result.
+Investigations of the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results. %but slightly differed when it comes to details.
+%%
+The overall explanatory power on the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the data w.r.t. the expected result.
 %%
@@
 %%
-We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
+We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast fat data set.
 %%
@@
 %%
-Static features can be added at will using the formula input without increasing the complexity of the visualization.
-}
-
-As next step, we want to introduce more regression types, which model different kinds of correlations.
+Static features can be added using the formula input without increasing the complexity of the visualization.
+}
+
+As a next step, we want to introduce more regression types, which model different kinds of correlations.
 %%
@@
 %%
-We want to support opening up knowledge discovery to allow a diverse group of domain experts to derive insight into their data and support the notion of open science.
+We want to support opening up knowledge discovery to allow a diverse group of domain experts to derive insight into their data. %and support the notion of open science.
 %%

### 2015-06-23 draft.tex
shortened the paper

@@
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
-%%
-He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
+Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
+%%
+He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
 %%
@@
 \\\\
-The difference of the presented approach in comparison with our previous and related work is twofold:
-%%
-\begin{enumerate}
-	%%
-	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
-	%%
-	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
-	%%
-\end{enumerate}
+The difference of the presented approach in comparison with our previous and related work is twofold.
+%%
+(1) We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+%%
+(2) The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
+%%
+% \begin{enumerate}
+% 	%%
+% 	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+% 	%%
+% 	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
+% 	%%
+% \end{enumerate}
 %%
@@
 %%
-In a data set with $n$ (e.g., 100) features, the regression formula
-\begin{equation}
-Cancer \sim X + Y
-\end{equation}
+In a data set with $n$ (e.g., 100) features, the regression formula $Cancer \sim X + Y$
+%\begin{equation}
+%
+%\end{equation}
 would yield $n^2$ (10,000) regression models, describing all possible combinations of two features in the data describing the target feature $Cancer$.
@@
 %%
-Also we provide a zoom functionality using the mouse wheel input.
+%Also we provide a zoom functionality using the mouse wheel input.
 %%
@@
  %%
- The web-server back-end (right) is written using \emph{Node.js}, hosting is carried out using \emph{Heroku}.
+ The web-server (right) is written using \emph{Node.js} and hosted on \emph{Heroku}.
  %%
@@
 %%
-\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+%\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+\texttt{Angular.js} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
 %%
@@
 %%
-The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements.
-%with proper stylings.
-%%
-The 2D heat map is implemented using \texttt{Data driven Documents} (\texttt{D3.js}), which is popular in information visualization using vector graphics \cite{D3}.
+%The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements.
+\texttt{Twitter Bootstrap} handles the page layout and provides a rich set of user interface elements.
+%%
+The 2D heat map is implemented using the \texttt{D3.js}~\cite{D3} information visualization library.
 %%
@@
 %%
-The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
+%The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
+The 3D plot is created using the WebGL-based library.
 %%
@@
 %%
-The web-server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
-%%
-It is hosted on \texttt{Heroku}\footnote{Owned by Salesforce.com, \href{https://www.heroku.com/}{\texttt{heroku.com}}}, a cloud application platform.
+%The web-server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
+The web-server is written in \texttt{Javascript} using \texttt{Node.js}, running on Googles V8 Javascript runtime environment.
+%%
+%It is hosted on \texttt{Heroku}\footnote{Owned by Salesforce.com, \href{https://www.heroku.com/}{\texttt{heroku.com}}}, a cloud application platform.
+It is hosted on \texttt{Heroku}, a cloud application platform.
 %%
@@
 %%
-They rely on the statistical programming language \texttt{R}.\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}
+%They rely on the statistical programming language \texttt{R}.\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}
+They rely on the statistical programming language \texttt{R}.
 %%
@@
 %%
-The ability of a system to discover knowledge is difficult to measure.
-%%
-Lam et al.~\cite{Lam2012} propose the \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique, which is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
-%%
-\emph{VDAR} can be carried out based on case studies using thinking-aloud techniques to comprehend the reasoning and thought process of the user.
+The knowledge discovery capabilities of a system are difficult to measure.
+%%
+The \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique proposed by Lam et al.~\cite{Lam2012} is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
+%%
+\emph{VDAR} can be carried out based on case studies using thinking-aloud techniques to comprehend the users reasoning and thought process.
 %%
@@
 Starting an analysis using these techniques takes about 5-10 minutes of setup time.
-
+%%
 The sessions started with an initial overview of the system, showcasing its features and functionality.
@@
 All sessions were video recorded to be processed later on.
-%%
-
 %%
@@
 %%
-
 We published all associated code and provide a freely accessible analysis platform open to heterogenous data types.

### 2015-06-23 draft.tex
revised figures to save space, removed workflow figure

@@
 %%
-\begin{figure}[htb]
- \centering
- \includegraphics[width=1.0\linewidth]{figures/workflow}
- \caption{
- Different workflow types of the analysis using \emph{3D Regression Heat Maps}.
- %%
- [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
- %%
- [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
- %%
- The user has the option to either adjust the formula or to derive details on demand on a specific regression.
- %%
- [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
- %%
- The latter is used to compare 3D Regression Heat Maps.
- %%
- [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
- }
-  \label{fig:Workflow}
-\end{figure}
+% \begin{figure}[htb]
+%  \centering
+%  \includegraphics[width=1.0\linewidth]{figures/workflow}
+%  \caption{
+%  Different workflow types using the \emph{3D Regression Heat Map}.
+%  %%
+%  [\textbf{1.AF}] The workflow starts by declaring a formula to specify a hypothesis, or to use a predefined formula for a hypothesis-free analysis.
+%  %%
+%  [\textbf{2.SI}] The \emph{3D Regression Heat Map} is then visualized.
+%  %%
+%  The user has the option to either adjust the formula, adjust the transfer function or to derive details on demand on models.
+%  %%
+%  [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula or a selection of a difference view.
+%  %%
+%  The latter is used to compare \emph{3D Regression Heat Maps}.
+%  %%
+%  [\textbf{4.DD}] Details about features using the 2D heat map representation yield insights and hypotheses about feature relations.
+%  }
+%   \label{fig:Workflow}
+% \end{figure}
 %%
@@
 %%
-As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
+%As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
+The workflow is highly iterative.
 %%
@@
  %%
- Breast fat data set loaded into the \emph{3D Regression Heat Map} prototype.
- %%
- (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
+ Breast fat data set loaded into our prototype.
+ %%
+ (a) Using the formula input, the user specifies the dependent feature and calculation rules.
  %%

### 2015-06-23 draft.tex
further cuts to fit the length specifications

@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{Interactive Visual Analysis, regression, Epidemiology, Breast Cancer, Hepatic Steatosis}
+\keywords{Interactive Visual Analysis, Regression Analysis, Heat Map, Epidemiology, Breast Cancer, Hepatic Steatosis}
 
@@
 %%
-This is achieved by providing template regression formulas, which are applied to all potential feature combinations.
+Therefore we provide template regression formulas, which are applied to all potential feature combinations.
 %%
@@
 Difference views between regression formulas allow to assess the influences of individual features on the process.
-
+%%
 Our contributions are:
@@
 %%
-\\\\
+
 Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
@@
 %%
-Other confounders have to be selected by epidemiologists specific to the investigated condition.
+Confounders have to be selected by epidemiologists specific to the investigated condition.
+%%
+\paragraph{The Study of Health in Pomerania (SHIP).}
+The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
+%%
+%It does not focus on a specific disease, making the data set open for many diseases.
+%%
+Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
+%%
+A second cohort, \texttt{SHIP-TREND} was started in 2012.
+%%
+Data for both cohorts are examined in a 5-year time span.
+%%
+New parameters are added in each iteration, extending the range of investigated diseases.
+%%
+%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
+%%
+Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
 
@@
 The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
+%%
 
@@
 %%
-\begin{equation}
-Dependent \sim Independent_1 + ... + Independent_n
-\label{eq:RegressionNotation}
-\end{equation}
-%%
-An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
+$Dependent \sim Independent_1 + ... + Independent_n$.
+% \begin{equation}
+% Dependent \sim Independent_1 + ... + Independent_n
+% \label{eq:RegressionNotation}
+% \end{equation}
+%%
+An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
 The most commonly used regression operators comprise:
@@
 %%
-\subsection{The Study of Health in Pomerania (SHIP)}
-The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
-%%
-%It does not focus on a specific disease, making the data set open for many diseases.
-%%
-Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
-%%
-A second cohort, \texttt{SHIP-TREND} was started in 2012.
-%%
-Data for both cohorts are examined in a 5-year time span.
-%%
-New parameters are added in each iteration, extending the range of investigated diseases.
-%%
-For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
-%%
-However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
-%%
 \section{Prior and Related Work}
@@
 %%
-Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
+%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
 %The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
@@
 %%
-We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
-%%
-M\"{u}hlebacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
+%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
+%%
+M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
 %%
@@
 %%
-Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
+%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
 %%
@@
 %%
-We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
+We apply their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}.
+%, meaning that the system has to adapt to different requirements in different analysis steps.
 %%
@@
 %%
-Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
+%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
 %%
@@
 %%
-In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
+In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insights.
 %%
@@
 %%
-
-While the presented techniques were well received by the epidemiological experts, the explanatory power w.r.t. back pain was limited.
+The presented techniques were well received by the epidemiological experts, but the explanatory power w.r.t. back pain was limited.
 %%
@@
  %%
- The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
- %%
- (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
+ The $R^2$ metrics extracted from the regression model are mapped to color saturation (a saturated color indicates a strong correlation).
+ %%
+ (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat maps.
  %%
@@
 \magenta{
-As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
-%%
-To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
+As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features important to the domain expert.
+%%
+To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence.
 %%
@@
 
-\paragraph{3D regression heat map comparison.}
+\paragraph{3D Regression Heat Map comparison.}
 \design{
@@
  %%
- (b) 3D heat map visualization, showing values above the matrix diagonal as overview.
+ (b) 3D heat map, showing values above the matrix diagonal as overview.
  %%
@@
  %%
- (c) 2D heat map visualization of the selected slice for feature \emph{Pain/Discomfort}.
+ (c) 2D heat map of the selected slice for feature \emph{Pain/Discomfort}.
  %%
@@
 %%
-By design, web technology is based on a client-server architecture.
+Web technology is based on a client-server architecture.
 %%
@@
 %%
-This allows for a fast learning of the individual system parts.
+This allows for a fast learning of the system.
 %%
@@
 	%%
-	\item The \emph{File Upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
-	%%
-	\item The \emph{Regression Heat Map Visualization} consisting of the 2D heat map as well as a 3D representation of all regression models \add{with facilities to change the represented regression metric and its range} [2.SI].
-	%%
-	\item The \emph{Formula Editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
+	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
+	%%
+	\item The \emph{Regression Heat Map visualization} consisting of the 2D heat map as well as a 3D representation of all regression models \add{with facilities to change the represented regression metric and its range} [2.SI].
+	%%
+	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
 	%%
@@
 %%
-The visualization and interaction with the \emph{3D Regression Heat Map} is the prototype core.
+The visualization and interaction with the \emph{3D Regression Heat Map} is the core of the prototype.
 %%
@@
 %%
-3D information visualizations are often criticized for introducing occlusions and interaction problems.
+3D information visualizations are criticized for introducing occlusions and interaction problems.
 %%
@@
 %%
-Since the regression metric (e.g. $R^2$) values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
+The regression metric (e.g. $R^2$) values are mapped on data point opacity, highlighting large values in the prism, which guides the focus to the respective slices.
 %%
@@
 %%
-Overlapping is still an issue, but greatly reduced in its effect to the readability of the visualization.
+Overlapping is still an issue, but greatly reduced in its effect to the visualization readability.
 %%
@@
  \caption{
- The front-end (left) is realized with \emph{HTML5}/\emph{CSS3}/\emph{Javascript} and different Javascript libraries, such as \emph{Angular.js}, \emph{Three.js} and D3.js.
- %%
- The web-server (right) is written using \emph{Node.js} and hosted on \emph{Heroku}.
- %%
- \emph{R} and \emph{OpenCPU} constitute the statistical back-end (top) to compute the \emph{3D Regression Heat Maps}.
+ The front-end (left) is realized with \texttt{HTML5}/\texttt{CSS3}/\texttt{Javascript} and different Javascript libraries, such as \href{https://www.angularjs.org/}{\texttt{Angular.js}}, \href{http://threejs.org}{\texttt{Threejs}} and \href{http://d3js.org/}{\texttt{D3.js}}.
+ %%
+ The web-server (right) is written using \href{http://nodejs.org}{\texttt{Node.js}} and hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}.
+ %%
+ \href{http://r-project.org}{\texttt{R}} and \href{https://www.opencpu.org/}{\texttt{OpenCPU}} constitute the statistical back-end (top) to compute the \emph{3D Regression Heat Maps}.
  %%
@@
 %%
-We rely on web-based technologies for our prototype.
+\noindent We rely on web-based technologies for our prototype.
 %%
@@
 %\texttt{Angular.js}\footnote{Open Source; Maintained by Google, \href{https://www.angularjs.org/}{\texttt{angularjs.org}}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
-\texttt{Angular.js} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
+\href{https://www.angularjs.org/}{\texttt{Angular.js}} abstracts web application into models and views, allowing for a responsive way to combine \texttt{HTML} and \texttt{Javascript}.
 %%
@@
 %The page layout is handled using \texttt{Twitter Bootstrap}\footnote{Open Source; Maintained by Twitter, \href{http://getbootstrap.com}{\texttt{getbootstrap.com}}}, which also provides a rich set of user interface elements.
-\texttt{Twitter Bootstrap} handles the page layout and provides a rich set of user interface elements.
-%%
-The 2D heat map is implemented using the \texttt{D3.js}~\cite{D3} information visualization library.
+\href{http://getbootstrap.com}{\texttt{Twitter Bootstrap}} handles the page layout and provides a rich set of user interface elements.
+%%
+The 2D heat map is implemented using the \href{http://d3js.org/}{\texttt{D3.js}}~\cite{D3} information visualization library.
 %%
@@
 %The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
-The 3D plot is created using the WebGL-based library.
+The 3D plot is created using the WebGL-based \href{http://threejs.org}{\texttt{Threejs}} library.
 %%
@@
 %The web-server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
-The web-server is written in \texttt{Javascript} using \texttt{Node.js}, running on Googles V8 Javascript runtime environment.
+The web-server is written in \texttt{Javascript} using \href{http://nodejs.org}{\texttt{Node.js}}, running on Googles V8 Javascript runtime environment.
 %%
 %It is hosted on \texttt{Heroku}\footnote{Owned by Salesforce.com, \href{https://www.heroku.com/}{\texttt{heroku.com}}}, a cloud application platform.
-It is hosted on \texttt{Heroku}, a cloud application platform.
-%%
-
+It is hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}, a cloud application platform.
+%%
 The statistical computations are performed on the second structure.
@@
 %They rely on the statistical programming language \texttt{R}.\footnote{Open Source; \href{http://r-project.org}{\texttt{r-project.org}}}
-They rely on the statistical programming language \texttt{R}.
+They rely on the statistical programming language \href{http://r-project.org}{\texttt{R}}.
 %%
@@
 %%
-\texttt{OpenCPU} is an \texttt{R} package and provides an API for accessing it via HTTP calls \cite{Ooms}.
-%%
-This way, any computer, which runs \texttt{R} can be turned into a statistical processor for our project.
+\href{https://www.opencpu.org/}{\texttt{OpenCPU}} is an \href{http://r-project.org}{\texttt{R}} package and provides an API for accessing it via HTTP calls \cite{Ooms}.
+%%
+This way, any computer, which runs \href{http://r-project.org}{\texttt{R}} can be turned into a statistical processor for our project.
 %%
@@
 %%
-A running instance of the \emph{3D Regression Heat Map} prototype can be found under \href{http://regressioncube.herokuapp.com/}{\texttt{regressioncube.herokuapp.com}}.
+A running instance of the \emph{3D Regression Heat Map} prototype can be found under \href{http://regressionheatmap.herokuapp.com/}{\texttt{regressionheatmap.herokuapp.com}}.
 %%
@@
 %%
-Instructions and code on how to create a setup running the \emph{3D Regression Heat Map} statistical back-end through a \texttt{Ubuntu} server using \texttt{OpenCPU} are referenced in the repository.
-%%
-The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{www.heroku.com}} by cloning the repository into a Heroku app.
+Instructions and code on how to create a setup running the statistical back-end through a \href{http://www.ubuntu.com/}{\texttt{Ubuntu}} server using \href{https://www.opencpu.org/}{\texttt{OpenCPU}} are referenced in the repository.
+%%
+The front-end can be deployed using \href{https://www.heroku.com/}{\texttt{Heroku}} by cloning the repository into a Heroku app.
 %%
@@
 %%
-
-%%
 Niemann et al.~\cite{Niemann2014} show different class distributions of liver fat concentrations of women and men.
@@
 %%
-The latter were derived from MRI data for each subject, which was manually segmented by radiologists.
+The latter were derived from MRI data for each subject, which was manually segmented by radiologists \cite{Hegenscheid2009, Ivanovska2014}.
 %%
@@
 %%
-\emph{UN} could reproduce most results, some features exhibit lower correlations though, e.g., \emph{creatinine} magnitudes.
+\emph{UN} could reproduce most results.
+%%
+Some features exhibit lower correlations, e.g., \emph{creatinine} magnitudes.
 %%
 A slight influence of \emph{age} on the target feature could be observed for women ($R^2$ of $0.09$ for females compared to $0.02$ for males).
-%% TODO Feedback of Henry!
+%%
 Relationships not described by Niemann et al.~\cite{Niemann2014} were found, such as enzymes indicating liver dysfunctions, e.g., \emph{aspartate aminotransferase}.
@@
 \emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
-%%
-
 %%

### 2015-06-24 draft.tex
added Bernhards second revision; added Bibs

@@
 %%
-A qualitative analysis was conducted on a breast fat data set.
-%%
-We were able to derive new hypotheses about relations between breast fat density and breast lesions with breast cancer.
+A qualitative analysis was conducted on a breast density data set.
+%%
+We were able to derive new hypotheses about relations between breast density density and breast lesions with breast cancer.
 %%
@@
 %%
-Therefore we provide template regression formulas, which are applied to all potential feature combinations.
+For this purpose we provide template regression formulas, which are applied to all potential feature combinations.
 %%
@@
 %%
-Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
+\emph{Hypothesis generation} benefits from overview visualizations of feature correlations, which are not supported by standard statistical processors.
 %%
@@
 %%
-There are many data acquisition modalities available to epidemiologists. %, which are suitable for different analysis types.
+There are many epidemiological instruments available to acquire data. %, which are suitable for different analysis types.
 %%
@@
 %%
-These features are gathered through different modalities, such as questionnaires, medical examinations or laboratory analyses.
+These features are gathered through questionnaires, medical examinations and laboratory analyses.
 %%
@@
 %%
-However, this should be avoided, since it reduces the information space.
-%%
-It also introduces a new information bias, as assumptions are modeled through the discretization.
-%%
-
-Medical image data are also acquired and analyzed in modern population studies.
+However, this reduces the information space and introduces a information bias, as assumptions are modeled through the discretization.
+%%
+
+Medical image data are also analyzed in modern population studies.
 %%
@@
 %%
-\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
-%%
+\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases are more likely for older subjects.
+%% TODO Shorten next sentence if needed
 It also influences the general body condition and thereby almost all features acquired through population studies.
@@
 %%
-Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
+Unique for the \texttt{SHIP} is the acquisition of medical image data.
 %%
@@
 %%
-Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
+Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
 %%
@@
 %%
-The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models.
+The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models \cite{Akaike}.
 %%
@@
 %%
-Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
+Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
 %%
@@
 %%
-Existing regression models can be validated and compared using 3D views and 2D slice-views.
+Existing regression models can be validated and compared using 3D views and 2D slice views.
 %%
@@
 %%
-M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
+M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
 %%
@@
 %%
-Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
+The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of visual analysis of cohort study data.
 %%
@@
 %%
-We apply their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}.
-%, meaning that the system has to adapt to different requirements in different analysis steps.
+We incorporate their requirements, which consist of \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
 %%
@@
 %%
-In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insights.
+In contrast, we focus on a fast large scale correlation analysis incorporating many features to derive insights.
 %%
@@
 %%
-They are therefore suitable for \emph{small} epidemiological data sets.
+%They are therefore suitable for \emph{small} epidemiological data sets.
 %%
 \magenta{
-Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlation of features with spatial and non-spatial geriatric data.
+Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
 %%
@@
 %%
-A multi-view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+A multi view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
 %%
@@
 %%
-Their research agenda acts as a guideline for our design.
+Their refer to their research agenda in our design.
 %%
@@
 %%
-We extended the feature selection step (recall Sec.
-\ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
+We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection using expert input as well as clustering methods.
 %%
@@
 \\\\
-The difference of the presented approach in comparison with our previous and related work is twofold.
+The difference of the presented approach in comparison with related work is twofold.
 %%
@@
 %%
-A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
+A regression model is based on expert knowledge.
+%%
+There is no rule how to apply models to a given set of features. Thus, they have to be applied with care.
 %%
@@
 %%
-\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
+\design{Expert knowledge modeling} is carried out using \emph{regression formulas}.
 %Expert knowledge is integrated into the analysis using regression formulas.
 %%
-As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
+The formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target (recall Sect.~\ref{sec:RegressionAnalysis}).
 %%
@@
 %\end{equation}
-would yield $n^2$ (10,000) regression models, describing all possible combinations of two features in the data describing the target feature $Cancer$.
-%%
-This notation comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
+yields $n^2$ (10,000) regression models, describing all combinations of two features describing $Cancer$.
+%%
+This notation is natural to anyone familiar with regression analysis, since it is the standard way of expression.
 %%
@@
 %%
-When each regression takes about 50~ms of calculation time, we roughly have to wait 14~h for the calculation to complete.
-%%
-Therefore, the computational complexity needs to be reduced.
-%%
-An approach for this is presented in the following section.
+When each regression takes about 50~ms of calculation time, the calculation lasts roughly 14~h.
+%%
+%Therefore, the computational complexity needs to be reduced.
+%%
+%An approach for this is presented in the following section.
 %%
@@
 %%
-In general, many of them exhibit a low or no correlation at all in view of the target feature.
+In general, many of them exhibit a low or no correlation at all w.r.t. the target feature.
 %%
@@
 %%
-The CFS algorithm will select the feature, which has the largest explanatory power and discards the other dimensions.
+The CFS algorithm will select the feature, which has the largest explanatory power and discards the other features.
 
@@
 \magenta{
-The tradeoff of the CFS algorithm is the potential removal of features of interest for the domain expert.
-%%
-This problem is handled in the next subsection as part of the 3D representation of the regression results.
+The tradeoff involved using the CFS algorithm is the potential removal of interesting features for the domain expert.
+%%
+This problem is discussed in the next subsection as part of the 3D representation of the regression results.
 }
@@
 %%
-This allows to infer about the predictive quality of the independent features included in the model.
-%%
-As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
+This allows to infer the predictive quality of the independent features included in the model.
+%%
+The $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment (recall Sect.~\ref{sec:RegressionAnalysis}).
 %%
@@
 %%
-This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
+This encodes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
 %%
@@
 %%
-$AIC$ values are not restricted to a normalized scale.
-%%
-They have to be normalized in order to map it onto color saturation.
+$AIC$ values have to be normalized in order to map them onto color saturation.
 %%
@@
 %%
-To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
+To include users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
+%To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
 %%
@@
 %%
-Epidemiologists with background in statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+Epidemiologists argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
 %%
@@
 \magenta{
-As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features important to the domain expert.
-%%
-To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence.
+As mentioned previously, the feature reduction using the CFS algorithm potentially removes important features. %important to the domain expert.
+%%
+The $z$ dimension of the visualization contains \emph{all} features of the data set, allowing to assess their influence.
 %%
@@
 %%
-The \emph{hypothesis-free} analysis allows to derive new insight into the data, such as identifying confounding features or potential targets (e.g.
-deriving risk factors for breast cancer associated features).
+The \emph{hypothesis-free} analysis allows to derive new insights, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
 %%
@@
 
-%%
+%% TODO: First example can be cut if needed
 $Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
@@
  %%
- Breast fat data set loaded into our prototype.
+ Breast density data set loaded into our prototype.
  %%
@@
 %%
-Since they are usually practicing clinicians, they usually have little time to wrangle software.
+They usually have little time to wrangle software.
 %%
@@
 %%
-Epidemiologists have little time slots available for their research.
-%%
-Therefore, complex systems, which require substantial training to work with are not suitable. %for the task of deriving overview visualizations.
+%Epidemiologists have little time slots available for their research.
+Epidemiologists will not adapt complex systems, which require substantial training and time. %to work with are not suitable. %for the task of deriving overview visualizations.
 %%
@@
 %%
-\magenta{Using a check-box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.}
+\magenta{Using a check box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.}
 
@@
 %%
-$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
+%$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
 %%
@@
 %%
-Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
+Along the diagonal, $X$ and $Y$ represent the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
 %%
@@
 %%
-Hence, UI elements for controlling them were introduced to the system.
+Hence, UI elements for controlling them were introduced.
 %%
@@
 %%
-\emph{Adjusted} $R^2$ values are only available for linear regression, logistic regression results are represented via $R^2$ values in this mode.
+\emph{Adjusted} $R^2$ values are only available for linear regression.
+%%
+Logistic regression results are represented via $R^2$ values in this mode.
 %%
@@
 %Since the regression metrics are mirrored along the diagonal, we decided to use this space to support better identification of the current slice.
-Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the heat map diagonal is used to display information about the currently selected plane.
+Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the heat map diagonal is used to display the 2D heat map of the currently selected plane.
 }
@@
 %%
-In is easily expandable by forcing developers to write modularized code.
+It is easily expandable by forcing developers to write modularized code.
 %It forces developers to write modularized code, which makes the components easier expandable. %while keeping the code maintainable by including unit tests.
@@
 %%
-It is widely adopted in the statistical analysis community, yielding a rich support of fast state-of-the-art statistics algorithms as well newly published methods.
+It is widely adopted in the statistical analysis community, yielding a rich support of state-of-the-art statistics algorithms as well newly published methods.
 %%
@@
 %%
-Instructions and code on how to create a setup running the statistical back-end through a \href{http://www.ubuntu.com/}{\texttt{Ubuntu}} server using \href{https://www.opencpu.org/}{\texttt{OpenCPU}} are referenced in the repository.
+Instructions and code to setup running the statistical back-end through a \href{http://www.ubuntu.com/}{\texttt{Ubuntu}} server using \href{https://www.opencpu.org/}{\texttt{OpenCPU}} are included in the repository.
 %%
@@
  %%
- The hypothesis-free analysis of the breast fat data set (right) w.r.t.
+ The hypothesis-free analysis of the breast density data set (right) w.r.t.
 the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
@@
 %%
-Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
-%%
-The breast fat data set is the basis for an explorative analysis w.r.t. the influencing parameters of the parenchyma tissue of the female breast.
+We try to reproduce the prior results from the analysis as proof-of-concept of our method.
+%%
+The female breast density data set is the basis for an explorative analysis w.r.t. the influencing parameters of the breast cancer related parenchyma tissue ratio.
 %%
@@
 %%
-\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast fat data set.
+\emph{KH}, a clinician (10 years of experience) with focus on epidemiological research, is the domain expert for the breast density data set.
 %%
@@
 %%
-An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
+An appropriate cut-off value of 52 years, which is the approximately the entry age for the menopause was set, yielding the most homogeneous class distribution within the resulting subsets.
 %%
@@
 %%
-\subsection{The Breast Fat Data Set}
-The breast fat data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
+\subsection{The Breast Density Data Set}
+%%
+The breast density data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
 %%
@@
 %%
-It contains 231 dimensions, holding information about somatometric features (e.g., body size and weight) consumption behavior, personal and medical history (e.g., occupation and prior diseases), women-specific features (e.g., number of born children and contraception type) as well as mammography features (e.g., fat content and parenchyma tissue proportion to volume).
+It contains 231 features, holding information about somatometric features (e.g., body size and weight) consumption behavior, personal and medical history (e.g., occupation and prior diseases), women-specific features (e.g., number of born children and contraception type) as well as mammography features (e.g., fat content and parenchyma tissue proportion to volume).
 %%
@@
 %%
-Analyzing the 2D heat map [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.
+Analyzing the 2D heat map [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used as renal retention parameter) and \emph{uric acid} (used as gout and diabetes risk factors) magnitudes.
 %%
@@
 %%
-\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
-%%
-The analysis aims to find relationships on the breast fat data using mammography analysis features.
+\subsection{Case 2: Hypothesis-free Analysis of the Breast Density Data Set}
+%%
+The analysis aims to find relationships on the breast density data using mammography analysis features.
 %%
@@
 %%
-Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
+Women with higher \emph{body fat} also have a larger \emph{breast density percentage}, which also correlates with other somatometric features.
 %%
@@
 %%
-The correlation was noted and will be further investigated in the near future using an extensive data set.
+The correlation was noted and will be further investigated using an extensive data set.
 %%
@@
 %%
-We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast fat data set.
+We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast density data set.
 %%

### 2015-06-24 draft.tex
minor revisions

@@
 %%
+\magenta{
+Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
+}
+%%
 The class of the target feature restricts the regression type.
 %%
-Different regression types are available, we focus on the following:
+Different regression types are available, we focus on the following for describing linear relationships:
 %%

### 2015-06-24 draft.tex
revised with Steffis feedback

@@
 	%%
-  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
+  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
 	%%
@@
 % 	%%
-% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
+% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details-on-demand techniques for in-depth investigations of feature relationships.
 % 	%%
@@
 %  %%
-%  The user has the option to either adjust the formula, adjust the transfer function or to derive details on demand on models.
+%  The user has the option to either adjust the formula, adjust the transfer function or to derive details-on-demand on models.
 %  %%
@@
 
-\textbf{Details on demand} [\textbf{4.DD}].
+\textbf{Details-on-demand} [\textbf{4.DD}].
 Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g.
@@
 %%
-\design{The distinction between overview and details on demand using two different representations was well received with our domain experts.}
+\design{The distinction between overview and details-on-demand using two different representations was well received with our domain experts.}
 %%
@@
 %%
-Details on demand for each model allow for a detailed assessment of regression models.
+Details-on-demand for each model allow for a detailed assessment of regression models.
 %%

### 2015-06-25 draft.tex
added revision of Sylvia

@@
 %%
-The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to them being not considered at all.
+The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
 %%
@@
 %%
-A multi view system linked spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
 %%
@@
 %%
-It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
+It allowed them to reproduce their knowledge about relationships by observing correlations they would expect as well as discovering new correlations.
 %%
@@
 %%
-$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
-%%
-$AIC$ values have to be normalized in order to map them onto color saturation.
+\emph{Adjusted} $R^2$ can be represented in the same way, since they share the scale.
+%%
+$AIC$ values have to be normalized in order to map them on color saturation.
 %%
@@
 %%
-Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
+Hence, we inverted the transfer function color mapping, assigning low $AIC$ features to saturated colors.
 %%
@@
 %%
-\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
+\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
 %%
@@
 %%
-Epidemiologists with focus on statistics pointed out, that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
+Epidemiologists with focus on statistics pointed out that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
 %%
@@
 %%
-Investigating more dynamic variables can be achieved by projecting the resulting high-dimensional space into a three-dimensional representation.
+Investigating more dynamic variables can be achieved by projecting the high-dimensional space into a three-dimensional representation.
 %%

### 2015-06-25 draft.tex
added Kais feedback

@@
 	%%
-  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
-	%%
-  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
-	%%
-  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
-	%%
-  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
+  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis generation w.r.t. a target feature using regression models.
+	%%
+  %\item Applying domain knowledge in the analysis using freely adjustable regression formulas.
+	\item Incorporation of prior domain knowledge by using freely adjustable regression formulas.
+	%%
+  %\item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
+	\item Metrics selection for analyzing regression models and for details-on-demand representations.
+	%%
+  %\item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
+	\item An open-source web-application that can be used with data of different application domains.
 	%%
@@
 %%
-The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models \cite{Akaike}.
+The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
 %%
@@
 %%
-The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
+%The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
+The $f$-$statistic$ measures the improvement of the model if independent variables were added. 
 %%
@@
 \magenta{
-Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
+Sedlmair et al.~\cite{Sedlmair} presented a conceptual taxonomy of parameter space analysis.
 %%
@@
 %%
-They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
+They provided a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
 %%
@@
 %%
-M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
+M\"{u}hlbacher et al. focused on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
 %%
@@
 \magenta{
-Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
+Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
 %%
@@
 %%
-A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
+A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
 %%
@@
 \magenta{
-The \emph{3D Regression Heat Map} is designed to provide overview visualizations to support hypothesis generation.
+The \emph{3D Regression Heat Map} is designed to provide an overview visualization to support hypothesis generation.
 %%
@@
 %%
-\emph{Adjusted} $R^2$ can be represented in the same way, since they share the scale.
+\emph{Adjusted} $R^2$ can be represented in the same way, since they are also scaled between $[0, 1]$.
 %%
@@
 %%
-This additionally allows to filter models with desired features, such as only very high $R^2$ values.
+This allows to filter models with desired features, such as only very high $R^2$ values.
 }
@@
 %%
-By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-scores and AIC values}.
+By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-statistics and AIC values}.
 %%

### 2015-06-25 draft.tex
removed color codes

@@
 Our contributions are:
-
-\design{
+%%
 \begin{itemize}
@@
 \end{itemize}
-}
+%%
 % \begin{itemize}
@@
 %%
-\add{
-%%
 The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
@@
 %%
-}
 $R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
@@
 %%
-\magenta{
 Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
-}
 %%
@@
 The \emph{dependent} variable has to be of \emph{continuous type}.
-%%
-\add{
 %%
@@
 %%
-}
-
 \paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
@@
 %%
-\magenta{
 Sedlmair et al.~\cite{Sedlmair} presented a conceptual taxonomy of parameter space analysis.
@@
 We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
-}
-
-\magenta{
+%%
 %% A Partition-based Framework for Building and Validating Regression Models
@@
 %We also take categorical targets and regression formula inputs into account.
-%%
-}
 
@@
 %%
-\magenta{
-%%
 Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
@@
 %%
-}
 %% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
@@
 %%
-\magenta{
 Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
@@
 Instead of applying several linked views, we focus on one view of abstracted data.
-}
 
@@
 %%
-\magenta{
 We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
 %%
-}
+
 %%
@@
 \section{3D Regression Heat Map Analysis of Population Study Data}
-\magenta{
+%%
 The \emph{3D Regression Heat Map} is designed to provide an overview visualization to support hypothesis generation.
@@
 Relationships observed using such techniques are subject of detailed statistical testing by statisticians with background in epidemiology using statistical processors, such as \texttt{SPSS}.
-}
 % \design{
@@
 %%
-\design{
-%%
 The \emph{3D Regression Heat Map} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
@@
 %In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
-}
 %%
@@
 %%
-\design{Expert knowledge modeling} is carried out using \emph{regression formulas}.
+Expert knowledge modeling is carried out using \emph{regression formulas}.
 %Expert knowledge is integrated into the analysis using regression formulas.
@@
 %%
-\magenta{
 The tradeoff involved using the CFS algorithm is the potential removal of interesting features for the domain expert.
@@
 This problem is discussed in the next subsection as part of the 3D representation of the regression results.
-}
 %%
@@
 %%
-The $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment (recall Sect.~\ref{sec:RegressionAnalysis}).
+The $R^2$, adjusted $R^2$ and AIC value are metrics allowing for this kind of assessment (recall Sect.~\ref{sec:RegressionAnalysis}).
 %%
@@
 %%
-\design{
-%%
 Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
@@
 Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
-%%
-}
-%%
-\add{
 %%
@@
 %%
-}
-\design{
-%%
 To include users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
@@
 %%
-}
-
 \paragraph{3D view.}
@@
 %%
-\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
-%%
-\magenta{Object size is not used to encode information because it would result in a cluttered view.}
-%%
-\design{
+The selected metric (per default set to $R^2$) of each heat map entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
+%%
+Object size is not used to encode information because it would result in a cluttered view.
 %%
@@
 %%
-}
-%%
 %The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
@@
 %%
-\magenta{
 As mentioned previously, the feature reduction using the CFS algorithm potentially removes important features. %important to the domain expert.
@@
 The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
-}
 %%
@@
 %%
-\design{
-%%
 Early analysis sessions yielded two approaches of analyzing the data.
@@
 %%
-}
-%%
 %Input formulas reflect \emph{hypotheses} about the data.
@@
 \paragraph{3D Regression Heat Map comparison.}
-\design{
 %%
@@
 Epidemiologists with focus on statistics pointed out that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
-%%
-}
 %%
@@
 %%
-\magenta{This is, however, out of the scope of this work.}
+This is, however, out of the scope of this work.
 %%
@@
 %%
-\design{
 %A key requirement was little set up time to reduce the effort on the domain expert side, as they are mostly practicing clinicians with small time-frames for epidemiological research.
@@
 %%
-}
 %Design choices and spaces are restricted by the underlying technologies and devices.
@@
 %%
-\magenta{
-%%
 This architecture is also prone to security issues, such as the storage of confidential data, especially in the epidemiological context.
@@
 %%
-}
-%%
 %The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
@@
 \subsection{System Paradigm and Components}
-%%
-\design{
 %%
@@
 %%
-}
-%%
 %The \emph{3D Regression Heat Map} design focuses on a clean appearance, reducing the amount of user interface elements as much as possible.
@@
 	%%
-	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
-	%%
-	\item The \emph{Regression Heat Map visualization} consisting of the 2D heat map as well as a 3D representation of all regression models \add{with facilities to change the represented regression metric and its range} [2.SI].
+	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [\textbf{1.AF}].
+	%%
+	\item The \emph{Regression Heat Map visualization} consisting of the 2D heat map as well as a 3D representation of all regression models with facilities to change the represented regression metric and its range [\textbf{2.SI}].
 	%%
@@
 	%%
-	It also allows to select a reference formula for creating difference models [1.AF, 3.ZF].
+	It also allows to select a reference formula for creating difference models [\textbf{1.AF}, \textbf{3.ZF}].
 	%%
@@
 %%
-\magenta{Using a check box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.}
+Using a check box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.
 
@@
 \paragraph{Selecting and scaling the descriptive regression metric.}
-\design{
 %%
@@
 %%
-}
-\add{
-%%
 The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
@@
 This allows to filter models with desired features, such as only very high $R^2$ values.
-}
-
+%%
 \paragraph{3D prism as data mini-map.}
-\design{
 %%
@@
 %%
-}
-%%
-The \emph{3D Regression Heat Map} representation \design{was redesigned to} act as an overview over the whole data set.
+The \emph{3D Regression Heat Map} representation was redesigned to act as an overview over the whole data set.
 %%
@@
 %%
-\design{The distinction between overview and details-on-demand using two different representations was well received with our domain experts.}
+The distinction between overview and details-on-demand using two different representations was well received with our domain experts.
 %%
@@
 %%
-\magenta{Variables are ordered the same way in the 2D and 3D heat map to preserve the mental model and make them visually analogous.}
+Variables are ordered the same way in the 2D and 3D heat map to preserve the mental model and make them visually analogous.
 
@@
 	%%
-	\design{Early prototypes only provided this method to select a slice of interest, which was inefficient when the user was looking for a specific slice.
+	Early prototypes only provided this method to select a slice of interest, which was inefficient when the user was looking for a specific slice.
 	%%
@@
 	%%
-	}
-	%%
 	%\item
@@
 %%
-\design{
 Early prototypes rendered the whole \emph{3D Regression Heat Map}, which made it hard to assess the position of the plane.
@@
 Since the regression metrics are mirrored along the diagonal, the space available from visualizing only the prism generated from the upper half of the heat map diagonal is used to display the 2D heat map of the currently selected plane.
-}
-%%
-The \magenta{regression metric values} are projected on this plane to provide an overlapping-free view.
+%%
+The regression metric values are projected on this plane to provide an overlapping-free view.
 %%
@@
 %%
-By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-statistics and AIC values}.
-%%
-\add{
+By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals f-statistics and AIC values.
 %%
@@
 Epidemiologists use such plots to validate models w.r.t. the model assumptions, such as homogeneity, normality, and independence \cite{GraphicModelAssessment}.
-}
-
+%%
 \section{Implementation} \label{implementation}
@@
 
-\magenta{
+%%
 %One limitation of the proposed method is that the regression metrics always only capture a part of the underlying model.
@@
 Static features can be added using the formula input without increasing the complexity of the visualization.
-}
+%%
 

### 2015-06-26 draft.tex
added Steffis revision

@@
 %%
-We were able to derive new hypotheses about relations between breast density density and breast lesions with breast cancer.
+We were able to derive new hypotheses about relations between breast density and breast lesions with breast cancer.
 %%
@@
 %%
-Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
+Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS}, are not tailored to a specific target feature.
 %%
@@
   %\item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
-	\item An open-source web-application that can be used with data of different application domains.
+	\item An open-source web application that can be used with data of different application domains.
 	%%
@@
 %%
-As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
+As described by Thew et al.~\cite{Thew2009}, the epidemiological workflow is divided into these different steps:
 \begin{enumerate}
-	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
+	\item Clinicians/epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
 	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
@@
 %%
-Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
+Reproducibility of results is an epidemiological key requirement and guides all analysis steps.
 %%
@@
 %%
-However, this reduces the information space and introduces a information bias, as assumptions are modeled through the discretization.
+However, this reduces the information space and introduces an information bias, as assumptions are modeled through the discretization.
 %%
@@
 %%
-A second cohort, \texttt{SHIP-TREND} was started in 2012.
+A second cohort, \texttt{SHIP-TREND}, was started in 2012.
 %%
@@
 %%
-Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
+Regression analysis is the most important statistical tool when analyzing epidemiological data and is the basis of this work.
 %%
@@
 %%
-Different regression types are available, we focus on the following for describing linear relationships:
+Different regression types are available; we focus on the following for describing linear relationships:
 %%
@@
 %% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
-Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
+Tukey already stated in 1977 that data are too often analyzed solely using a confirmatory data analysis \cite{Tukey}.
 %%
@@
 %%
-In this section we present prior and related work trying to achieve this goal.
+In this section, we present prior and related work trying to achieve this goal.
 %%
@@
 %%
-This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
+This limits the method to comparing a few models at once, as the plot matrix gets complex with increasing feature number.
 %%
@@
 %%
-The model view allows for assessing different models by color coding distances to the line-of-fit.
-%%
-The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
+The model view allows for assessing different models by color-coding distances to the line-of-fit.
+%%
+The model parameters can then be fine-tuned using line graphs, histograms and model projections.
 %%
@@
 %%
-The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of visual analysis of cohort study data.
-%%
-They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
+The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of a visual analysis of cohort study data.
+%%
+They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of a cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
 %%
@@
 %%
-We incorporate their requirements, which consist of \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
+We incorporate their requirements, which consist of a \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
 %%
@@
 %% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
-Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
+Steenwijk et al.~\cite{Steenwijk} focus on the \emph{hypothesis-free} exploration of cohort data.
 %%
@@
 
-Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
+Generalized pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
 %%
@@
 %% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
-Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
+Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high-dimensional data.
 %%
@@
 %% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
-Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
+Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high-dimensional data into subspaces, which are then described by their most reliable linear regression model.
 %%
@@
 %%
-They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
+They focus on an algebraic representation for efficient regression model calculation to find the best fit for a subspace.
 %%
@@
 %% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
-Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
+Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high-dimensional data sets with brushing facilities to create subsets.
 %%
@@
 %%
-We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
+We try to reproduce their results and further investigate findings presented by Niemann et al.~\cite{Niemann2014}.
 %%
@@
 %%
-(1) We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+(1) We focus on the large-scale analysis of a vast number of linear and logistic regression models by assessing their quality-of-fit using descriptive metrics.
 %%
@@
 % 	%%
-% 	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
+% 	\item We focus on the large-scale analysis of a vast number of linear and logistic regression models by assessing their quality-of-fit using descriptive metrics.
 % 	%%
@@
 %%
-For explorative analysis, they preferred an approach, which highlights associations w.r.t.
+For explorative analysis, they preferred an approach which highlights associations w.r.t.
 various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
@@
 %%
-The CFS algorithm will select the feature, which has the largest explanatory power and discards the other features.
+The CFS algorithm will select the feature which has the largest explanatory power and discards the other features.
 
@@
 %%
-Epidemiologists argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
+Epidemiologists argued that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
 %%
@@
 %%
-The \emph{hypothesis-free} analysis allows to derive new insights, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
+The \emph{hypothesis-free} analysis allows to derive new insights, such as identifying confounding features or potential targets (e.g. deriving risk factors for breast cancer-associated features).
 %%
@@
 %%
-Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
+Using the operators, dynamic variables and data set features, many different assumptions can be expressed.
 %%
@@
 %% TODO: First example can be cut if needed
-$Cancer \sim X + Y + Z$ is the formulation of a hypothesis, where the specific feature \textit{Cancer} is analyzed.
+$Cancer \sim X + Y + Z$ is the formulation of a hypothesis where the specific feature \textit{Cancer} is analyzed.
 %%
@@
  %%
- (b) 3D heat map, showing values above the matrix diagonal as overview.
+ (b) 3D heat map showing values above the matrix diagonal as overview.
  %%
@@
 %Epidemiologists have little time slots available for their research.
-Epidemiologists will not adapt complex systems, which require substantial training and time. %to work with are not suitable. %for the task of deriving overview visualizations.
+Epidemiologists will not adapt complex systems that require substantial training and time. %to work with are not suitable. %for the task of deriving overview visualizations.
 %%
@@
 %%
-To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
+To allow other users even outside the epidemiological application domain to access our tool, we use standard ASCII-based CSV files.
 %%
@@
 %%
-Using a check box, the user can disable the CFS preprocessing step, which is useful for small data sets, where the user does not want to reduce the number of features.
+Using a check box, the user can disable the CFS preprocessing step, which is useful for small data sets where the user does not want to reduce the number of features.
 
@@
 %%
-The visualization is sparse, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
+The visualization is sparse, since the majority of the regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
 %%
@@
 	%%
-	We, however, still display the whole 3D object instead of cutting away information. %down to the slice position.
+	However, we still display the whole 3D object instead of cutting away information. %down to the slice position.
 	%%
@@
 %%
-To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation-based feature selection.
+To reduce visual clutter, the 2D view only shows dimensions which are retrieved through the correlation-based feature selection.
 %%
@@
 %%
-By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals f-statistics and AIC values.
+By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals, f-statistics and AIC values.
 %%
@@
  %%
- The web-server (right) is written using \href{http://nodejs.org}{\texttt{Node.js}} and hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}.
+ The web server (right) is written using \href{http://nodejs.org}{\texttt{Node.js}} and hosted on \href{https://www.heroku.com/}{\texttt{Heroku}}.
  %%
@@
 %%
-%The web-server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
-The web-server is written in \texttt{Javascript} using \href{http://nodejs.org}{\texttt{Node.js}}, running on Googles V8 Javascript runtime environment.
+%The web server is written in \texttt{Javascript} using \texttt{Node.js}\footnote{Open Source; Maintained by Joyent Inc, \href{http://nodejs.org}{\texttt{nodejs.org}}}, running on Googles V8 Javascript runtime environment.
+The web server is written in \texttt{Javascript} using \href{http://nodejs.org}{\texttt{Node.js}}, running on Googles V8 Javascript runtime environment.
 %%
@@
 %%
-This way, any computer, which runs \href{http://r-project.org}{\texttt{R}} can be turned into a statistical processor for our project.
+This way, any computer which runs \href{http://r-project.org}{\texttt{R}} can be turned into a statistical processor for our project.
 %%
@@
 %%
-The female breast density data set is the basis for an explorative analysis w.r.t. the influencing parameters of the breast cancer related parenchyma tissue ratio.
+The female breast density data set is the basis for an explorative analysis w.r.t. the influencing parameters of the breast cancer-related parenchyma tissue ratio.
 %%
@@
 %%
-The \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique proposed by Lam et al.~\cite{Lam2012} is focused on the characterization of a systems' ability to generate hypotheses and explore the data in order to extract information.
-%%
-\emph{VDAR} can be carried out based on case studies using thinking-aloud techniques to comprehend the users reasoning and thought process.
+The \emph{Visual Data Analysis and Reasoning} (\emph{VDAR}) technique proposed by Lam et al.~\cite{Lam2012} is focused on the characterization of a system's ability to generate hypotheses and explore the data in order to extract information.
+%%
+\emph{VDAR} can be carried out based on case studies using thinking-aloud techniques to comprehend the user's reasoning and thought process.
 %%
@@
 %%
-We conducted a web-based analysis by using an online-meeting software, which features voice chat as well as screen sharing.
+We conducted a web-based analysis by using an online meeting software, which features voice chat as well as screen sharing.
 %%
@@
 %%
-Afterwards, the experts use the system on their own computers.
-%%
-The screen-sharing function was still used to observe the actions of the expert.
-%%
-All sessions were video recorded to be processed later on.
+Afterwards, the experts used the system on their own computers.
+%%
+The screen-sharing function was still used to observe the actions of the experts.
+%%
+All sessions were video-recorded to be processed later on.
 %%
@@
 %%
-She is a radiologist, responsible for the SHIP-MRI acquisition and also responsible of the mammography analysis.
+She is a radiologist responsible for the SHIP-MRI acquisition and also for the mammography analysis.
 %%
@@
 %%
-The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
+The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience), who assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
 %%
@@
 %%
-Apart from the target feature, the data set contains 199 features, comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol and tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
+Apart from the target feature, the data set contains 199 features comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol and tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
 %%
@@
 %%
-An appropriate cut-off value of 52 years, which is the approximately the entry age for the menopause was set, yielding the most homogeneous class distribution within the resulting subsets.
+An appropriate cut-off value of 52 years, which is the approximate entry age for the menopause was set, yielding the most homogeneous class distribution within the resulting subsets.
 %%
@@
 
-The data of each cohort was presented as individual \texttt{SPSS} files.
+The data of each cohort were presented as individual \texttt{SPSS} files.
 %%
@@
 %%
-The finding is subject of further analysis.
+The finding is subject to further analysis.
 %%
@@
 %%
-Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. its strong influence to the parenchyma tissue.
-%%
-A surprising finding was the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
+Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. its strong influence on the parenchyma tissue.
+%%
+A surprising finding was the strong correlation with \emph{kidney disorder} ($R^2$ values around $0.9$).
 %%
@@
 %%
-Lastly, \emph{KH} assessed the influence of contraception-related features, such as use of \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
+Lastly, \emph{KH} assessed the influence of contraception-related features, such as the use of \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
 %%
@@
 %%
-Results from automatic `black-box' methods, such as data mining algorithms are more often obscure to the expert.
+Results from automatic `black-box' methods, such as data mining algorithms, are more often obscure to the expert.
 %%
@@
 %%
-This, however, increases the cognitive load and complexity of the analysis substantially and needs to be accompanied by techniques, which simplify this approach.
+This, however, increases the cognitive load and complexity of the analysis substantially and needs to be accompanied by techniques that simplify this approach.
 %%
@@
 We want to support opening up knowledge discovery to allow a diverse group of domain experts to derive insight into their data. %and support the notion of open science.
-%%
+
 \begin{small}
@@
 %%
-Sylvia Gla{\ss}er and Kai Lawonn are funded by the Federal Ministry of Education and Research within the Forschungs-campus STIMULATE under grant number '13GW0095A'.
+Sylvia Gla{\ss}er and Kai Lawonn are funded by the Federal Ministry of Education and Research within the Forschungscampus STIMULATE under grant number '13GW0095A'.
 }

### 2015-06-27 draft.tex
minor last revisions

@@
 %%
-The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of a visual analysis of cohort study data.
+The work of Zhang et al.~\cite{Gotz2014, Zhang2014} is closest to ours regarding the application of a visual analysis of population study data.
 %%
@@
 
-As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
+Regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data (recall Sect.~\ref{sec:RegressionAnalysis}).
 %%
@@
 %%
-\noindent Our \emph{3D Regression Heat Map} model is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
+\noindent Our \emph{3D Regression Heat Map} is well suited for different workflow analysis techniques, based on the Visual Analytics (VA) Mantra of Keim et al.~\cite{Keim}:
 
@@
 %%
-Starting an analysis using these techniques takes about 5-10 minutes of setup time.
+Starting an analysis using these techniques took about 5-10 minutes of setup time.
 %%
@@
 %%
-Apart from the target feature, the data set contains 199 features comprising sociodemographic features (gender, age), consumption behavior (e.g., alcohol and tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
+Apart from the target feature, the data set contains 199 features comprising sociodemographic features (e.g., gender, age), consumption behavior (e.g., alcohol and tobacco), laboratory data (e.g., sera concentrations), and two features depicting the liver ultrasound.
 %%

### 2015-06-27 draft.tex
consistent naming of Three.js

@@
  \caption{
- The front-end (left) is realized with \texttt{HTML5}/\texttt{CSS3}/\texttt{Javascript} and different Javascript libraries, such as \href{https://www.angularjs.org/}{\texttt{Angular.js}}, \href{http://threejs.org}{\texttt{Threejs}} and \href{http://d3js.org/}{\texttt{D3.js}}.
+ The front-end (left) is realized with \texttt{HTML5}/\texttt{CSS3}/\texttt{Javascript} and different Javascript libraries, such as \href{https://www.angularjs.org/}{\texttt{Angular.js}}, \href{http://threejs.org}{\texttt{Three.js}} and \href{http://d3js.org/}{\texttt{D3.js}}.
  %%
@@
 %The 3D plot is created using the WebGL-based \texttt{Threejs}\footnote{Open Source; Originally developed by R. Cabello, \href{http://threejs.org}{\texttt{threejs.org}}} library.
-The 3D plot is created using the WebGL-based \href{http://threejs.org}{\texttt{Threejs}} library.
+The 3D plot is created using the WebGL-based \href{http://threejs.org}{\texttt{Three.js}} library.
 %%
